### 2021-03-05 analysis/scripts/get_data/creating_input_data.Rmd
Initial commit

---
title: "Creating Haul-out Timeline Input Data"
description: |
  document the processing of the 'raw' telemetry data into a usable 
  data structure for downstream analysis procedures
author:
  - name: Josh M. London 
    url: https://github.com/jmlondon
    orcid_id: 0000-0003-0645-5666
    affiliation: NOAA Alaska Fisheries Science Center
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    theme: theme.css
params:
   make_data: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(conflicted)
library(rmarkdown)
library(odbc)
library(DBI)
library(dplyr)
conflict_prefer("filter", "dplyr")
conflict_prefer("selec", "dplyr")
library(dbplyr)
library(pingr)
library(keyringr)
library(lubridate)
conflict_prefer("month","lubridate")
library(sf)
library(glue)
library(tibble)
library(tidyr)
library(purrr)
library(forcats)
library(gt)
library(here)
library(ggplot2)
```

## Introduction

The focus of this document is not to analyze the data, but instead to document the
processing of the 'raw' telemetry data into a usable data structure for downstream
analysis procedures in support of the manuscript *\<insert manuscript title here\>*. The
key focus of this step is to merge separate data streams for estimated seal locations,
observed seal haul-out behavior (in the form of percent-dry timelines), and capture
details into a single table ready for ingestion to the PostgreSQL database. This table
should, generally, provide hourly percent-dry records for each seal along with reasonable
location estimates. We will rely on a separate process within the database for, then,
assigning reanalysis weather covariates to each record in this spatial time series.

## Workflow

This process is managed with a `make_data.R` script. We rely on a series of functions to
pull data from the database and, then, tidy and merge into our needed data structure. We
will briefly describe each functional step below. Each function can be examined in detail
by opening the corresponding R file within the R directory.

```{r}
if (params$make_data) {
source(here::here('analysis/scripts/get_data/make_data.R'))
} else {
  load(here::here('data/locs_sf.Rdata'))
  load(here::here('data/timeline_data.Rdata'))
  load(here::here('data/tbl_percent_locs.Rdata'))
}
```

### `get_timeline_data()`

This is a relatively simple pull and merge of telemetry data from the PEP PostgreSQL
database. The focus is on the *percent_dry* timeline data and joining those records with
additional details regarding seal species, age, sex, etc. We also filter the data to only
include records for *Bearded seals*, *Ribbon seals*, and *Spotted seals*. Also, only
records from March, April, May, June, and July are retained. The last step is to create a
*unique_day* column that represents the unique combination of year and day-of-year. Since
we will be calculating daily location averages, this field will be crucial for properly
joining the timeline data with those locations.

The final data structure:

```{r}
glimpse(timeline_data)
```

### `get_locs_sf()`

This is also a simple pull and merge of telemetry data from PostgreSQL database. The focus
here is on location estimates. As such, we rely on the `sf::read_sf()` function and a
custom query:

``` {.sql}
SELECT  deployid,
        locs_dt,
        type,
        quality,
        error_radius,
        geom
FROM telem.geo_wc_locs_qa
```

After this query, the process is similar to the `get_timeline_data()` function. We merge
with the additional details (e.g. seal species, age, sex) and filter to only include our
months of interest (March, April, May, June). Note we also calculate the *unique_day*
field that will become the basis for our location averaging.

The final data structure:

```{r}
glimpse(locs_sf)
```

### `create_db_input_data()`

This is the workhorse function for this part of our workflow. We need to accomplish a few
key objectives with this function:

1.  determine an appropriate error radius value for each location record

    1.  for records with only Argos location quality, use standard error values

    2.  for records with error_radius values, prefer over Argos location quality values

2.  transform the geographic coordinates to the *LAEA Bering Sea* / *EPSG:3571* projection

3.  calculate a weighted mean daily x and y coordinate where the weights are
    error_radius\^-1

4.  full join of the timeline and daily location records

5.  fill any missing coordinates with the most recent known daily location

6.  write the result back to the PEP PostgreSQL database

Our data represents deployments across several years and differing data formats for Argos
locations. Only the newer (since \~2010) deployments provide an record-specific estimate
of error radius (also, error ellipse information but we are simplifying the approach here
and just using radius). Earlier deployments only provide the location quality values

The following table represents our map of Argos location quality values to estimated error
radius (in meters).

```{r}
tribble(
    ~quality, ~error_radius,
    "3", 250,
    "2", 500,
    "1", 1500,
    "0", 2500,
    "A", 2500,
    "B", 2500
  ) 
```

Once we have an error radius value for each record, we can proceed to transform our
spatial coordinates from geographic (latitude, longitude) into an equal-area projection
appropriate for the Bering Sea. We chose the LAEA Bering Sea (espg:3571) projection. This
transforms our spatial coordinates into units of meters and allows for a straightforward
calculation of daily mean coordinates. Since we have our error radius information, we can
take a more thorough approach to calculation of our daily mean and apply the inverse of
the error radius value as a weight to our mean calculation.

At this point, we have two tables: daily, weighted mean locations and hourly percent-dry
values. A full join of these two tables based on seal identification (i.e. *speno*) and
the *unique_day* provides an full table of all timeline records and, where there is a
matching daily mean location, spatial estimates of seal location. In more than a few
cases, there are observed percent-dry data without a corresponding location for that same
day. In these situations, the previously known location is carried forward to fill in the
unknown days.

## Data

First, the table structure:

```{r}
glimpse(db_input_data)
```

Nothing unexpected here. The key is that this dataset is a Simple Features Collection with
each record represented by a geospatial point. This geometry along with the haul-out
date-time data will allow us to properly link these records with spatially explicit
weather covariates.

Next, let's summarize the number deployments for each species by age and sex along with
the number of seal hours represented. This table allows us to compare the spread and
amount of data for each with our expectations.

```{r}
db_input_data %>% 
  st_set_geometry(NULL) %>% 
  mutate(age = dplyr::case_when(
      age == "YEARLING" ~ "SUBADULT",
      TRUE ~ age
    ),
    sex = dplyr::case_when(
      sex == "F" ~ "Female",
      sex == "M" ~ "Male",
      TRUE ~ sex
    )) %>% 
  mutate(age = fct_relevel(age, "PUP","YOUNG OF YEAR","SUBADULT","ADULT")) %>%  
  dplyr::mutate(data_year = lubridate::year(haulout_dt)) %>% 
  dplyr::group_by(speno,data_year,species, sex, age) %>% 
  dplyr::summarise(min_dt = min(haulout_dt),
            max_dt = max(haulout_dt),
            n_days = difftime(max_dt, min_dt, units = "days"),
            n_hours = n()
  ) %>% 
  dplyr::select(-c("min_dt","max_dt")) %>% 
  dplyr::group_by(species, sex, age) %>% 
  dplyr::summarise(
    n = n(),
    seal_hours = sum(n_hours)
  ) %>% 
  dplyr::mutate(
    deploy_stats = glue::glue("{n} ({seal_hours} seal hours)")
  ) %>% 
  dplyr::select(-c("n","seal_hours")) %>% 
  tidyr::pivot_wider(names_from = age, values_from = c(deploy_stats)) %>% 
  group_by(species) %>% 
  gt(rowname_col = "sex") %>% 
  tab_header(
    title = "Summary of bio-logger deployment data across seal 
  species, sex, and age classification during the months of March, April, May, 
  and June.",
    subtitle = "total seal hours represents the sum of available data across all
  seals"
  ) %>% 
  tab_options(
    table.border.top.width = px(0),
    heading.title.font.size = px(16),
    column_labels.border.top.width = 3,
    column_labels.border.top.color = "black",
    column_labels.border.bottom.width = 3,
    column_labels.border.bottom.color = "black",
    table_body.border.bottom.color = "black",
    table.border.bottom.width = px(0),
    table.width = pct(100),
    table.background.color = "white",
    row_group.border.top.color = "black",
    row_group.border.bottom.width = px(0)
  ) %>% 
  tab_style(
    style = list(
      # remove horizontal lines
      cell_borders(
        sides = c("top", "bottom", "right"),
        weight = px(0)
      )),
      #do this for all columns and rows
    locations = list(cells_body(
      columns = everything(),
      rows = everything()),
    cells_stub())
    ) %>% 
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_row_groups()
  )
```

It's also helpful to visually examine the temporal spread of the data to make sure it
meets our expectations and to identify any temporal limits we might need to consider
during the analysis process.

```{r, fig.height=6.5, fig.width=8}
seal_hours_tbl <- db_input_data %>% 
  sf::st_set_geometry(NULL) %>% 
  dplyr::mutate(wkofyr = lubridate::epiweek(haulout_dt),
                dayofwk = lubridate::wday(haulout_dt, label = TRUE, abbr = TRUE) %>% 
                  forcats::fct_rev()
  ) %>% 
  dplyr::group_by(species, wkofyr, dayofwk) %>% 
  dplyr::count()

ggplot(data = seal_hours_tbl) + 
  geom_point(aes(x = wkofyr, y = dayofwk, color = n, size = n)) +
  scale_color_distiller(palette = "Blues", direction = 1) +
  scale_x_continuous(breaks = c(13,18,23),
                     labels = c("April", "May", "June")) +
  guides(color = "none",
         size = guide_legend(override.aes = list(color = "#3182bd"),
                             title.position = "bottom",
                             title.hjust = 0.5)) +
  facet_grid(species ~ .
             # labeller=labeller(species = c(Eb = "Bearded",
             #                               Hf = "Ribbon",
             #                               Pl = "Spotted"))
             ) + 
  theme_minimal(12) + 
  theme(axis.text.y = element_blank(),
        panel.grid = element_blank()) +
  theme(plot.title.position = "plot") +
  theme(legend.position = "bottom") +
  labs(x = "", y = "day of week", color = "seal-hours",
       size = "seal hours") +
  ggtitle("Distribution in seal-hours of haul-out behavior data",
          subtitle = "collated across years (2005-2017)")
```

lastly, let's plot a map of our data so we can get a sense of the geographic distribution
for each species

```{r}
haulout_locs <- db_input_data %>% 
  filter(lubridate::hour(haulout_dt) == 12)

hexgrid <- sf::st_make_grid(st_bbox(haulout_locs) %>% st_as_sfc(), cellsize = 50*1000,
                            what = "polygons", square = FALSE) 
hexgrid <- st_sf(index = 1:length(lengths(hexgrid)), hexgrid)

hexbin <- st_join(haulout_locs, hexgrid, join = st_intersects) %>% 
  group_by(species)

locs_count <- hexgrid %>%
  left_join(
    count(hexbin, index) %>%
      as_tibble() %>%
      dplyr::select(index, ct=n)
  ) %>% drop_na()

ggplot() +
  geom_sf(
    data = locs_count, size = 0.125,
    aes(fill = ct)
  ) +
  scale_fill_viridis_c(
    option = "magma", 
    trans = "log10", aesthetics = "fill",
    name = "# locations"
  )

```

We can see that there are two locations well outside the study area. these are likely a
result of very poor quality Argos locations and the averaging used to calculate daily
locations for this analysis. We can remove these by filtering based on the speno
(*PL2018_1002*) and the two dates (*2018-06-04* and *2018-06-06*).

```{r}
haulout_locs <- db_input_data %>% 
  filter(lubridate::hour(haulout_dt) == 12,
         speno != 'PL2018_1002' & !lubridate::date(haulout_dt) %in% c('2018-06-04','2018-06-06'))

hexgrid <- sf::st_make_grid(st_bbox(haulout_locs) %>% st_as_sfc(), cellsize = 50*1000,
                            what = "polygons", square = FALSE) 
hexgrid <- st_sf(index = 1:length(lengths(hexgrid)), hexgrid)

hexbin <- st_join(haulout_locs, hexgrid, join = st_intersects)

locs_count <- hexgrid %>%
  left_join(
    count(hexbin, index, species) %>%
      as_tibble() %>%
      dplyr::select(species, index, ct=n)
  ) %>% drop_na()
```

With the erroneous locations removed, we can plot the distribution of haulout locations
for each of our three species

```{r, fig.height=3, fig.width=8}
library("rnaturalearth")
library("rnaturalearthdata")
library(ggspatial)

world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(st_crs(locs_count))

ggplot() +
  annotation_spatial(world) +
  layer_spatial(
    data = locs_count, size = 0.125,
    aes(fill = ct)
  ) +
  scale_fill_viridis_c(
    option = "magma", 
    trans = "log10", aesthetics = "fill",
    name = "# locations"
  ) + facet_grid(~ species)
```