### 2021-05-07 data_raw/adfg/adfg_tidydata.Rmd
ADDED ADDITIONAL ADFG DATA SOURCES
up to this point, some ADFG data sources were not included in the analysis. This is the initial attempt to include those data.

---
title: "Import and Tidy ADFG Supplied Location and Haul-out Data"
description: | 
  This article describes the import and tidying of location and haul-out data provided to
  MML by ADFG for use in manuscript analyzing haul-out behavior of bearded, ribbon, and
  spotted seals in the Bering and Chukchi seas
date: May 6, 2021
author:
  - name: "Josh M. London"
    url: https://github.com/jmlondon
    affiliation: Alaska Fisheries Science Center
    affiliation_url: https://www.fisheries.noaa.gov/about/alaska-fisheries-science-center
    orcid_id: 0000-0002-3647-5046
  - name: "Justin Crawford"
    url: http://www.adfg.alaska.gov/index.cfm%3Fadfg%3Dmarinemammalprogram.icesealresearch
    affiliation: Alaska Department of Fish & Game
    affiliation_url: http://www.adfg.alaska.gov
output: distill::distill_article
---

```{r packages, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
```

## Introduction

The focus of this document is not to analyze the data, but instead to document
the processing of the 'raw' data (as provided) into a usable data structure for
downstream merging with other datasets and analysis procedures. We will focus
on reading in the csv files, tidying any data structures or values that need it,
and then merging into a single usable tibble.

## Read the CSV Files

There are three files provided by ADF&G

1. ADFG_SealDeploymentTable3.csv
2. ADFG_SealLocationData3.csv
3. ADFG_SealTimelineData3.csv

### ADFG Seal Deployments

Let's read this file in and take a look at the structure

```{r}
adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv')
```

Let's look at some of the data and see if we might want to adjust some of the
automatically determined column specifications

```{r}
glimpse(adfg_deployments)
```

Looks like the only thing we would want to change is the `DateTagged` column.
This is clearly a date-time field, so we'll create a custom `readr::cols`
function to specify our column types. Note, we will then set the proper time
zone to *America/Anchorage* and then convert to a `Date` object.

```{r}
adfg_deploy_cols <- readr::cols(
  DeployIDs = readr::col_character(),
  MapID = readr::col_character(),
  DateTagged = readr::col_datetime(format = "%m/%d/%Y %H:%M"),
  Species = readr::col_character(),
  Age = readr::col_character(),
  Sex = readr::col_character(),
  PrimaryTagType = readr::col_character(),
  FlipperTagType = readr::col_character()
)

adfg_deployments <- readr::read_csv('ADFG_SealDeploymentTable3.csv',
                                    col_types = adfg_deploy_cols) %>% 
  dplyr::mutate(DateTagged = lubridate::force_tz(DateTagged, 
                                                 'America/Anchorage')) %>% 
  dplyr::mutate(DateTagged = lubridate::as_date(DateTagged, 
                                                tz = 'America/Anchorage'))

glimpse(adfg_deployments)
```

The first author (JML) likes to stick with snake case for column names and variables.
So, we'll do some additional editing of the data to rename columns. While at it,
we'll also convert some of the columns to lower case to limit typo issues down the
road.

Another issue we need to tidy up is the way satellite tag type and attachment
position is recorded in the dataset. These data are stored within two columns:
`primary_tag_type` and `flipper_tag_type`. For a proper tidy structure, we need
to get these data into two columns: `tag_position` and `tag_type`. To do this,
we'll use the `tidyr::gather` function.

```{r}
adfg_deployments <- adfg_deployments %>% 
  dplyr::rename(deployid = DeployIDs,
                mapid = MapID,
                date_tagged = DateTagged,
                primary_tag_type = PrimaryTagType,
                flipper_tag_type = FlipperTagType
                ) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::mutate(species = tolower(species),
                age = tolower(age),
                sex = tolower(sex)) %>% 
  tidyr::gather(tag_position, tag_type, c(primary_tag_type,flipper_tag_type)) %>% 
  dplyr::mutate(tag_position = ifelse(tag_position == "primary_tag_type",
                   "primary","flipper")) %>% 
  dplyr::filter(!is.na(tag_type)) %>% 
  dplyr::arrange(date_tagged,deployid)

glimpse(adfg_deployments)
```

### ADFG Seal Locations

Next, we'll go through a similar process with the locations data.

```{r}
adfg_locations <- readr::read_csv('ADFG_SealLocationData3.csv')
```

```{r}
glimpse(adfg_locations)
```

As with the deployment data, the `Date` column needs to be converted to a proper
date-time type. Additionally, we can see that `Ptt` was brought in as a `double`
when, in fact, this is an `integer`. There is no way to do the `Date` or `Ptt` type conversion via the `readr::col_*` functions, so we will do a mutate after 
reading the data in.

```{r}
adfg_locs_cols <- readr::cols(
  DeployID = readr::col_character(),
  LocationQuality = readr::col_character(),
  Date = readr::col_double(),
  Latitude = readr::col_double(),
  Longitude = readr::col_double(),
  Ptt = readr::col_double(),
  TagType = readr::col_character(),
  MonthNumb = readr::col_character()
)

adfg_locations <- readr::read_csv('ADFG_SealLocationData3.csv',
                                  col_types = adfg_locs_cols) %>% 
  dplyr::mutate(Date = as.Date(Date, origin = "1899-12-30", tz ="UTC") %>% 
                  lubridate::as_datetime(),
                Ptt = as.integer(Ptt))

glimpse(adfg_locations)
```

As before, we'll rename a few columns to snake_case and then all to lower case.

```{r}
adfg_locations <- adfg_locations %>% 
  dplyr::rename(deployid = DeployID,
                location_quality = LocationQuality,
                date_time = Date,
                tag_type = TagType) %>% 
  dplyr::filter(location_quality != "Z") %>% 
  dplyr::select(-MonthNumb) %>% 
  dplyr::rename_all(tolower)

glimpse(adfg_locations)
```

At this point, we have our locations data assembled. Let's take a moment and plot
all the data on a map

```{r}
library(sf)
library(rnaturalearth)
library(ggspatial)

world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(3571)

locs_sf <- adfg_locations %>% 
  dplyr::left_join(adfg_deployments, by = c("deployid","tag_type")) %>% 
  sf::st_as_sf(coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(4326) %>% sf::st_transform(3571)

ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(
    data = locs_sf, size = 0.125
  ) +
  facet_grid(~ species) +
  theme(legend.position = "bottom") +
  ggtitle("ADFG Locations (Argos)",
            subtitle = "all locations from bearded, ringed, and spotted seals")

```

### ADFG Seal Timelines

Last, we'll go through a similar process with the timeline data.

```{r}
adfg_timelines <- readr::read_csv('ADFG_SealTimelineData3.csv')
```

For this dataset, we'll need to convert the `GMTDate` to a proper date-time
format. And, as before the same with the `Ptt` as an integer.

```{r}
adfg_timeline_cols <- cols(
  .default = readr::col_double(),
  DeployIDs = readr::col_character(),
  TagType = readr::col_character(),
  HistType = readr::col_character(),
  MonthNumb = readr::col_character(),
  LocationQuality = readr::col_character(),
  Sum = readr::col_character()
)

adfg_timelines <- readr::read_csv('ADFG_SealTimelineData3.csv',
                                  col_types = adfg_timeline_cols) %>% 
  dplyr::mutate(GMTDate = as.Date(GMTDate, origin = "1899-12-30", 
                                     tz = "UTC") %>% 
                  lubridate::as_datetime(),
                Ptt = as.integer(Ptt),
                TagType = toupper(TagType)) %>% 
  dplyr::rename(deployid = DeployIDs,
                tag_type = TagType,
                hist_type = HistType,
                date_time = GMTDate) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(-c(monthnumb,latitude,longitude,locationquality,sum,hist_type))

```

We need to do some additional processing of the timeline data to work with the
odd non-tidy data structure. In this case, each `bin*` column represents an hour
of each day.

```{r}
bins <- list(bin = paste("bin", 1:24, sep=""), hour=0:23) %>% as_tibble()

adfg_timelines <- adfg_timelines %>% 
        tidyr::gather(bin, percent_dry, starts_with('bin')) %>%
        dplyr::left_join(bins, by = "bin") %>%
        dplyr::mutate(date_time = date_time + lubridate::hours(hour),
                      data_date = lubridate::floor_date(date_time,"day")) %>%
        dplyr::select(c(deployid,ptt,tag_type,date_time,data_date,percent_dry)) %>% 
        dplyr::arrange(deployid,date_time)
```

```{r}
glimpse(adfg_timelines)
```

```{r}
usethis::use_data(adfg_deployments, overwrite = TRUE)
usethis::use_data(adfg_timelines, overwrite = TRUE)
usethis::use_data(adfg_locations, overwrite = TRUE)
```