### 2021-08-12 analysis/paper/paper.Rmd
CONTINUED UPDATES TO PAPER/FIGURES

---
title: "Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas"
author:
  - Josh M. London:
      email: josh.london@noaa.gov
      institute: [MML]
      correspondence: true
  - Paul B. Conn:
      institute: [MML]
      correspondence: false
  - Stacie K. Hardy:
      institute: [MML]
      correspondence: false
  - Erin L. Richmond:
      institute: [MML]
      correspondence: false
  - Jay M. Ver Hoef:
      institute: [MML]
      correspondence: false
  - Michael F. Cameron:
      institute: [MML]
      correspondence: false
  - Peter L. Boveng:
      institute: [MML]
      correspondence: false
institute:
  - MML: Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA
output: 
  officedown::rdocx_document:
    reference_docx: ../templates/jml_template.docx
    pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
bibliography: references.json
link-citations: true
csl: ../templates/royal-society-open-science.csl
always_allow_html: true
params:
   run_models: FALSE
---

## Abstract

Ice-associated seals rely on sea ice for a variety of activities, including breeding,
molting, pupping, and resting. In the Arctic, many of these activities occur in spring
(April-June) as sea ice begins to melt and retreat northward. Rapid acceleration of
climate change in Arctic ecosystems is therefore of concern as the quantity and quality of
suitable habitat is forecast to decrease. In this paper, we use data collected from
satellite-linked bio-loggers deployed between 2005 and 2020 to investigate the seasonal
timing and environmental factors affecting sea ice use by seals (specifically, bearded,
ribbon, and spotted seals) in the Bering and Chukchi seas. In addition to providing
baseline data on phenology, these data also allow us to quantify 'availability', which is
needed to accurately estimate abundance from aerial survey counts of seals basking on ice
(i.e., to correct for the proportion of animals that are in the water while surveys are
conducted). Using generalized linear mixed pseudo-models to properly account for temporal
autocorrelation, we fit models with a variety of covariates (e.g., day-of-year, solar
hour, age-sex class, wind speed, barometric pressure, temperature, precipitation) to
examine their ability to explain variation in hourly haul-out records. We found evidence
for strong diurnal and within-season patterns in haul-out behavior, as well as strong
weather effects (particularly wind and temperature). In general, seals were more likely to
rest on ice in the middle of the day and when wind speed was low and temperatures were
higher. Haul-out propensity increased through March and April, peaking in May and early
June before declining again. The timing and frequency of haul-out events also varied based
on species and age-sex class. For ribbon and spotted seals, models with year effects were
highly supported, indicating that the timing and magnitude of haul-out behavior varied
among years. However, we did not find evidence that haul-out timing was linked to annual
sea ice extent. Our analysis emphasizes the importance of accounting for variation in
haul-out behavior when interpreting the number of seals counted in aerial surveys.

keywords: availability; generalized linear mixed pseudo-model; haul-out behavior;
phenology; Phocidae; sea ice

```{r setup, include = FALSE}
library(conflicted)
library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
library(lubridate)
conflict_prefer("month","lubridate")
library(glue)
library(scales)
library(DBI)
library(dbplyr)
library(sf)
library(flextable)
library(patchwork)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(ggsci)
library(ggrepel)
library(gghighlight)
library(ggtext)
library(glmmLDTS)
library(mgcv)
library(ragg)

knitr::opts_chunk$set(dev = "ragg_png")

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  include = FALSE,
  fig.path = "../figures/",
  dpi = 400,
  fig.width = 6.5,
  out.width = "100%"
)

knitr::opts_chunk$set(
  tab.cap.style = "Table Caption",
  tab.cap.pre = "Table ",
  tab.cap.sep = ": "
  )

knitr::opts_chunk$set(
  fig.cap.style = "Image Caption",
  fig.cap.pre = "Figure ",
  fig.cap.sep = ": "
  )

options(scipen = 3, digits = 3)

theme_set(theme_minimal(base_family = 'Source Sans Pro'))
theme_replace(plot.title.position = "plot")
```

# Introduction

The global climate disruption is causing considerable reduction in Arctic sea ice extent,
volume, and seasonal presence [@kwok2018erl, @meier2014rg, @wang2017dsrpitsio]. These
changes have ripple effects on Arctic organisms, ecosystems, and the human communities who
live in the region [@huntington2020ncc]. Such disruptions are a particular cause of
concern for ice-associated seals (ringed (*Pusa hispida*), bearded (*Erignathus
barbatus*), spotted (*Phoca largha*), and ribbon seals (*Histriophoca fasciata*)) that
depend on spring and early summer sea ice (March-June) in the Bering and Chukchi seas as a
platform for important functions, such as pupping, nursing, breeding, and molting
[@boveng2009a; @boveng2013; @cameron2010a; @kelly2010]. Limited data and large knowledge
gaps complicate predictions about the ultimate effects of changes in sea ice on the
behavior, abundance, and distribution of these seals. Knowledge about evolutionary
constraints on the timing of reproductive and molting behavior is generally lacking, so it
is difficult to predict the readiness with which ice-associated seal species will be able
to adapt to future changes (e.g., by adjusting pupping or molting schedules to earlier
dates or different locales). Additionally, trends in abundance of these species are
largely unknown, so it is difficult to say what effect, if any, declines in sea ice
habitat have had, or will have, on seal densities.

For ice-associated seals, haul out is tightly linked with sea ice. Ribbon seals rest out
of the water almost exclusively on sea ice and are mostly pelagic outside the spring
pupping, breeding, and molting season [@burns1981]. While spotted and bearded seals rest
on beaches during ice-free months, they strongly prefer sea ice as a resting platform
during the spring and early summer [@burns2009]. The remoteness of the Bering and Chukchi
seas mean direct scientific observation of seal behavior is impractical. Bio-logging
devices are, thus, especially useful tools for collecting key information on movement and
haul-out behavior for these species.

Bio-logging records of time spent out of the water provide valuable data for identifying
covariates that explain variation in haul-out behavior. For instance, in the Antarctic,
Bengtson and Cameron [-@bengtson2004pb] relied on bio-logging data to demonstrate greater
haul-out propensity in juvenile crabeater seals (*Lobodon carcinophaga*) than adults, with
highest probabilities in February and at times close to solar noon. In the Arctic, Von
Duyke et al. [-@vonduyke2020ee] used seal-borne, satellite-linked loggers to corroborate
seasonal changes between diurnal and nocturnal haul-out behavior of ringed seals
previously described by Kelly et al. [-@kelly2010] using labor-intensive VHF radio tags
and direct observation. Bengtson et al. [-@bengtson2005pb] documented a higher propensity
for ringed seal basking near solar noon, as did Ver Hoef et al. [-@verhoef2014sm] in an
analysis of bearded, ribbon, and spotted seals using much larger sample sizes. Olnes et
al. [-@olnes2020pb] showed that the proportion of time bearded seals spent hauled out
progressively increased through spring and summer. And, Ver Hoef et al. [-@verhoef2014sm]
found haul-out probabilities increased gradually starting in March and peaking in May and
June for bearded, ribbon, and spotted seals.

While bio-logging of haul-out patterns clearly has utility for description of
ice-associated seals' natural history and ecology, some studies have focused, as we do
here, specifically on quantifying the probability of seals being hauled out as a means of
accounting for seals in the water, and therefore missed, during aerial surveys. These
studies typically use logistic regression-style analyses to estimate the time-specific
probability of being hauled out base on 'wet/dry' data relayed by bio-loggers. In these
models, haul-out probabilities were expressed as a function of predictive covariates, such
as time of day, day-of-year, sex, age class, and environmental conditions (e.g.,
[@reder2003pb], [@bengtson2004pb], [@bengtson2005pb], [@udevitz2009cjz], [@verhoef2014sm],
and [@southwell2008]). However, sample sizes have often been low and insufficient to
permit strong inference about life history and/or seasonal variation in haul-out
probabilities. For instance, Bengtson and Cameron's [-@bengtson2004pb] study included 5
adult and 2 juvenile crabeater seals, while Bengtson et al.'s [-@bengtson2005pb] study was
based on 6 telemetered ringed seals in the Chukchi Sea. These studies were often further
limited by logistical constraints on tagging fieldwork and the attachment duration or
operational life of satellite tags. In this study, we address some of these limitations by
deploying small bio-loggers designed for attachment on a portion of the study individuals.
These devices are designed to collect data through the molt period (when those adhered to
the hair would fall off) and, in some situations, provide multiple years of data.

Ultimately, knowledge of trends in phenology and abundance (or life history surrogates
such as survival and recruitment) will be necessary to make credible quantitative
predictions about the effects of the climate crisis on the abundance and distribution of
Arctic seal populations. Before we can construct a trend, however, we first require a
baseline. Several studies have contributed estimates of the distribution and abundance of
ice-associated seal species in the Arctic using aerial surveys (e.g., [@bengtson2005pb],
[@conn2014mee], and [@verhoef2014sm]). Such studies were conducted over very large areas
and estimation of absolute abundance required making inference about numerous nuisance
processes affecting the observation of seals on ice. This included availability (only
seals basking on ice were available to be counted), detection probability (observers or
automated detection systems may have missed some seals on ice), species misclassification,
and possible disturbance of seals by aircraft [@conn2014mee; @verhoef2014sm].

In this study, we used data from 15 years of bio-logger records in a comprehensive
investigation of haul-out behavior by 3 species of ice-associated seals in the Bering and
Chukchi seas. We focused our effort on bearded, spotted, and ribbon seals and did not
include ringed seals in the analysis. Ringed seals are unique in their use of snow lairs
in the spring months, and they warrant a separate study that includes additional
environmental covariates, such as the onset of snow melt and snow depth. Our goals are
threefold. First, we wish to establish baseline estimates for the chronology of haul-out
behavior in the critical spring season for each species across different age and sex
classes. Second, we wish to refine estimates of haul-out availability corrections for
aerial surveys in order to improve estimates of seal abundance. Previously estimated
availability correction factors (e.g., [@bengtson2005pb], [@conn2014mee], and
[@verhoef2014sm]) accounted for variables such as the time-of-day and day-of-year, but did
not investigate meteorological variables that have been shown to influence haul-out
behavior of walruses [@reder2003pb; @udevitz2009cjz]. Third, we explored the annual
variability in haul-out timing and possible linkage to dramatic changes in the
availability and extent of seasonal sea ice between 2005 and 2020. Our work extends the
scope of previous haul-out analyses, includes the influence of meteorological variability,
and investigates the potential impact of a changing icescape on the behavior of these
species.

# Methods

## Data collection

```{r getdata}
con <- dbConnect(
  odbc::odbc(),
  dsn = "PostgreSQL pep",
  uid = keyringr::get_kc_account("pgpep_londonj"),
  pwd = keyringr::decrypt_kc_pw("pgpep_londonj"),
  sslmode = "disable"
)

# qry removes two Sea of Okhotsk bearded seals
qry <- {
  "SELECT * 
  FROM telem.res_iceseal_haulout_cov 
  WHERE 
  EXTRACT(MONTH FROM haulout_dt) IN (3,4,5,6,7) AND
  rast_vwnd IS NOT NULL AND
  species != 'Ph' AND
  speno != 'EB2009_7010' AND
  speno != 'EB2009_3002' "
}

dat.sf <- sf::st_read(con,query = qry) %>% 
  filter(lubridate::yday(haulout_dt) < 197 ) %>% 
  mutate(age = case_when(
    age %in% c("pup", "PUP") ~ "YOUNG OF YEAR",
    age %in% c("yearling","YEARLING","subadult","Subadult") ~ "SUBADULT",
    age %in% c("adult","ADULT","Adult") ~ "ADULT",
    TRUE ~ age),
    species = case_when(
    species == "bearded" ~ "Bearded seal",
    species == "spotted" ~ "Spotted seal",
    TRUE ~ species),
    sex = case_when(
    sex %in% c("Female","female") ~ "F",
    sex %in% c("Male","male") ~ "M",
    TRUE ~ sex)
    ) %>% 
  filter(!is.na(age))

load(here::here('data/locs_sf.Rdata'))

deploy_meta <- locs_sf %>% 
  as_data_frame() %>% 
  select(speno, deploy_dt) %>% 
  filter(speno %in% !!unique(dat.sf$speno)) %>% 
  group_by(speno) %>% 
  summarise(deploy_dt = min(deploy_dt, na.rm = TRUE))

dat.sf <- dat.sf %>% left_join(deploy_meta) %>% 
  mutate(bday = as.Date(paste0(lubridate::year(deploy_dt),"-03-01")),
         age_days = difftime(haulout_dt, bday, tz="UTC", units = "days"),
         age = case_when(
           age_days > 360 & age == "YOUNG OF YEAR" ~ "SUBADULT",
           TRUE ~ age
         )) %>% 
  mutate(age = forcats::fct_relevel(age, c("ADULT", "SUBADULT", "YOUNG OF YEAR"))) %>% 
  dplyr::select(-c(bday,age_days)) %>% 
  relocate(deploy_dt, .after = age) %>% 
  filter(!(species == "Bearded seal" & age == "YOUNG OF YEAR"))

# read in world land polygons and buffer in by 100m; want to ensure only locations well
world <- ne_countries(scale = "large", returnclass = "sf") %>% 
  filter(continent != "Antarctica") %>% 
  st_wrap_dateline() %>% st_transform(st_crs(dat.sf)) %>% 
  st_buffer(-100)

not_intersects <- function(x,y) {
  !sf::st_intersects(x,y)
}

dat.sf <- dat.sf %>% 
  sf::st_filter(st_combine(world), .predicate = not_intersects)

save("dat.sf", file = here::here("data/dat_sf.Rdata"))

grid.sf <- sf::st_read(con,
                       query = "SELECT * FROM base.geo_analysis_grid") %>% 
  st_filter(st_bbox(dat.sf) %>% st_as_sfc())

dbDisconnect(con)

deploy_tbl <- dat.sf %>% 
  tibble::as_tibble() %>% 
  dplyr::group_by(speno,species, sex, age) %>% 
  dplyr::summarise(min_dt = min(haulout_dt),
            max_dt = max(haulout_dt),
            n_days = difftime(max_dt, min_dt, units = "days"),
            n_hours = n()
  ) %>% 
  dplyr::select(-c("min_dt","max_dt"))

```

Haul-out behavior data from `r dplyr::n_distinct(deploy_tbl$speno)` bio-loggers deployed
on bearded, spotted, and ribbon seals was filtered to include only records from 1 March to
15 July between `r min(lubridate::year(dat.sf$haulout_dt))` and
`r max(lubridate::year(dat.sf$haulout_dt))`. Bio-loggers were of the 'SPLASH' or 'SPOT'
family of tags developed by Wildlife Computers (Redmond, Washington, USA) and either
adhered to the hair on the seal or attached to the rear flipper inter-digital webbing. The
use of bio-loggers adhered to the back or head provide some benefits over flipper mounted
devices (e.g. increased satellite transmittal rates, locations at sea) but these are lost
during the annual molt, which limits the effective length of haul-out timelines.
Additionally, sensors attached to the head or dorsal region are often dry while the seal
is at the surface and can slightly bias the percent-dry values recorded by the device. For
this study, in cases where both tag types were deployed, percent dry observations from the
flipper tag were preferred. Sex and age class (non-dependent *young-of-the-year*, sexually
immature *subadults*, and mature *adults*) were determined at the time of deployment by
subjective judgement based on length, and pelage characteristics for some species. Field
identification of age class can be inexact, particularly when discerning subadults from
adults. In the case of ribbon seals, pelage differences between subadults and adults are
distinguishable, with subadults often having less defined ribbons. Bearded seal subadults
will often have a spotted pattern in the pelage that is not seen in adults. Spotted seal
pelage is monomorphic and can not be used to discern age class. Despite these challenges,
we feel the age classifications used in this analysis are useful categories in testing for
age-related effects on haul-out behavior. Seals determined to be less than one year were
classified as young-of-the-year. Subadults are those seals likely greater than one year of
age but not yet sexually mature. Adults are likely sexually mature and older than
approximately four years. For those tags deployed on young-of-the-year and transmitting
into the next year, the age class was advanced to subadult on 1 March of the following
year. Table \@ref(tab:tblDeploy) provides a summary of these deployments and data received
from them.

```{r tblDeploy, include=TRUE, tab.id = "tblDeploy", tab.cap = "Summary of bio-logger deployments across seal species and age classification from 1 March to 15 July. Total seal hours represents the sum of available data across all seals."}

deploy_tbl %>% 
  dplyr::group_by(species, sex, age) %>% 
  dplyr::summarise(
    n = n(),
    seal_hours = sum(n_hours)
  ) %>% 
  dplyr::mutate(
    deploy_stats = glue::glue('{n} ({format(seal_hours, big.mark=",")} seal hours)')
  ) %>% 
  dplyr::select(-c("n","seal_hours")) %>% 
  tidyr::pivot_wider(names_from = age, values_from = c(deploy_stats)) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(Species = species, Sex = sex, `Young-of-the-Year` = `YOUNG OF YEAR`,
                `Subadult` = SUBADULT, Adult = ADULT) %>% 
  flextable() %>% 
  add_header(`Young-of-the-Year` = "Age Class",
   Subadult = "Age Class", Adult = "Age Class",
   top = TRUE ) %>% 
  merge_h(part = "header") %>% 
  align(align = "center", part = "header") %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  autofit()
```

```{r seal-hours}
seal_hours_tbl <- dat.sf %>% 
  tibble::as_tibble() %>% 
  dplyr::mutate(wkofyr = lubridate::epiweek(haulout_dt),
                dayofwk = lubridate::wday(haulout_dt, label = TRUE, abbr = TRUE) %>% 
                  forcats::fct_relevel("Sun","Mon","Tue","Wed","Thu","Fri","Sat")
  ) %>% 
  dplyr::group_by(species, wkofyr, dayofwk) %>% 
  dplyr::count()
```

Tags that fall off due to molt or mortality and remain on ice or land can still send data
to satellites, and we did not want to include data from these tags in our analyses. As
such, end times of each deployment were identified by experts who examined several
simultaneous data streams (e.g., tag locations and dive behavior) to determine when tags
ceased providing data consistent with seal behavior. Sensor readings outside of these
start and end times were discarded prior to analysis. After approximately 9 months, 7
devices deployed on the rear flipper of bearded seals reported implausible sensor data.
All data after the first instance of unrealistic values were censored from this analysis.
Figure \@ref(fig:dataCal) shows a calendar distribution of these data across the study
season for each species. Observations are concentrated in the months of May and June due
to the timing of deployment (April and May) and the timing of molt (May and June). During
molt, seals (and their attached bio-loggers) spend more time out of the water and more
data is transmitted. Molt timing also impacts when many deployments end as any tags
adhered to the hair will fall off.

```{r seal-hours-calendar, include=TRUE, fig.height = 4.75, fig.id = "dataCal", fig.cap="Distribution of haul-out behavior data from 1 March to 15 July for each species. The higher density of data from May and June coincides with peak molting when seals (and their attached bio-loggers) are more likely hauled out. Additionally, a majority of deployments started in April and May. Data were collated across all years between 2005 and 2020."}
ggplot(data = seal_hours_tbl) + 
  geom_point(aes(x = wkofyr, y = dayofwk, color = n, size = n), stroke = 0, shape = 15) +
  scale_size(range = c(1,4.3), breaks = c(100,500,1000,1500)) +
  scale_color_gradient(low = "#6baed6", high = "#08306b") +
  scale_x_continuous(breaks = c(13,18,23,27),
                     labels = c("April", "May", "June", "July")) +
  scale_y_discrete(limits = rev) +
  guides(color = "none",
        size = guide_legend(override.aes = list(color = "#2171b5"),
                            title.position = "bottom", title.hjust = 0.5)
        ) +
  facet_grid(species ~ .) + 
  theme(
        #axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(x = "", y = "", color = "seal-hours",
       size = "seal hours of bio-logger data") +
  ggtitle("Haul-out behavior data is concentrated in the months of May and June",
          subtitle = "data presented in seal-hours collated across years (2005-2020)")
```

Haul-out behavior data was recorded and transmitted via the Argos satellite network as
hourly percent-dry timelines. For each hour of a day, the wet/dry sensor was polled by the
tag firmware every few seconds and a percent of the hour in a dry state was calculated
(Figure \@ref(fig:examplePlot)). Each data transmission message consisted of a complete
24-hour record. Tags were, generally, deployed and programmed in a manner to maximize data
transmission during the spring pupping and molting period, however, not all days during
the deployment were successfully transmitted. This is due to a variety of factors
including satellite coverage, tag availability (i.e. tags mounted to the rear flipper are
not available to the surface while the seal is at sea), tag performance, duty cycling, and
extra-terrestrial interference.

```{r examplePlot, include=TRUE, fig.id="examplePlot", fig.cap="Haul-out behavior observations recorded by a bio-logger deployment on a ribbon seal over two years during the months of March, April, and May. Empty areas of the plot within a month represent missing data that were not successfully transmitted from the tag."}
d <- dat.sf %>% 
  filter(speno == "HF2016_1002") %>% 
  dplyr::select(speno, haulout_dt, percent_dry) %>% 
  arrange(haulout_dt) %>% st_drop_geometry() %>% 
  mutate(year = lubridate::year(haulout_dt),
         month = lubridate::month(haulout_dt,label=TRUE),
         day = lubridate::day(haulout_dt),
         hour = lubridate::hour(haulout_dt)) %>% 
  filter(month != 'Jun')

p <- ggplot(d,aes(day,hour,fill=percent_dry))+
  geom_tile(color= "white",size=0) + 
  scale_fill_distiller(palette = "YlGnBu",
                       name = "hourly percent dry", aesthetics = "fill",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                           barheight = 0.5, title.hjust = 0.5)
    )

p <- p + facet_grid(year~month, scales = "free_x")
#p <- p + scale_y_continuous(trans = "reverse", expand = c(0,0))
p <- p + scale_x_continuous(breaks = c(5,15,25), expand = c(0,0)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  theme(panel.spacing = unit(0,"line")) +
  ggtitle("Observation records are incomplete because not all data are transmitted",
          subtitle = "percent-dry data for a single ribbon seal across two years") +
  xlab("day of month") + ylab("hour (UTC)")
p
```

Of key interest in this study is the relationship between haul-out behavior and a variety
of environmental covariates that vary with space and time. We explored the use of a
movement model to predict locations at specific times. However, the sparse nature of data
from some tags, especially those mounted to the rear flipper, resulted in poor modeling
performance or convergence issues. For this study, we calculated a weighted average daily
location where the inverse of the estimated Argos location error was used for the weight.
Any days where haul-out observations were present without any location data were filled
from the last calculated weighted average daily location, and any days where the location
intersected with land were removed from the dataset. Figure \@ref(fig:dataMap) shows the
spatial distribution of determined locations and available haul-out behavior data for each
species across the study area.

```{r locs-count-mapdata}
haulout_locs <- dat.sf %>% 
  filter(lubridate::hour(haulout_dt) == 12)

hexgrid <- sf::st_make_grid(st_bbox(haulout_locs) %>% st_as_sfc(), cellsize = 50*1000,
                            what = "polygons", square = FALSE) 
hexgrid <- st_sf(index = 1:length(lengths(hexgrid)), hexgrid)

hexbin <- st_join(haulout_locs, hexgrid, join = st_intersects)

locs_count <- hexgrid %>%
  left_join(
    count(hexbin, index, species) %>%
      as_tibble() %>%
      dplyr::select(species, index, ct=n)
  ) %>% drop_na()
```

```{r dataMap, fig.height = 3.75, include=TRUE, fig.id="dataMap", fig.cap="Spatial distribution of haul-out data during the months of March through July for each of the three species. Data were collated across all years between 2005 and 2020."}

world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(st_crs(locs_count))

ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(
    data = locs_count, size = 0.125,
    aes(fill = ct)
  ) +
  scale_fill_distiller(type = "seq", palette = "YlOrRd", direction = 1,
                    trans = "log10", aesthetics = "fill",
    name = "number of locations",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15, 
                           barheight = 0.5, title.hjust = 0.5)) + 
  facet_grid(~ species) +
  theme(legend.position = "bottom") +
  ggtitle("Spatial distribution of haul-out behavior data")
```

## Explanatory variables

In addition to sex and age class, we gathered additional variables that might help explain
variation in haul-out probabilities. These included day-of-year (for seasonal effects) and
local solar hour (for diurnal effects). Solar hour was calculated using the {solaR}
package [@perpinan2012jss] within the R statistical environment [@rcoreteam2019] using the
daily locations. We also compiled several meteorological variables shown to affect
haul-out behavior in other Arctic pinnipeds [@reder2003pb; @udevitz2009cjz; @perry2017].
In particular, we linked locations from satellite tags to key weather values from the
North American Regional Reanalysis (NARR) model produced by the National Centers for
Environmental Prediction [@mesinger2006bams]. The NARR model assimilates observational
data to produce a long-term picture of weather over North America. Numerous weather
variables are made available across the region 8 times daily. For this study, NARR weather
values were subset to the extent of our study area over the Bering and Chukchi Seas at
3-hr intervals based on the native grid resolution of 32 km (1024 sq. km). The following
meteorological variables were interpolated and assigned to seal locations using a bilinear
method: 1) air temperature at 2m above the earth's surface, 2) wind consisting of
northerly and easterly vector components converted to wind speed using the Euclidean norm,
3) barometric pressure at sea level, and 4) precipitation (Table
\@ref(tab:tblCovariates)).

Models for all species include the following effects: day-of-year, solar hour,
temperature, wind speed, barometric pressure, precipitation, wind chill (represented by a
*wind:temperature* interaction [@udevitz2009cjz]), and day-of-year and time of day
interactions to permit diurnal patterns to change throughout the year. Spotted and ribbon
seal models included age-sex class and interactions between day-of-year and age-sex class,
but we omitted these from bearded seal models due to poor representation of age-sex
classes (Table \@ref(tab:tblDeploy)). Bearded seal models included a latitudinal effect
(and an interaction with day-of-year), since bearded seals occupy a substantial range and
we were interested in possible differences in the timing of haul-out behavior along a
latitudinal gradient. We omitted the latitudinal effect from ribbon and spotted seal
models since, during the spring, these species are most prevalent near the southern ice
edge in the Bering Sea [@conn2014mee].

```{r wxCovMap, fig.height = 5.5, fig.id="wxCovMap", fig.cap="Example spatial distribution of weather covariates across the study region for 05 May 2012. The NARR was used to predict weather covariates for each hour in a percent dry timeline record."}

load(file = here::here('data/cov_means_daily_2012.Rdata'))
grid_plot <- cov_means_daily_2012 %>% filter(year == 2012,
                           yday == 125) %>%
  inner_join(grid.sf, by = 'cell') %>%
  sf::st_as_sf()

world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(st_crs(grid_plot))

p_precip <- ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(grid_plot, aes(fill = precip), size = 0) +
  scale_fill_distiller(palette = "RdYlBu") +
  # scale_fill_viridis_c(option = "C") +
  labs(subtitle = "precipitation") +
  theme(legend.position = "none")

p_pressure <- ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(grid_plot, aes(fill = pressure), size = 0) +
  scale_fill_distiller(palette = "RdYlBu") +
  # scale_fill_viridis_c(option = "C") +
  labs(subtitle = "barometric pressure") +
  theme(legend.position = "none")

p_wind <- ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(grid_plot, aes(fill = wind), size = 0) +
  scale_fill_distiller(palette = "RdYlBu") +
  # scale_fill_viridis_c(option = "C") +
  labs(subtitle = "wind") +
  theme(legend.position = "none")

p_temp2 <- ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(grid_plot, aes(fill = temp2), size = 0) +
  scale_fill_distiller(palette = "RdYlBu") +
  # scale_fill_viridis_c(option = "C") +
  labs(subtitle = "temperature") +
  theme(legend.position = "none")

(p_precip + p_pressure) / (p_wind + p_temp2) +
  plot_annotation(
    title = 'Weather varies across the Bering and Chukchi seas',
    caption = "North American Regional Reanalysis (NARR), 2012 May 05")
```

Notably missing from this list of explanatory variables is any spatial-temporal
representation of sea ice concentration, area, or extent. This may seem counterintuitive
when modeling the haul-out behavior of seal species with such a close association to sea
ice; seals haul out in the presence of sea ice, and we could assess the local
concentration of sea ice during these events (see [@olnes2020]). However, a major focus of
this study is to develop models applicable for aerial survey correction factors and using
sea ice as a covariate would almost certainly bias haul-out predictions towards those
seals that are on or near ice and therefore more likely to haul out. Since aerial surveys
can only detect seals on ice, abundance estimates would be missing a correction for those
seals that are away from ice (e.g., on foraging trips). Additionally, many of the
bio-logger deployments used in this study were single attachments to the rear flipper of
the seal and do not provide at-sea locations, limiting our ability to fully evaluate
habitat preferences related to sea-ice. It should be noted that the aerial survey counts
are modeled with sea-ice and other habitat parameters. So, any final abundance estimates
that combine models of survey counts and our model for proportion of seals hauled out on
ice do account for the strong association between these species and sea-ice in the Bering
and Chukchi seas. Lastly, our study was limited to the spring season when seal haul-out
tendencies are strongly influenced by pupping, nursing, breeding, and molt and these
drivers are likely more influential than specific sea-ice concentration. Crawford et al
[-@crawford2019] compared haul-out probability models for bearded seals and found those
that only included season (and not sea-ice concentration) were the most parsimonious. For
these reasons, we have elected not to use sea ice concentration as a predictor for
haul-out probability; studies to evaluate habitat selection in relation to sea-ice
parameters and other environmental covariates are in progress.

We explored whether the annual variation in maximum spring sea ice extent in the Bering
Sea influences the seasonal peak of seal haul-out behavior. In particular, we used sea ice
concentration data from the Nimbus-7 SMMR and DMSP SSM/I-SSMIS Passive Microwave Dataset,
Version 1 [@cavalieri1996] to calculate maximum sea ice extent. All sea ice concentration
grid cells (25 km^2^) in the study area with greater than 15% concentration were summed to
get the total sea ice extent for each day between 15 February and 15 July across all
years. Maximum spring sea ice extent was simply the largest daily value for each year.

```{r tblCovariates, include=TRUE, tab.id = "tblCovariates", tab.cap = "Explanatory covariates used in analyses of binary haul-out records for ice-associated seals.  Note that we also considered select interactions (see article text) between these primary covariates.  For instance, wind chill was represented by the interaction _temperature:wind_."}
covariate_tbl <- tibble::tribble(
  ~Covariate, ~Type, ~Source, ~Description,
  "Age-sex class",   "Categorical", "Field Assessment", 
  "young-of-the-year, subadult, adult male and adult female",
  "Hour", "Continuous; Fourier basis", "Bio-logger", 
  "local solar hour using 6 variables of a Fourier-series basis",
  "Day",   "Continuous", "Bio-logger", 
  "linear, quadratic, and cubic effects of day-of-year",
  "Precip", "Continuous", "NARR", 
  "convective precipitation (kg/m2)",
  "Pressure", "Continuous", "NARR", 
  "atmospheric pressure at sea level (kPa)",
  "Temp", "Continuous", "NARR",
  "air temperatures at 2m above the earth’s surface",
  "Wind", "Continuous", "NARR", 
  "northerly and easterly vector components for wind converted into a single wind speed via the Euclidean norm",
  "Northing", "Continuous", "Bio-logger","latitude divided by the mean latitude across all locations (for bearded seals only)"
)

covariate_tbl %>% 
  flextable() %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  set_table_properties(width = .95, layout = "autofit")
```

## Haul-out modeling

```{r runmodels}
if (params$run_models) {
source(here::here('analysis/scripts/model.R'))
} else {
  load(here::here('data/fit_ribbon.Rdata'))
  load(here::here('data/fit_bearded.Rdata'))
  load(here::here('data/fit_spotted.Rdata'))
}
```

Haul-out records for seals are often characterized by multiple hours spent basking on ice
alternating with long periods of foraging (Figure \@ref(fig:examplePlot)). Commonly used
statistical models for binary data (e.g. logistic regression) assume independence among
responses, an assumption that is clearly violated if hourly responses are modeled. Any
analysis that ignores temporal dependence in responses will thus have overstated precision
[@betts2006].

To properly account for temporal dependence within a computationally tractable framework,
we used generalized linear mixed pseudo-models (GLMPMs; [@verhoef2010cs]) to model
variation in haul-out behavior as a function of (1) covariate predictors, (2) temporally
autocorrelated random effects, and (3) individual random effects representing
heterogeneity in individual behavior. We used the {glmmLDTS} package [@verhoef2010cs] to
implement GLMPMs in the R programming environment [@rcoreteam2021]. We explored two
different model formulations for our data, and owing to the large number of records, we
fit separate models to bearded, ribbon, and spotted seal data sets. In our first model
formulation, for each species, we fitted a year-independent model that predicted average
haul-out behavior as a function of demographic, environmental, seasonal, and diurnal
effects. Second, for ribbon and spotted seals (which had considerably more data than
bearded seals), we fitted models that included all the effects from the first model, but
also permitted annual variation in haul-out timing. This second class of models was used
to examine whether haul-out patterns varied by year and to determine the annual timing of
apparent peaks in haul-out behavior.

We followed Ver Hoef et al. [-@verhoef2014sm] in using linear, quadratic, and cubic
effects of day-of-year to represent seasonal changes in behavior. However, unlike previous
models for harbor seals [@london2012] and ice-associated seals [@verhoef2014sm], which
treated hour-of-day as a 24-level categorical variable to capture diurnal cycles, we
adopted a continuous formulation based on Fourier series that provides a flexible model
while preserving the inherent circularity needed for time-of-day effects (i.e., hour 0
should be equal to hour 24). It also represents hour-of-day with 6 parameters, which is a
considerable reduction when compared to a 24 parameter variable, especially when fitting
models with interactions between hour-of-day and other variables (e.g., age-sex class,
day-of-year). According to this approach, we used the following specification for
hour-of-day effects:

$$
H_t=\alpha_1 \cos(\frac{\pi t}{4}) + \alpha_2 \sin(\frac{\pi t}{4}) + \alpha_3 \cos(\frac{\pi t}{6}) + \alpha_4 \sin(\frac{\pi t}{6}) + \alpha_5 \cos(\frac{\pi t}{12}) + \alpha_6 \sin(\frac{\pi t}{12})
$$ 
where $H_t$ gives the effect for solar hour $t$ and $α_{i}$ are estimated parameters
(regression coefficients).

For the second set of models examining inter-annual variation in sea ice use, we fitted
models with year by day-of-year interactions. However, in this case we only included
*year:day* and *year:day2*, omitting the main effects of year as well as *year:day3*
interactions because models with the latter effects were numerically unstable. However,
the modeled interactions were sufficient to allow shifts in haul-out distribution, as one
can show mathematically that a simple horizontal shift in timing of haul-out distributions
does not affect the main effects or cubic terms in a polynomial regression model.

```{r runyearmodels}
if (params$run_models) {
 source(here::here('analysis/scripts/model_year_hf.R'))
 source(here::here('analysis/scripts/model_year_pl.R'))

} else {
 load(here::here('data/fit_ribbon_year.Rdata'))
 load(here::here('data/fit_spotted_year.Rdata'))
}
```

A typical model fitting exercise would also include a model selection process. However,
AIC (and similar criteria) is not suitable when using pseudo-likelihoods, because
pseudo-data generated in the model fitting process [@verhoef2010cs] differ between models
[@teneyck2017]. After fitting GLMPM models, we instead used "type III" $F$-tests to
calculate $p$-values [@verhoef2010cs] to evaluate model performance and important terms.
We also produced predictions of haul-out behavior as a function of influential predictors
(e.g. solar hour, day-of-year, age-sex). Weather covariates for these predictions were
based on daily or hourly smoothed weather covariate values across the study region. Such
predictions were then used to develop haul-out probability surfaces, explore conditional
effects of weather covariates, and determine annual peaks in haul-out activity. The timing
of peak haul-out behavior was further used to regress against the annual maximum sea ice
extent in the study region.

# Results

```{r createnewdata}
# function to create a prediction data frame for a given age_sex group
create_newdata <- function(data, age_sex) {
  df_list <- vector(mode = "list", length = length({{age_sex}} ))

  for (a_s in {{age_sex}}) {
  range_yday <- data %>% filter(age_sex == a_s ) %>%
    summarize(start_day = min(yday),
              end_day = max(yday))
  start_day <- range_yday$start_day
  end_day <- range_yday$end_day
  n_days = (end_day - start_day) + 1

  # for wx covariates we'll use a gam to get values by day/hour since wx is likely
  # to vary w/in day over the season

  gam.baro <-
    gam(pressure ~ s(yday), data = data)
  gam.temp <-
    gam(temp2 ~ s(yday) + s(as.numeric(solar_hour)), data = data)
  gam.wind <-
    gam(wind ~ s(yday) + s(as.numeric(solar_hour)), data = data)
  gam.precip <-
    gam(precip ~ s(yday), data = data)

  wx_new_data <- data.frame(
    solar_hour = rep(0:23, each = n_days),
    yday = rep(start_day:end_day, times = 24)
  )

  temp_pred <- predict(gam.temp, newdata = wx_new_data)
  wind_pred <- predict(gam.wind, newdata = wx_new_data)
  baro_pred <- predict(gam.baro, newdata = wx_new_data)
  precip_pred <- predict(gam.precip, newdata = wx_new_data)

  df <- data.frame(
    age_sex = a_s,
    solar_hour = rep(0:23, each = n_days),
    yday = rep(start_day:end_day, times = 24),
    northing = mean(data$northing),
    temp2 = temp_pred,
    wind = wind_pred,
    pressure = baro_pred,
    precip = precip_pred
  ) %>%
    mutate(
      sin1 = sin(pi * solar_hour / 12),
      cos1 = cos(pi * solar_hour / 12),
      sin2 = sin(pi * solar_hour / 6),
      cos2 = cos(pi * solar_hour / 6),
      sin3 = sin(pi * solar_hour / 4),
      cos3 = sin(pi * solar_hour / 4),
    ) %>%
    mutate(day = (yday - 120) / 10,
           day2 = day ^ 2,
           day3 = day ^ 3)
  df_list[[a_s]] <- df
  }
  if(length({{age_sex}}) > 1) {
  df_out <- bind_rows(df_list) %>%
    mutate(age_sex = forcats::fct_relevel(
      age_sex,c("ADULT.F","ADULT.M","SUBADULT","YOUNG OF YEAR"))
    )
  } else {
    df_out <- bind_rows(df_list)
  }
}
```

Models omitting year effects suggested that day-of-year, solar hour, age-sex class,
temperature, and wind substantially influenced haul-out behavior of all three species,
with $F$ tests producing $p$-values less than 0.05 for variables embodying these effects
and/or their interactions. Haul-out probabilities typically increased throughout March and
April, reaching a peak in May and early June before declining again. Diurnal patterns were
present, with maximum haul-out behavior centered around local solar noon.

## Ribbon Seals

```{r hfnewdata}
# get the ribbon seal fit object
load(here::here("data/fit_ribbon.Rdata"))
ribbon_newdata <- create_newdata(fit_ribbon$dataset,
                                 age_sex = levels(fit_ribbon$dataset$age_sex))

# create the model matrix
ribbon_mm <- model.matrix(fit_ribbon$fixed.formula[-2],
                data = ribbon_newdata)

#clean up extra intercept and get coef
fit_coef <- fit_ribbon$fixed.effects %>%
  filter(!is.na(std.err)) %>%
  pull(estimate)

ribbon_newdata <- ribbon_newdata %>%
  mutate(logit_fits = as.vector(ribbon_mm %*% fit_coef),
         logit_fits_var = diag(ribbon_mm %*% fit_ribbon$covb %*% t(ribbon_mm)),
         logit_fits_lo95 = logit_fits - 1.96*sqrt(logit_fits_var),
         logit_fits_up95 = logit_fits + 1.96*sqrt(logit_fits_var),
         ho_prob = plogis(ribbon_mm %*% fit_coef),
         lower95 = plogis(logit_fits_lo95),
         upper95 = plogis(logit_fits_up95)
         )
```

Ribbon seals show a pattern of gradually increasing haul-out probability in April that
peaks in late May for subadults and in early June for adults (Figure
\@ref(fig:ribbonHOCal)). The behavior is clearly centered around local solar noon and
expands to other hours later in the season as seals enter their molting period. Subadults
showed an earlier start and more intense haul-out activity in April and May. The
young-of-the-year records begin after weaning and the model predictions demonstrate the
ontogeny of in-water activities (e.g. diving, foraging) in May. Adult females have a more
protracted haul-out season compared to males, and more time spent resting in June and
July. The model suggests adult male ribbon seals complete their molt by the end of June.
Lastly, for adults of both sexes there is some indication for a shift to a crepuscular
haul-out pattern in late June and July.

```{r ribbonHOCal, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOCal", fig.cap="Predicted haul-out probability of ribbon seals from 1 March to 15 July for each age and sex class used in the model. Adult females are indicated by 'ADULT.F' and 'ADULT.M' indicates adult males. 'SUBADULT' and 'YOUNG OF YEAR' age classes include both sexes."}
plot_df <- ribbon_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label = TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df, aes(day,solar_hour,fill = ho_prob)) +
  geom_tile(color = "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour = "white")) +
  theme(axis.ticks = element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Ribbon seals rest on sea ice in late spring and around solar noon.")
p
```

```{r type3}
ribbon_type3 <- tibble(
  fit_ribbon$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>% 
  column_to_rownames(var = "Effect") 
ribbon_type3 <- setNames(split(ribbon_type3, 
                               seq(nrow(ribbon_type3))), 
                         rownames(ribbon_type3))

spotted_type3 <- tibble(
  fit_spotted$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
spotted_type3 <- setNames(split(spotted_type3, 
                               seq(nrow(spotted_type3))), 
                         rownames(spotted_type3))

bearded_type3 <- tibble(
  fit_bearded$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
bearded_type3 <- setNames(split(bearded_type3, 
                               seq(nrow(bearded_type3))), 
                         rownames(bearded_type3))
```

The haul-out probability for ribbon seals was strongly influenced by each of the weather
covariates included within the model except precipitation. The impacts of temperature
($F_{`r paste(ribbon_type3$temp2$Num.df, ribbon_type3$temp2$Den.df, sep=",")`}$ =
`r ribbon_type3$temp2$F.value`; $p$ = `r ribbon_type3$temp2$Prob.F`), wind
($F_{`r paste(ribbon_type3$wind$Num.df, ribbon_type3$wind$Den.df, sep=",")`}$ =
`r ribbon_type3$wind$F.value`; $p$ = `r ribbon_type3$wind$Prob.F`), and barometric
pressure
($F_{`r paste(ribbon_type3$pressure$Num.df, ribbon_type3$pressure$Den.df, sep=",")`}$ =
`r ribbon_type3$pressure$F.value`; $p$ = `r ribbon_type3$pressure$Prob.F`) were especially
noticeable as ribbon seals were less likely to haul out at higher winds and lower pressure
values and more likely to be on the ice when temperatures were relatively warm. Wind chill
($F_{`r paste(ribbon_type3$temp2_wind$Num.df, ribbon_type3$temp2_wind$Den.df, sep=",")`}$
= `r ribbon_type3$temp2_wind$F.value`; $p$ = `r ribbon_type3$temp2_wind$Prob.F`) showed a
moderately negative influence on haul-out probability.

Visualizing the marginal effect of each weather covariate on haul-out probability is difficult
in this analysis because of the collinearity between covariates as well as the temporal variation 
in covariate values during the study period. The relationship of each weather covariate with haul-out
probability, averaged over the other weather conditions, is more variable than simply
showing model coefficients would imply. Figure \@ref(fig:ribbonHOwx)) presents the
predicted haul-out probability of ribbon seals across the range of weather conditions
encountered in the observed data. The collinearity and temporal variation are apparent
in the visual patterns that emerge. By

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Variability in predicted haul-out probability of ribbon seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  geom_smooth(se = FALSE) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Ribbon seal haul-out behavior is influenced by weather')
```

## Spotted Seals

```{r plnewdata}
# get the spotted seal fit object
load(here::here("data/fit_spotted.Rdata"))

spotted_newdata <- create_newdata(fit_spotted$dataset,
                                 age_sex = levels(fit_spotted$dataset$age_sex))

# create the model matrix
spotted_mm <- model.matrix(fit_spotted$fixed.formula[-2],
                          data = spotted_newdata)

#clean up extra intercept and get coef
fit_coef <- fit_spotted$fixed.effects %>%
  filter(!is.na(std.err)) %>%
  pull(estimate)

spotted_newdata <- spotted_newdata %>%
  mutate(logit_fits = as.vector(spotted_mm %*% fit_coef),
         logit_fits_var = diag(spotted_mm %*% fit_spotted$covb %*% t(spotted_mm)),
         logit_fits_lo95 = logit_fits - 1.96*sqrt(logit_fits_var),
         logit_fits_up95 = logit_fits + 1.96*sqrt(logit_fits_var),
         ho_prob = plogis(spotted_mm %*% fit_coef),
         lower95 = plogis(logit_fits_lo95),
         upper95 = plogis(logit_fits_up95)
  )
```

Compared to ribbon seals, spotted seals showed a longer spring haul-out season that is
less intensely centered on solar noon (Figure \@ref(fig:spottedHOCal)). Adults of both sexes spend considerable time in
April and May resting on the ice. As with ribbon seals, the young-of-the-year records
begin after weaning and the model predictions demonstrate the ontogeny of in-water
activities (e.g. diving, foraging) in May.

```{r spottedHOCal, fig.height = 5.5, include=TRUE, fig.id = "spottedHOCal", fig.cap="Predicted haul-out probability of spotted seals from 1 March to 15 July for each age and sex class used in the model. Adult females are indicated by 'ADULT.F' and 'ADULT.M' indicates adult males. 'SUBADULT' and 'YOUNG OF YEAR' age classes include both sexes."}
plot_df <- spotted_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df,aes(day,solar_hour,fill=ho_prob))+
  geom_tile(color= "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Spotted seals rest on sea ice in late spring and around solar noon.")
p
```

Spotted seal haul-out behavior appears most strongly influenced by temperature
($F_{`r paste(spotted_type3$temp2$Num.df, spotted_type3$temp2$Den.df, sep=",")`}$ =
`r spotted_type3$temp2$F.value`; $p$ = `r spotted_type3$temp2$Prob.F`) and wind
($F_{`r paste(spotted_type3$wind$Num.df, spotted_type3$wind$Den.df, sep=",")`}$ =
`r spotted_type3$wind$F.value`; $p$ = `r spotted_type3$wind$Prob.F`) with barometric
pressure having a moderate influence
($F_{`r paste(spotted_type3$pressure$Num.df, spotted_type3$pressure$Den.df, sep=",")`}$ =
`r spotted_type3$pressure$F.value`; $p$ = `r spotted_type3$pressure$Prob.F`). Spotted
seals were less likely to haul out at higher winds and more likely to be on the ice when
temperatures were relatively warm. Wind chill (*temperature:wind*) and precipitation were
not as influential as the other covariates. Differences in the magnitude of response
between the age-sex classes are present and consistent across each of the weather
covariates (Figure \@ref(fig:spottedHOwx)).

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Variability in the predicted haul-out probability of spotted seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}
library(ggsci)

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Spotted seal haul-out behavior is influenced by weather')
```

## Bearded Seals

Age and sex class were not included in the model for bearded seals and, as such, results
are shown for all ages. Unlike ribbon and spotted seals, bearded seals were predicted to
have a bi-modal distribution of haul-out probability across the day (Figure \@ref(fig:beardedHOCal)). In addition to a peak
around local solar noon, the bearded seal model predicts additional haul-out activity
around local midnight. The haul-out behavior is also more protracted throughout the spring
season compared to ribbon and spotted seals.

```{r ebnewdata}
# get the bearded seal fit object
load(here::here("data/fit_bearded.Rdata"))

eb_data <- fit_bearded$dataset %>% 
  mutate(age_sex = forcats::as_factor("ALL AGES"))

bearded_newdata <- create_newdata(eb_data,
                                  age_sex = levels(eb_data$age_sex))

# create the model matrix
bearded_mm <- model.matrix(fit_bearded$fixed.formula[-2],
                           data = bearded_newdata)

#clean up extra intercept and get coef
fit_coef <- fit_bearded$fixed.effects %>%
  filter(!is.na(std.err)) %>%
  pull(estimate)

bearded_newdata <- bearded_newdata %>%
  mutate(logit_fits = as.vector(bearded_mm %*% fit_coef),
         logit_fits_var = diag(bearded_mm %*% fit_bearded$covb %*% t(bearded_mm)),
         logit_fits_lo95 = logit_fits - 1.96*sqrt(logit_fits_var),
         logit_fits_up95 = logit_fits + 1.96*sqrt(logit_fits_var),
         ho_prob = plogis(bearded_mm %*% fit_coef),
         lower95 = plogis(logit_fits_lo95),
         upper95 = plogis(logit_fits_up95)
  )
```

```{r beardedHOCal, fig.height = 3.5, include=TRUE, fig.id = "beardedHOCal", fig.cap="Predicted haul-out probability of bearded seals from 1 March to 15 July for all and sex classes."}
plot_df <- bearded_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df,aes(day,solar_hour,fill = ho_prob))+
  geom_tile(color= "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Bearded seals rest on sea ice around solar noon and solar midnight")
p
```

Like ribbon and spotted seals, bearded seal haul-out probability appears most influenced
by wind ($F_{`r paste(bearded_type3$wind$Num.df, bearded_type3$wind$Den.df, sep=",")`}$ =
`r bearded_type3$wind$F.value`; $p$ = `r bearded_type3$wind$Prob.F`) and temperature
($F_{`r paste(bearded_type3$temp2$Num.df, bearded_type3$temp2$Den.df, sep=",")`}$ =
`r bearded_type3$temp2$F.value`; $p$ = `r bearded_type3$temp2$Prob.F`). Additionally, wind
chill was also a significant influence
($F_{`r paste(bearded_type3$temp2_wind$Num.df, bearded_type3$temp2_wind$Den.df, sep=",")`}$
= `r bearded_type3$temp2_wind$F.value`; $p$ = `r bearded_type3$temp2_wind$Prob.F`). Any
influence of barometric pressure or precipitation is less apparent (Figure \@ref(fig:beardedHOwx)).

```{r beardedHOwx,fig.height = 5.5, include=TRUE, fig.id = "beardedHOwx", fig.cap="Variability in predicted haul-out probability of bearded seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df, aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df, aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df, aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

((p_temp + p_wind) / (p_pressure + p_precip)) +
  plot_annotation(
    title = 'Bearded seal haul-out behavior is less influeced by weather')
```

## Annual variation in haul-out

```{r createnewdatayr}
create_newdata_year <- function(data, age_sex) {
  df_list <- vector(mode = "list", length = length({{age_sex}} ))
  df_list_years <- vector(mode = "list", length = length(levels(data$year)))

  for (a_s in {{age_sex}}) {
  for (y in levels(data$year)) {
    d <- data %>% filter(year == y, age_sex == a_s)
    if(nrow(d) == 0) { next }
    range_yday <- d %>%
      summarize(start_day = min(yday),
                end_day = max(yday))
    if(range_yday$start_day > 152) {next}
    if(range_yday$end_day < 121) {next}
    print(range_yday)
    start_day <- range_yday$start_day
    end_day <- range_yday$end_day
    n_days = (end_day - start_day) + 1

    # for wx covariates we'll use a gam to get values by day/hour since wx is likely
    # to vary w/in day over the season

    gam.baro <-
      gam(pressure ~ s(yday), data = data)
    gam.temp <-
      gam(temp2 ~ s(yday) + s(as.numeric(solar_hour)), data = data)
    gam.wind <-
      gam(wind ~ s(yday) + s(as.numeric(solar_hour)), data = data)
    gam.precip <-
      gam(precip ~ s(yday), data = data)

    wx_new_data <- data.frame(
      solar_hour = 12,
      yday = start_day:end_day
    )

    temp_pred <- predict(gam.temp, newdata = wx_new_data)
    wind_pred <- predict(gam.wind, newdata = wx_new_data)
    baro_pred <- predict(gam.baro, newdata = wx_new_data)
    precip_pred <- predict(gam.precip, newdata = wx_new_data)

    df <- data.frame(
      age_sex = a_s,
      year = y,
      solar_hour = 12,
      yday = start_day:end_day,
      northing = mean(data$northing),
      temp2 = temp_pred,
      wind = wind_pred,
      pressure = baro_pred,
      precip = precip_pred
    ) %>%
      mutate(
        sin1 = sin(pi * solar_hour / 12),
        cos1 = cos(pi * solar_hour / 12),
        sin2 = sin(pi * solar_hour / 6),
        cos2 = cos(pi * solar_hour / 6),
        sin3 = sin(pi * solar_hour / 4),
        cos3 = sin(pi * solar_hour / 4),
      ) %>%
      mutate(day = (yday - 120) / 10,
             day2 = day ^ 2,
             day3 = day ^ 3)
    df_list_years[[y]] <- df
  }
    df_list[[a_s]] <- bind_rows(df_list_years)
  }
  df_out <- bind_rows(df_list) %>%
    mutate(age_sex = forcats::fct_relevel(
      age_sex,c("ADULT.F","ADULT.M","SUBADULT","YOUNG OF YEAR")),
      year = forcats::as_factor(as.numeric(year))
    )

  return(df_out)
}
```

```{r hfpvyear}
load(here::here('data/fit_ribbon_year.Rdata'))

ribbon_newdata_year <- create_newdata_year(fit_ribbon_year$dataset,
                                           age_sex = levels(fit_ribbon_year$dataset$age_sex))

# create the model matrix
ribbon_mm <- model.matrix(fit_ribbon_year$fixed.formula[-2],
                          data = ribbon_newdata_year)

#clean up extra intercept and get coef
fit_coef <- fit_ribbon_year$fixed.effects %>%
  filter(!is.na(std.err)) %>%
  pull(estimate)

# multiply by the model coefficients and convert to probabilities with plogis
# (this came from J Laake code for Hood Canal estimates)
ribbon_newdata_year <- ribbon_newdata_year %>%
  mutate(logit_fits = as.vector(ribbon_mm %*% fit_coef),
         logit_fits_var = diag(ribbon_mm %*% fit_ribbon_year$covb %*% t(ribbon_mm)),
         logit_fits_lo95 = logit_fits - 1.96*sqrt(logit_fits_var),
         logit_fits_up95 = logit_fits + 1.96*sqrt(logit_fits_var),
         ho_prob = plogis(ribbon_mm %*% fit_coef),
         lower95 = plogis(logit_fits_lo95),
         upper95 = plogis(logit_fits_up95)
  )

load(here::here('data/fit_spotted_year.Rdata'))

spotted_newdata_year <- create_newdata_year(fit_spotted_year$dataset,
                                            age_sex = levels(fit_spotted_year$dataset$age_sex))

# create the model matrix
spotted_mm <- model.matrix(fit_spotted_year$fixed.formula[-2],
                           data = spotted_newdata_year)

#clean up extra intercept and get coef
fit_coef <- fit_spotted_year$fixed.effects %>%
  filter(!is.na(std.err)) %>%
  pull(estimate)

# multiply by the model coefficients and convert to probabilities with plogis
# (this came from J Laake code for Hood Canal estimates)
spotted_newdata_year <- spotted_newdata_year %>%
  mutate(logit_fits = as.vector(spotted_mm %*% fit_coef),
         logit_fits_var = diag(spotted_mm %*% fit_spotted_year$covb %*% t(spotted_mm)),
         logit_fits_lo95 = logit_fits - 1.96*sqrt(logit_fits_var),
         logit_fits_up95 = logit_fits + 1.96*sqrt(logit_fits_var),
         ho_prob = plogis(spotted_mm %*% fit_coef),
         lower95 = plogis(logit_fits_lo95),
         upper95 = plogis(logit_fits_up95)
  )
```

```{r type3_year}
ribbon_year_type3 <- tibble(
  fit_ribbon_year$typeIII.hypoth
) %>% 
  mutate(# Prob.F = ifelse(Prob.F<0.001, "<0.01",Prob.F),
         Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>% 
  column_to_rownames(var = "Effect") 
ribbon_year_type3 <- setNames(split(ribbon_year_type3, 
                               seq(nrow(ribbon_year_type3))), 
                         rownames(ribbon_year_type3))

spotted_year_type3 <- tibble(
  fit_spotted_year$typeIII.hypoth
) %>% 
  mutate(# Prob.F = ifelse(Prob.F<0.001, "<0.01",Prob.F),
         Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
spotted_year_type3 <- setNames(split(spotted_year_type3, 
                               seq(nrow(spotted_year_type3))), 
                         rownames(spotted_year_type3))
```

The second set of models that included annual variation in haul-out patterns uncovered
significant contributions for linear and quadratic interactions between day and year in
both ribbon seals (day $\times$ year,
$F_{`r paste(ribbon_year_type3$day_year$Num.df, ribbon_year_type3$day_year$Den.df, sep=",")`}$
= `r ribbon_year_type3$day_year$F.value`; $p$ = `r ribbon_year_type3$day_year$Prob.F`;
day^2^:year,
$F_{`r paste(ribbon_year_type3$day2_year$Num.df, ribbon_year_type3$day2_year$Den.df, sep=",")`}$
= `r ribbon_year_type3$day2_year$F.value`; $p$ = `r ribbon_year_type3$day2_year$Prob.F`)
and spotted seals (day $\times$ year,
$F_{`r paste(spotted_year_type3$day_year$Num.df, spotted_year_type3$day_year$Den.df, sep=",")`}$
= `r spotted_year_type3$day_year$F.value`; $p$ = `r spotted_year_type3$day_year$Prob.F`;
day^2^:year,
$F_{`r paste(spotted_year_type3$day2_year$Num.df, spotted_year_type3$day2_year$Den.df, sep=",")`}$
= `r spotted_year_type3$day2_year$F.value`; $p$ =
`r spotted_year_type3$day2_year$Prob.F`). Predicted distributions of haul-out activity
were largely unimodal, but varied some among and within years with respect to the timing
and magnitude of haul-out peaks (Figure \@ref(fig:annualHO)). It is important to note that
predicted variation in annual haul-out patterns likely reflects both process error and
sampling variability. While we did remove any years where only one deployment in a species
+ age-sex group was present, there are still some years where the pattern shown is
informed by a small number of individuals that may not represent population level
processes.

```{r annualHO, fig.height = 5.5, include=TRUE, fig.id = "annualHO", fig.cap="Annual variability in the timing of peak haul-out probability for ribbon and spotted seals across 10 years. Predictions are shown for local solar noon and under smoothed weather conditions. Only those groups (age-sex + year) that included observations from more than one seal are shown. Additionally, any groups where the underlying data started after 1 June or ended before 1 May are not included."}
select_groups <- fit_ribbon_year$dataset %>%
  filter(age_sex != "YOUNG OF YEAR") %>%
  group_by(age_sex, year) %>%
  summarise(n = n_distinct(speno), group_id = paste0(age_sex,year)) %>%
  filter(n >1) %>% pull(group_id)

plot_df <- ribbon_newdata_year %>%
  mutate(species = "Ribbon seal",
         date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug", paste0(age_sex,year) %in% select_groups)

peak_ho_df <- plot_df %>% filter(age_sex != "YOUNG OF YEAR",
                                 yday > 100) %>%
  group_by(species, age_sex, year) %>% slice_max(ho_prob)

select_groups <- fit_spotted_year$dataset %>%
  filter(age_sex != "YOUNG OF YEAR") %>%
  group_by(age_sex, year) %>%
  summarise(n = n_distinct(speno), group_id = paste0(age_sex,year)) %>%
  filter(n >1) %>% pull(group_id)

plot_df2 <- spotted_newdata_year %>%
  mutate(species = "Spotted seal",
         date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug", paste0(age_sex,year) %in% select_groups)

peak_ho_df2 <- plot_df2 %>% filter(age_sex != "YOUNG OF YEAR",
                                 yday > 100) %>%
  group_by(species, age_sex, year) %>% slice_max(ho_prob)

plot_df <- plot_df %>% bind_rows(plot_df2)
peak_ho_df <- peak_ho_df %>% bind_rows(peak_ho_df2)

p <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(yday, ho_prob)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.25, 
                  stroke = 0, color = "grey80") +
  geom_pointrange(data = peak_ho_df,
                  mapping = aes(yday, ho_prob, ymin = 0, ymax = ho_prob, color = year),
                  shape = 15) +
  scale_color_brewer(palette = "Paired") +
  scale_x_continuous(breaks = c(74,105,135,166),
                     labels = c("15-Mar", "15-Apr", "15-May", "15-Jun")) +
  facet_grid(age_sex~species) +
  theme(legend.position = "bottom") +
  xlab("") + ylab("haul-out probability") +
  ggtitle("Annual variability in the timing of peak haul-out probability")
 p
```

```{r}
load(here::here('data/seaice_extent.Rdata'))

extent_max <- extent_df %>% 
  group_by(year) %>% 
  slice_max(area_km2_epsg3571) %>% 
  ungroup() %>% 
  rename(max_extent_km2 = area_km2_epsg3571,
         max_extent_yday = yday) %>% 
  dplyr::select(-fdate)

extent_lookup <- extent_df %>% 
  dplyr::select(fdate, area_km2_epsg3571) %>% 
  rename(extent_km2 = area_km2_epsg3571)

peak_ho_df <- peak_ho_df %>% left_join(extent_max, 
                         by = "year") %>% 
  left_join(extent_lookup, by = c("date" = "fdate")) %>% 
  mutate(peak_diff_days = yday - max_extent_yday) %>% 
  distinct() %>% 
  mutate(max_extent_km2 = max_extent_km2/1000)

models <- peak_ho_df %>% 
  group_by(species, age_sex) %>%  
  nest() %>% 
  mutate(model = map(data, ~ lm(max_extent_km2 ~ yday, data = .x))) %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)

hf_adultF <- models %>% filter(species == "Ribbon seal" & age_sex == "ADULT.F") %>% 
  dplyr::select(r.squared, statistic, p.value)
hf_adultM <- models %>% filter(species == "Ribbon seal" & age_sex == "ADULT.M") %>% 
  dplyr::select(r.squared, statistic, p.value)
hf_subadult <- models %>% filter(species == "Ribbon seal" & age_sex == "SUBADULT") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_adultF <- models %>% filter(species == "Spotted seal" & age_sex == "ADULT.F") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_adultM <- models %>% filter(species == "Spotted seal" & age_sex == "ADULT.M") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_subadult <- models %>% filter(species == "Spotted seal" & age_sex == "SUBADULT") %>% 
  dplyr::select(r.squared, statistic, p.value)

```

The timing of peak haul-out probability for ribbon and spotted seals appeared to have only
a limited to no relationship with the amount of yearly maximum sea ice extent. Adult
female and subadult spotted seals show a negative trend line, but there is no indication
that the observed trend is meaningful (spotted seal adult female: $R^{2}$ =
`r pl_adultF$r.squared`, $p$ = `r pl_adultF$p.value`; spotted seal subadults: $R^{2}$ =
`r pl_subadult$r.squared`, $p$ = `r pl_subadult$p.value`). For ribbon seals and adult male
spotted seals, $p$-values were substantially larger than 0.05 (ribbon seal adult females;
$R^{2}$ = `r hf_adultF$r.squared`, $p$ = `r hf_adultF$p.value`; ribbon seal adult males:
$R^{2}$ = `r hf_adultM$r.squared`, $p$ = `r hf_adultM$p.value`; ribbon seal subadults:
$R^{2}$ = `r hf_subadult$r.squared`, $p$ = `r hf_subadult$p.value`; spotted seals adult
males: $R^{2}$ = `r pl_adultM$r.squared`, $p$ = `r pl_adultM$p.value`).

```{r maxextentHO, fig.height = 5.5, include=FALSE, fig.id = "maxextent", fig.cap="Annual timing of peak haul-out behavior analyzed against the yearly maximum area of sea ice extent in the study area. While spotted seal adult females and subadults show a negative trend line, none of the relationships were determined to be significant."}
ggplot(data = peak_ho_df, aes(y = yday, x = max_extent_km2, color = species)) + 
  geom_smooth(method = "lm", alpha = 0.2, aes(fill = species)) +
  geom_point() +
  facet_grid(age_sex~.) +
  scale_y_continuous(breaks = c(125,135,145, 156),
                     labels = c("05-May", "15-May", "25-May", "05-Jun")) +
  scale_color_manual(values = c('#E66100','#5D3A9B'), name = "") +
  scale_fill_manual(values = c('#E66100','#5D3A9B'), name = "") +
  theme(legend.position = "top",
        legend.justification = "right") +
  labs(y = "date of peak haul-out", x = "maximum sea ice extent (km<sup>2</sup> / 1000)",
       title = "The connection between yearly maximum sea ice extent and the \ntiming of peak haul-out probability is limited") +
  theme(axis.title.x = element_markdown())

```

# Discussion

In this paper, we modeled data from bio-logging sensors deployed on bearded, spotted, and
ribbon seals to examine factors affecting haul-out behavior on sea ice in the Bering and
Chukchi seas. Our analysis shows seals use sea ice progressively more through the
spring, with peak use near mid-May to early June before declining again. This pattern
aligns well with what has been previously documented and confirms our bio-logging
deployments are likely capturing population-level behavioral patterns. Seals
preferentially haul out on ice shortly after solar noon, which allows seals to maximize
absorption of solar radiation [@feltz1966]. Interestingly, bearded seals appear to have
two peaks in haul-out activity across the day, one shortly after solar noon, and one
centered near solar midnight. A similar bi-modal pattern has been seen in ringed seals
[@vonduyke2020] and suggests that bearded and ringed seals may be operating under
different constraints than ribbon and spotted seals. Bearded and ringed seals are
distributed across higher latitudes and the extended daylight hours may allow more
flexibility in optimizing resting periods with foraging.

Unlike previous analyses of seal haul-out data (e.g. [@verhoef2010cs], [@conn2014mee]), we
also investigated the influence of sex-age class on haul-out probabilities for all species
except bearded seals because of low sample size. While both ribbon and spotted seals
exhibited a unimodal diel haul-out pattern generally centered around local solar noon,
there were key differences across species, age, and sex that match expectations given what
we know of their ecological behavior. Spotted seals are known to form triads during the
breeding season [@burns2009, @mcclintock2015] where a female and dependent pup are
accompanied on the ice by a suitor male. The male waits for the female to wean the pup and
enter estrus, and fends off any other males that may be interested. Triad formation results
in both males and females spending a large portion of the day hauled out on ice and a
protracted spring haul-out season for both sexes. Because mating occurs in the water,
females may also be less inclined to venture out for foraging trips while still nursing
the pup. We see this reflected in the predicted haul-out behavior, with both males and
females exhibiting a broad distribution of time out of the water throughout the solar day
and the season. Ribbon seals are not known to form triads and our model predicts a
progression of increased haul-out behavior with females starting earlier in the season
than males. Notably, female ribbon seals spend a large portion of the day in the water
during the pupping period, aligning with the hypothesis that ribbon seal females continue
foraging while nursing. In the case of both ribbon and spotted seals, subadults are the
first to begin consistent haul-out behavior and follow a typical phocid pattern where
subadults molt first as they do not have any reproductive constraints.

We also investigated the influence of environmental covariates on haul-out probabilities,
including wind speed, temperature, barometric pressure, precipitation, and wind chill.
These have been investigated for walrus (e.g. Udevitz et al. [@udevitz2009cjz]) and a few
select studies of ice-associated seals [@hamilton2018, @perry2017]. Ribbon seals seemed to
be the most influenced by weather, with wind, temperature, barometric pressure, and
precipitation all being important components of the model. Spotted seals were most
affected by wind and barometric pressures. And, for bearded seals, the model indicated
wind and temperature had the greatest impact. In general, and as might be expected, seals
were more likely to haul-out when daily temperatures were warmer, winds speeds were lower,
barometric pressure was higher, and precipitation was lower. These weather conditions
are likely most beneficial for energetic considerations and may also provide the best
situation for predator detection. These results highlight the importance for haul-out
behavior models (and subsequent calculations of availability during surveys) to
incorporate weather covariates.

Our model uncovered annual deviations in the timing and magnitude of haul-out peaks for
ribbon and spotted seals. The timing of peak haul-out activity appears to fall within a
relatively narrow time window of 3-4 weeks in late May and early June. This remarkable
level of consistency across 15 years is indicative of the adaptations these species have
evolved to synchronize their key life history stages with favorable environmental
conditions (e.g. availability of sea ice, increased oceanographic productivity, favorable
weather conditions). Unfortunately, this may also be indicative of a limited range of
plasticity as climatological conditions change. Additionally, while from an ecological
perspective the haul-out behavior appears consistent, the interannual differences in
timing and magnitude are large enough to have important ramifications on calculations of
abundance and trend.

Previous attempts to estimate the abundance of phocid seals from aerial survey data in the
Bering and Chukchi Seas (e.g. [@bengtson2005pb], [@conn2014mee], [@verhoef2014sm]) have
used estimated haul-out probabilities to correct for the proportion of animals that are in
the water and thus unavailable to be counted during aerial surveys. Although several of
these studies allowed haul-out probabilities to vary by day-of-year and time-of-day,
previous correction factors have not accounted for variability between years, different
weather conditions, and the age-sex class of the sample. In this paper, we have shown that
there can be considerable differences in the number of seals hauled out on ice based on
these factors. We recommend that future abundance analyses employ correction factors that
account for these factors. For instance, it is relatively straightforward to obtain
weather reanalysis products for times and locations that are surveyed and to construct a
relevant correction factor based on predictions of GLMPMs. The most challenging element in
developing availability correction factors is with annual variability. It can be difficult
to get a sufficient sample size to estimate year-specific correction factors, particularly
because research teams would likely need to tag seals and conduct aerial surveys
concurrently which requires considerably more personnel and money. One possible suggestion
is to estimate a "shift" parameter within models for aerial survey counts that allow the
peak of haul-out distributions to be adjusted earlier or later in the year based on the
frequency of counts observed over time. Regardless, researchers should anticipate there
being some unmodeled heterogeneity in availability probability present in abundance
estimates obtained from aerial surveys. This may make trend detection difficult, as one
will not know if moderate differences in abundance estimates are attributable to changes
in abundance or changes in haul-out frequencies.

Predictions of absolute haul-out probability in this paper were somewhat different than
those previously reported for these species, especially for bearded seals. For instance,
Ver Hoef et al. [-@verhoef2014sm)] and Conn et al. [-@conn2014mee] used haul-out
correction factors with maximums of 0.66 for bearded seals, 0.62 for ribbon seals, and
0.54 for spotted seals, where maximums corresponded to times near solar noon in mid-late
May. Applying models that ignore age, sex, and year effects, these probabilities were
0.38, 0.72, and 0.60, respectively, under the current analysis framework. Our current
estimates reflect increased sample sizes in terms of number of animals, but also
improvements to the way data were prepared prior to analysis.

We focused this paper on haul-out behavior of bearded, ribbon, and spotted seals. Ringed
seals are also present in the Bering and Chukchi Seas but exhibit qualitatively different
haul-out behavior. For instance, many ringed seals build subnivean layers in the snow on
top of the sea ice, which they frequently inhabit before snow melt causes their lairs to
collapse [@kelly1990]. Thus, the wet-dry sensor on a satellite tag could indicate that an
animal is hauled out, but it could still be unavailable to be detected because it is in a
lair. We hope to address availability of ringed seals using data from satellite tags,
replicate survey tracks, and auxiliary information about snow depth and timing of melt in
a future study.

# Author Contributions

**Josh M. London**: investigation, conceptualization, methodology, formal analysis,
validation, software, writing: original draft, writing: review and editing, visualization,
and data curation

**Paul B. Conn**: conceptualization, methodology, formal analysis,
software, validation, writing: original draft, writing: review and editing

**Stacie K. Hardy**: investigation, data curation, methodology, validation, writing: review and
editing

**Erin L. Richmond**: data curation, investigation, methodology, validation,
writing: review and editing

**Jay M. Ver Hoef**: conceptualization, methodology, software,
writing: review and editing

**Irina S. Trukhanova**: conceptualization, writing: review
and editing

**Michael F. Cameron**: investigation, project administration, writing: review
and editing

**Peter L. Boveng**: investigation, conceptualization, supervision, project
administration, writing: review and editing

# Data Availability

Data and code will be available as a complete research compendium on GitHub and major versions
archived at Zenodo. Original data sources for telemetry are archived at the United
States Animal Telemetry Network.

# Acknowledgements

We recognize that the species and ecosystems studied in this paper are within the
traditional range of the Inpuiaq and Yupik people. And, we would like to acknowledge and
thank them as the original stewards of these ecosystems. The deployment of bio-logging
devices used in this study were often done in collaboration with Alaska Native
communities. We would like to especially acknowledge the communities of Kotzebue,
Utqiagvik, and Ulguniq and the following individuals: James Adams, Jeff Barger, David
Barr, Wendell Booth, Cyrus Harris, Nereus 'Doc' Harris, Grover Harris, Lee Harris, Tom
Jones, Frank Garfield, Brenda Goodwin, Henry Goodwin, John Goodwin, Pearl Goodwin, Willie
Goodwin, Brett Kirk, Noah Naylor, Virgil Naylor Jr., Virgil Naylor Sr., Dan Savetilik,
Allen Stone, and Randy Toshavik from Kotzebue, Alaska; Billy Adams, Mary-Ellen Ahmaogak,
James Aiken, Tim Aiken, Sarah Coburn, Jason Herreman, Howard Kittick, Gilbert Leavitt,
Isaac Leavitt, J.R. Leavitt, Enoch Oktollik, Shawn Oktollik, Stacey Osborn, Fred Rexford,
Chuck Schaeffer, Ross Schaeffer, Bob Shears, and Joe Skin from the North Slope Borough,
Alaska;

We are grateful for the assistance in catching and sampling seals by James Bailey,
Michelle Barbieri, John Bengtson, Gavin Brady, Vladamir Burkanov, Cynthia Christman, Shawn
Dahle, Rob Delong, Stacy DiRocco, Deb Fauquier, Shannon Fitzgerald, Kathy Frost, Scott
Gende, Tracey Goldstein, Jeff Harris, Markus Horning, John Jansen, Shawn Johnson, Charles
Littnan, Lloyd Lowry, Erin Moreland, Lorrie Rea, Gay Sheffield, Brent Stewart, and Dave
Withrow. We also appreciate the commitment to science and safety by all officers and crew
of the NOAA Ship Oscar Dyson, the NOAA Ship MacArthur II, and the RV Thomas G. Thompson.

Telemetry data from the Alaska Department of Fish and Game and the North Slope Borough
Department of Wildlife Management were important contributions to the findings presented
here. Both agencies collaborate closely with Alaska Native communities and the
contributions of local hunters to the successful capture, deployment, and understanding of seal ecology are
invaluable. Deployments in the western Bering Sea and Sea of Okhotsk were done in
collaboration with Russian colleagues and North Pacific Wildlife.

The findings and conclusions in the paper are those of the author(s) and do not
necessarily represent the views of the National Marine Fisheries Service, NOAA. Any use of
trade, product, or firm names does not imply an endorsement by the U.S. Government.
Funding for this study was provided by the U.S. National Oceanic and Atmospheric
Administration. The field work was conducted under the authority of Marine Mammal
Protection Act Research Permits Nos. 782-1676, 782-1765, 15126, and 19309 issued by the
National Marine Fisheries Service, and Letters of Assurance of Compliance with Animal
Welfare Act regulations, Nos. A/NW 2010-3 and A/NW 2016-1 from the Alaska Fisheries
Science Center/Northwest Fisheries Science Center Institutional Animal Care and Use
Committee.

# References

::: {#refs}
:::