### 2021-12-18 analysis/paper/paper.Rmd
IMPROVE FIG 1
feedback re: fig 1 was that the week/day-of-week organization didn't make sense. So, re-org day-of-year into pseudo 10-day weeks and re-adjusted the plot

---
title: "Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas"
author:
  - Josh M. London:
      email: josh.london@noaa.gov
      institute: [MML]
      correspondence: true
  - Paul B. Conn:
      institute: [MML]
      correspondence: false
  - Stacie K. Hardy:
      institute: [MML]
      correspondence: false
  - Erin L. Richmond:
      institute: [MML]
      correspondence: false
  - Jay M. Ver Hoef:
      institute: [MML]
      correspondence: false
  - Michael F. Cameron:
      institute: [MML]
      correspondence: false
  - Justin Crawford:
      institute: [ADFG]
      correspondence: false
  - Andrew L. Von Duyke:
      institute: [NSB]
      correspondence: false
  - Lori T. Quakenbush:
      institute: [ADFG]
      correspondence: false
  - Peter L. Boveng:
      institute: [MML]
      correspondence: false
institute:
  - MML: Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA
  - ADFG: Arctic Marine Mammals Program, Alaska Department of Fish and Game, Fairbanks, Alaska, USA
  - NSB: Department of Wildlife Management, North Slope Borough, Utqiaġvik, Alaska, USA
output: 
  officedown::rdocx_document:
    reference_docx: ../templates/jml_template.docx
    pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
bibliography: references.json
link-citations: true
csl: ../templates/proceedings-of-the-royal-society-b.csl
always_allow_html: true
---

## Abstract

Ice-associated seals rely on sea ice for a variety of activities, including
breeding, molting, pupping, and resting. In the Arctic, many of these activities
occur in spring (April-June) as sea ice begins to melt and retreat northward.
Rapid acceleration of climate change in Arctic ecosystems is therefore of
concern as the quantity and quality of suitable habitat is forecast to decrease.
In this paper, we use data collected from satellite-linked bio-loggers deployed
between 2005 and 2020 to investigate the seasonal timing and environmental
factors affecting haul-out behavior by ice-associated seals (specifically, bearded,
ribbon, and spotted seals) in the Bering and Chukchi seas. In addition to
providing baseline data on phenology, these data also allow us to quantify
'availability', which is needed to accurately estimate abundance from aerial
survey counts of seals basking on ice (i.e., to correct for the proportion of
animals that are in the water while surveys are conducted). Using generalized
linear mixed pseudo-models to properly account for temporal autocorrelation, we
fit models with covariates of interest (e.g., day-of-year, solar hour, age-sex
class, wind speed, barometric pressure, temperature, precipitation) to examine
their ability to explain variation in hourly haul-out probability. We found evidence
for strong diurnal and within-season patterns in haul-out behavior, as well as
strong weather effects (particularly wind and temperature). In general, seals
were more likely to haul out on ice in the middle of the day and when wind speed
was low and temperatures were higher. Haul-out probability increased through
March and April, peaking in May and early June before declining again. The
timing and frequency of haul-out events also varied based on species and age-sex
class. For ribbon and spotted seals, models with year effects were highly
supported, indicating that the timing and magnitude of haul-out behavior varied
among years. However, we did not find evidence that haul-out timing was linked
to annual sea ice extent. Our analysis emphasizes the importance of accounting
for seasonal and temporal variation in haul-out behavior, as well as associated
environmental covariates, when interpreting the number of seals counted
in aerial surveys.

keywords: availability; generalized linear mixed pseudo-model; haul-out
behavior; phenology; Phocidae; sea ice

```{r setup, include = FALSE}
library(targets)
library(conflicted)
library(tidyverse)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
library(lubridate)
conflict_prefer("month","lubridate")
library(glue)
library(scales)
library(DBI)
library(dbplyr)
library(RPostgres)
library(pingr)
library(sf)
library(flextable)
library(patchwork)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggspatial)
library(ggsci)
library(ggrepel)
library(gghighlight)
library(ggtext)
library(glmmLDTS)
library(mgcv)
library(ragg)

knitr::opts_chunk$set(dev = "ragg_png")

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  include = FALSE,
  fig.path = "../figures/",
  dpi = 400,
  fig.width = 6.5,
  out.width = "100%"
)

knitr::opts_chunk$set(
  tab.cap.style = "Table Caption",
  tab.cap.pre = "Table ",
  tab.cap.sep = ": "
  )

knitr::opts_chunk$set(
  fig.cap.style = "Image Caption",
  fig.cap.pre = "Figure ",
  fig.cap.sep = ": "
  )

options(scipen = 3, digits = 3)

theme_set(theme_minimal(base_family = 'Source Sans Pro'))
theme_replace(plot.title.position = "plot")
```

# Introduction

The global climate disruption is causing considerable reduction in Arctic sea
ice extent, volume, and seasonal presence [@kwok2018;
@meier2014; @overland2021; @wang2017]. These changes have ripple effects on
Arctic organisms, ecosystems, and the human communities who live in the region
[@huntington2020]. Such disruptions are a particular cause of concern for
ice-associated seals (ringed (*Pusa hispida*), bearded (*Erignathus barbatus*),
ribbon seals (*Histriophoca fasciata*), and spotted (*Phoca largha*)) that
depend on spring and early summer sea ice (March-June) in the Bering and Chukchi
seas as a platform for important functions, such as pupping, nursing, breeding,
and molting [@boveng2009; @boveng2013a; @cameron2010; @kelly2010]. Limited data
and large knowledge gaps complicate predictions about the ultimate effects of
changes in sea ice on the behavior, health, abundance, and distribution of these
seals. To date, indices of seal health sampled during periods of declining sea
ice differ regionally [@crawford2015; @harwood2020]. Knowledge about
evolutionary constraints on the timing of reproductive and molting behavior is
generally lacking, so it is difficult to predict the readiness with which
ice-associated seal species will be able to adapt to future changes (e.g., by
adjusting pupping or molting schedules to earlier dates or different locales).
Additionally, trends in abundance of these species are largely unknown, so it is
difficult to assess the effect, if any, declines in sea ice habitat have had, or
will have, on seal densities.

For ice-associated seals, haul-out behavior is tightly linked with sea ice.
Ribbon seals haul out of the water almost exclusively on sea ice and are mostly
pelagic outside the spring pupping, breeding, and molting season [@boveng2018a].
While spotted and bearded seals rest on coastal features, they
strongly prefer sea ice as a resting platform during the spring and early summer
[@frost2018]. The remoteness of the Bering and Chukchi seas mean direct
scientific observation of seal behavior is impractical. Thus, bio-logging
devices are, thus, especially useful tools for collecting key information on
movement and haul-out behavior for these species.

Bio-logging records of time spent out of the water provide valuable data for
identifying covariates that explain variation in haul-out behavior. For
instance, in the Antarctic, Bengtson and Cameron [-@bengtson2004] relied on
bio-logging data to demonstrate greater haul-out propensity in juvenile
crabeater seals (*Lobodon carcinophaga*) than adults, with highest probabilities
in February and at times close to solar noon. In the Arctic, Von Duyke et al.
[-@vonduyke2020] used satellite-linked bio-loggers to corroborate seasonal
changes between diurnal and nocturnal haul-out behavior of ringed seals
previously described by Kelly et al. [-@kelly2010] using VHF radio tags and
direct observation. Bengtson et al. [-@bengtson2005a] documented a higher
propensity for ringed seal basking near solar noon, as did Ver Hoef et al.
[-@verhoef2014a] in an analysis of bearded, ribbon, and spotted seals using much
larger sample sizes. Olnes et al. [-@olnes2020] showed that the proportion of
time bearded seals spent hauled out progressively increased through spring and
summer. And, Ver Hoef et al. [-@verhoef2014a] found haul-out probabilities
increased gradually starting in March and peaking in May and June for bearded,
ribbon, and spotted seals.

While analyzing haul-out patterns clearly has utility for understanding natural
history and ecology, some studies have focused, as we do here, specifically on
quantifying the probability of seals being hauled out as a means of quantifying
seals in the water that are not counted during aerial surveys. These studies
typically use logistic regression-style analyses to estimate the time-specific
probability of being hauled out based on 'wet/dry' data relayed by bio-loggers.
In these models, haul-out probabilities were expressed as a function of
predictive covariates, such as time-of-day, day-of-year, sex, age class, and
environmental conditions (e.g., [@reder2003], [@bengtson2004], [@bengtson2005a],
[@udevitz2009], [@verhoef2014a], and [@southwell2008]). However, sample sizes
have often been insufficient to permit strong inference about life history
and/or seasonal variation in haul-out probabilities. For instance, Bengtson and
Cameron's [-@bengtson2004] study included 5 adult and 2 juvenile crabeater
seals, while Bengtson et al.'s [-@bengtson2005a] study was based on 6 ringed
seals in the Chukchi Sea. These studies were often further limited by logistical
constraints on fieldwork and the attachment duration or operational life
of bio-loggers. In this study, we address some of these limitations by deploying
small bio-loggers designed for longer-term attachment on a subset of the study
individuals. These devices are designed to collect data through the molt period
(when those adhered to the hair would fall off) and, in some situations, provide
multiple years of data.

Ultimately, knowledge of trends in phenology and abundance (or life history
surrogates such as survival and recruitment) will be necessary to make credible
quantitative predictions about the effects of climate disruptions on the
abundance and distribution of Arctic seal populations. Before we can construct a
trend, however, we first require a baseline. Several studies have contributed
estimates of the distribution and abundance of ice-associated seal species in
the Arctic using aerial surveys (e.g., [@bengtson2005a], [@conn2014a], and
[@verhoef2014a]). Such abundance studies were conducted over very large areas and
estimation of absolute abundance required making inference about numerous issues
affecting the observation of seals on ice. These included availability (only
seals basking on ice were available to be counted), detection probability
(observers or automated detection systems may have missed some seals on ice),
species misclassification, and possible disturbance of seals by aircraft
[@conn2014a; @verhoef2014a]. Refining these inferences will improve the
accuracy of abundance estimates in the Arctic.

In this study, we used 16 years of bio-logging data to investigate the haul-out
behavior of 3 species of ice-associated seals in the Bering and Chukchi seas. We
focused our effort on bearded, ribbon, and spotted seals and did not include
ringed seals in the analysis. Ringed seals are unique in their use of snow lairs
in the spring months, and they warrant a separate study that includes additional
environmental covariates, such as the onset of snow melt and snow depth. Our
goals were threefold. First, we wish to establish baseline estimates for the
chronology of haul-out behavior in the critical spring season for each species
across different age and sex classes. Second, we refined estimates of
haul-out availability corrections for aerial surveys in order to improve
estimates of seal abundance. Previously estimated availability correction
factors (e.g., [@bengtson2005a], [@conn2014a], and [@verhoef2014a]) accounted
for variables such as the time-of-day and day-of-year, but did not investigate
meteorological variables that have been shown to influence haul-out behavior of
walruses [@reder2003; @udevitz2009]. Third, we assessed the annual variability
in haul-out timing and possible linkage to changes in the extent of seasonal sea
ice between 2005 and 2020. Our work extends the scope of previous haul-out
analyses, includes the influence of meteorological variability, and investigates
the potential impact of a changing icescape on the behavior of these species.

# Methods

## Data collection

```{r getdata}
targets::tar_load(analysis_data, store = here::here("_targets"))
targets::tar_load(grid, store = here::here("_targets"))
targets::tar_load(deploy_table, store = here::here("_targets"))
```
Data used in our study came from bio-loggers deployed on bearded, ribbon, and
spotted seals in the Bering, Chukchi, and Beaufort seas by multiple
organizations as part of collaborative investigations during 2005 and 2020.
Seals were captured using nets and instrumented during ship-based capture trips
and from coastal Alaska communities. Ship-based capture events occurred during
spring near the southern ice edge in the Bering Sea from 20xx to 20xx (maybe
provide sample sizes for each species). Land-based capture events occurred from
May to October, generally between the coastal communities of Scammon Bay,
Alaska, in the Bering Sea, and Utqiaġvik, Alaska, near the junction of the
Chukchi and Beaufort seas, from 2005 to 2020. We refer readers to the primary
literature for detailed capture and bio-logger attachment methods (Supplemental
Material, Table).

Haul-out behavior data from `r dplyr::n_distinct(deploy_table$speno)`
bio-loggers deployed on bearded, ribbon, and spotted seals were subset to
include only records from 1 March to 15 July between 
`r min(lubridate::year(analysis_data$haulout_dt))` and 
`r max(lubridate::year(analysis_data$haulout_dt))`. Bio-loggers were of the
'SPLASH' or 'SPOT' family of tags developed by Wildlife Computers (Redmond,
Washington, USA) and either adhered to the hair on the seal or attached to the
rear flipper inter-digital webbing. The use of bio-loggers adhered to the back
or head provide some benefits over flipper mounted devices (e.g. increased
satellite transmittal rates, locations at sea) but these are lost during the
annual molt, which limits the duration of haul-out data they provide.
Additionally, bio-loggers attached to the head or dorsal region are often dry
while the seal is at the surface and can slightly bias the hourly percent-dry values
reported by the bio-logger. For this study, in cases where both bio-logger types were
deployed, hourly percent dry observations from the flipper tag were preferred. Sex and
age class (non-dependent *young-of-the-year*, sexually immature *subadults*, and
mature *adults*) were determined at the time of deployment by varying
combinations of length, claw annuli [@burns1969; @burns1983; @mclaren1958a], and pelage characteristics for some
species. Field identification of age class can be inexact, particularly when
discerning subadults from adults. In the case of ribbon seals, pelage
differences between subadults and adults are distinguishable, with subadults
often having less defined ribbons. Bearded seal subadults will often have a
spotted pattern in the pelage that is not seen in adults. Spotted seal pelage is
monomorphic and can not be used to discern age class. Despite these challenges,
we feel the age classifications used in this analysis are useful categories in
testing for age-related effects on haul-out behavior. Seals determined to be
less than one year were classified as young-of-the-year. Subadults are those
seals likely greater than one year of age but not yet sexually mature. Adults
are likely sexually mature and older than four years. For those
bio-loggers deployed on young-of-the-year and transmitting into the next year,
the age class was advanced to subadult on 1 March of the following year. Table
\@ref(tab:tblDeploy) provides a summary of these deployments and data received
from them.

```{r tblDeploy, include=TRUE, tab.id="tblDeploy",tab.cap = "Summary of bio-logger data across seal species and age classification from 1 March to 15 July 2005-2020. Total seal hours represents the sum of available data across all seals."}

deploy_table %>% 
  dplyr::group_by(species, sex, age) %>% 
  dplyr::summarise(
    n = n(),
    seal_hours = sum(n_hours)
  ) %>% 
  dplyr::mutate(
    deploy_stats = glue::glue('{n} ({format(seal_hours, big.mark=",")} seal hours)')
  ) %>% 
  dplyr::select(-c("n","seal_hours")) %>% 
  tidyr::pivot_wider(names_from = age, values_from = c(deploy_stats)) %>% 
  dplyr::ungroup() %>% 
  dplyr::rename(Species = species, Sex = sex, `Young-of-the-Year` = `YOUNG OF YEAR`,
                `Subadult` = SUBADULT, Adult = ADULT) %>% 
  flextable() %>% 
  add_header(`Young-of-the-Year` = "Age Class",
   Subadult = "Age Class", Adult = "Age Class",
   top = TRUE ) %>% 
  merge_h(part = "header") %>% 
  align(align = "center", part = "header") %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  autofit()
```

```{r seal-hours}
create_yday_groups <- function(analysis_data) {
  d <- analysis_data %>%
    tibble::as_tibble() %>%
    dplyr::mutate(yday = lubridate::yday(haulout_dt)) %>%
    dplyr::select(yday) %>%
    dplyr::distinct() %>%
    dplyr::arrange(yday)

  d <- split(d$yday, ceiling(seq_along(d$yday) / 8)) %>%
    dplyr::as_tibble() %>%
    tidyr::pivot_longer(
      cols = 1:18,
      names_to = "col_group",
      values_to = "yday",
      names_transform = list(col_group = as.integer),
      values_transform = list(yday = as.integer)
    ) %>%
    dplyr::arrange(yday) %>% 
    transform(row_group = letters[1:8]) %>% 
    dplyr::filter(col_group < 18)
  return(d)
}

yday_groups <- create_yday_groups(analysis_data)

seal_hours_tbl <- analysis_data %>% 
  tibble::as_tibble() %>% 
  dplyr::mutate(yday = lubridate::yday(haulout_dt)) %>% 
  dplyr::left_join(yday_groups) %>% 
  dplyr::group_by(species, col_group, row_group) %>% 
  dplyr::count()
```

Tags that fall off due to molt or mortality and remain on ice or land can still
send data to satellites; i.e., the bio-logger will be on ice or land and dry, therefore, it will record and transmit data suggesting the seal is hauled out. As such, end times of each deployment were identified by examining
bio-logger locations and dive behavior to determine when bio-loggers ceased providing data
consistent with seal behavior. Data outside of the deployment start and end
times were discarded prior to analysis. After approximately 9 months, 7 devices
deployed on the rear flipper of bearded seals reported implausible hourly percent-dry data.
All data after the first instance of unrealistic values were censored from this
analysis. Figure \@ref(fig:dataCal) shows a calendar-like distribution of all
haul-out data across the study season for each species. Observations for ribbon
and spotted seals are concentrated in the months of May and June due to the
timing of deployment (April and May) and the timing of molt (May and June).
During molt, seals (and their attached bio-loggers) spend more time out of the
water and more data are transmitted. Molt timing also impacts when many
deployments end as any bio-loggers adhered to the hair will fall off. Bearded seal
deployments are more evenly spread with fewer observations in May and June. The
majority of bearded seal deployments start later in the summer and by May bio-loggers
have either fallen off or the batteries are depleted.

```{r seal-hours-calendar, include=TRUE, fig.id = "dataCal", fig.height = 4.75, fig.cap = "Distribution of hourly percent-dry biologger data from 1 March to 15 July for each species. Data are grouped by day-of-year and presented in seal-hours collated across all years between 2005 and 2020. The higher density of data from May and June in ribbon and spotted seals coincides with peak molting when seals (and their attached bio-loggers) are more likely hauled out. Additionally, many bio-logger deployments started in April and May."}

ggplot(data = seal_hours_tbl) + 
  geom_point(aes(x = col_group, y = row_group, color = n, size = n), stroke = 0, shape = 15) +
  scale_size(range = c(1,3.5), breaks = c(100,500,1000,1500)) +
  scale_color_gradient(low = "#6baed6", high = "#08306b") +
  scale_x_continuous(breaks = c(1,4,8,12,16),
                     labels = c("March", "April", "May", "June", "July")) +
  scale_y_discrete(limits = rev) +
  guides(color = "none",
        size = guide_legend(override.aes = list(color = "#2171b5"),
                            title.position = "bottom", title.hjust = 0.5)
        ) +
  facet_grid(species ~ .) + 
  theme(
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom") +
  labs(x = "", y = "", color = "seal-hours",
       size = "seal hours of bio-logger data") +
  ggtitle("Density of haul-out behavior observations varies from March through July",
          subtitle = "data are grouped by day-of-year and presented in seal-hours collated across years (2005-2020)")
```

Haul-out behavior data was recorded and transmitted via the Argos satellite
network as hourly percent-dry timelines. For each hour of a day, the wet/dry
sensor was polled by the tag firmware every few seconds and a percent of the
hour in a dry state was calculated (Figure \@ref(fig:examplePlot)). Hourly
percent-dry data were usually rounded to the nearest 10 percent inclusive of 0
and 100 percent and additional values at 3 and 98 percent. This compression
resulted in additional data transmission as each message consisted of two
complete 24-hour records. Bio-loggers were, generally, deployed and programmed
in a manner to maximize data transmission during the spring pupping and molting
period, however, hourly percent-dry data for days during a deployment were not
successfully transmitted. This is due to a variety of factors including
satellite coverage, tag availability (i.e. tags mounted to the rear flipper are
not available to the surface while the seal is at sea), tag performance, duty
cycling, and extra-terrestrial atmospheric interference.

```{r examplePlot, include=TRUE, fig.id="examplePlot", fig.cap="Haul-out behavior observations recorded by a bio-logger deployment on a ribbon seal over two years during the months of March, April, and May. Areas of the plot indicated with an 'X' represent missing data that were not successfully transmitted from the tag."}

d <- analysis_data %>% 
  filter(speno == "HF2016_1002") %>% 
  dplyr::select(speno, haulout_dt, percent_dry) %>% 
  arrange(haulout_dt) %>% st_drop_geometry() %>% 
  mutate(year = lubridate::year(haulout_dt),
         month = lubridate::month(haulout_dt,label=TRUE),
         day = lubridate::day(haulout_dt),
         hour = lubridate::hour(haulout_dt)) %>% 
  filter(month != 'Jun')

d_deploy <- analysis_data %>% 
  sf::st_drop_geometry() %>% 
  filter(speno == "HF2016_1002") %>% 
  dplyr::select(speno,deploy_dt,haulout_dt,percent_dry) %>% 
  mutate(year = lubridate::year(deploy_dt),
         month = lubridate::month(deploy_dt, label = TRUE),
         day = lubridate::day(deploy_dt),
         hour = lubridate::hour(deploy_dt))

missing_hours <- tibble(haulout_dt = seq(min(d_deploy$haulout_dt),
              max(d_deploy$haulout_dt),by = "hour")) %>% 
  filter(!haulout_dt %in% d_deploy$haulout_dt) %>% 
  filter(lubridate::month(haulout_dt) %in% c(3,4,5)) %>% 
  mutate(percent_dry = 0) %>% 
  mutate(year = lubridate::year(haulout_dt),
         month = lubridate::month(haulout_dt,label=TRUE),
         day = lubridate::day(haulout_dt),
         hour = lubridate::hour(haulout_dt))

p <- ggplot(d,aes(day,hour,fill=percent_dry))+
  geom_tile(color= "white",size=0) + 
  scale_fill_distiller(palette = "YlGnBu",
                       name = "hourly percent dry", aesthetics = "fill",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                           barheight = 0.5, title.hjust = 0.5)
    )

p <- p + facet_grid(year~month, scales = "free_x")
p <- p + scale_x_continuous(breaks = c(5,15,25), expand = c(0,0)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  theme(panel.spacing = unit(0,"line")) +
  ggtitle("Observation records are incomplete because not all data are transmitted",
          subtitle = "percent-dry data for a single ribbon seal across two years") +
  xlab("day of month") + ylab("hour (UTC)") 
p <- p + 
  geom_point(aes(x = day, y = hour), data = missing_hours, 
             color = "black", shape = 4, size=1)
p
```

Of key interest in this study is the relationship between haul-out behavior and
environmental covariates that vary with space and time. We explored the use of a
continuous-time correlated random walk [@devins.johnson2008] movement model to
predict locations at specific times. However, the sparse nature of data from
some bio-loggers, especially those mounted to the rear flipper, resulted in poor
modeling performance or convergence issues. For this study, we calculated a
weighted average daily location where the inverse of the estimated Argos or
FastLoc GPS location error was used for the weight. Each Argos location estimate
was assigned an error radius based on either the categorical location quality (
_3_=250m, _2_=500m, _1_=1500m, _0_=2500m, _A_=2500m, _B_=2500m) or, when
available, the estimated error radius from the Argos Kalman filter algorithm.
Location estimates from FastLoc GPS were all assigned an error radius of 50m.
Any days where haul-out observations were present but location data were missing
we used the last calculated weighted average daily location, and any days where
the location intersected with land were removed from the data set. We recognized
that bearded and spotted seals haul out on land. However, assessing the
relationship between haul-out behavior and environmental covariates and their
availability for aerial surveys on land was outside the scope of this study.
Additionally, any daily locations on land are likely more reflective of
coordinate averaging and measurement error instead of actual use of coastal
features. Figure \@ref(fig:dataMap) shows the spatial distribution weighted
locations with available haul-out behavior data for each species across the
study area.

```{r locs-count-mapdata}
haulout_locs <- analysis_data %>% 
  filter(lubridate::hour(haulout_dt) == 12)

hexgrid <- sf::st_make_grid(st_bbox(haulout_locs) %>% st_as_sfc(), cellsize = 50*1000,
                            what = "polygons", square = FALSE) 
hexgrid <- st_sf(index = 1:length(lengths(hexgrid)), hexgrid)

hexbin <- st_join(haulout_locs, hexgrid, join = st_intersects)

locs_count <- hexgrid %>%
  left_join(
    count(hexbin, index, species) %>%
      as_tibble() %>%
      dplyr::select(species, index, ct=n)
  ) %>% drop_na()
```

```{r dataMap, fig.height = 3.75, fig.id="dataMap", include=TRUE, fig.cap="Spatial distribution of haul-out data during the months of March through July 15 for each of the three species. Data were collated across all years between 2005 and 2020."}

world <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  sf::st_transform(st_crs(locs_count))

ggplot() +
  annotation_spatial(world, 
                     size = 0, 
                     fill = "grey70") +
  layer_spatial(
    data = locs_count, size = 0.125,
    aes(fill = ct)
  ) +
  scale_fill_distiller(type = "seq", palette = "YlOrRd", direction = 1,
                    trans = "log10", aesthetics = "fill",
    name = "number of locations",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15, 
                           barheight = 0.5, title.hjust = 0.5)) + 
  facet_grid(~ species) +
  theme(legend.position = "bottom") +
  ggtitle("Spatial distribution of haul-out behavior data")
```

## Explanatory variables

In addition to sex and age class, we analyzed variables that might help explain
variation in haul-out probabilities. These included day-of-year (for seasonal
effects) and local solar hour (for diurnal effects). Solar hour was calculated
using the {solaR} package [@perpinan2012] within the R statistical environment
[@rcoreteam2021] using the daily locations. We included several meteorological
variables shown to affect haul-out behavior in other Arctic pinnipeds
[@reder2003; @udevitz2009; @perry2017]. In particular, we linked the weighted
average daily locations from our bio-loggers to key weather values from the
North American Regional Reanalysis (NARR) model produced by the National Centers
for Environmental Prediction [@mesinger2006]. The NARR model assimilates
observational data to produce a long-term picture of weather over North America.
Numerous weather variables are made available across the region 8 times daily.
For this study, NARR weather values were subset to the extent of our study area
over the Bering and Chukchi seas at 3-hr intervals based on the native grid
resolution of 32 km (1024 sq. km). The following meteorological variables were
interpolated and assigned to seal locations using a bilinear method: 1) air
temperature at 2m above the earth's surface, 2) wind consisting of northerly and
easterly vector components converted to wind speed using the Euclidean norm, 3)
barometric pressure at sea level, and 4) precipitation (Table
\@ref(tab:tblCovariates)).

Models for all species include the following effects: day-of-year, solar hour,
temperature, wind speed, barometric pressure, precipitation, wind chill
(represented by a *wind:temperature* interaction [@udevitz2009]), and
day-of-year and time-of-day (solar hour) interactions to permit diurnal patterns
to change from March to July. Ribbon and spotted seal models included age-sex
class and interactions between day-of-year and age-sex class, but we omitted
these from bearded seal models due to poor representation of age-sex classes
(Table \@ref(tab:tblDeploy)). Bearded seal models included a latitudinal effect
(and an interaction with day-of-year), since bearded seals occupy a substantial
range and we were interested in possible differences in the timing of haul-out
behavior along a latitudinal gradient. We omitted the latitudinal effect from
ribbon and spotted seal models since, during the spring, these species are most
prevalent near the southern ice edge in the Bering Sea [@conn2014a].

Notably missing from this list of explanatory variables is any spatial-temporal
representation of sea ice concentration, area, or extent. This may seem
counterintuitive when modeling the haul-out behavior of seal species with such a
close association to sea ice; seals haul out in the presence of sea ice, and we
could assess the local concentration of sea ice during these events (see
[@olnes2020]). However, a major focus of this study is to develop models
applicable for aerial survey correction factors and using sea ice as a covariate
would almost certainly bias haul-out predictions towards those seals that are on
or near ice and therefore more likely to haul out. Since aerial surveys can only
detect seals on ice, abundance estimates would be missing a correction for those
seals that are away from ice (e.g., on foraging trips). Additionally, many of
the deployments consisted of a single device attached to the rear flipper of the
seal which do not provide at-sea locations, limiting our ability to fully
evaluate habitat preferences related to sea-ice. It should be noted that the
aerial survey counts are modeled with sea-ice and other habitat parameters. So,
any final abundance estimates that combine models of survey counts and our model
for proportion of seals hauled out on ice do account for the strong association
between these species and sea-ice in the Bering and Chukchi seas. Lastly, our
study was limited to the spring season when seal haul-out tendencies are
strongly influenced by pupping, nursing, breeding behavior, and molt and these
drivers are likely more influential than specific sea-ice concentration.
Crawford et al [-@crawford2019] compared haul-out probability models for ringed
seals and found those that only included season (and not sea-ice concentration)
were the most parsimonious. Further, studies to evaluate habitat selection in
relation sea-ice parameters and other environmental covariates are in progress.
For these reasons, we have elected not to use sea ice concentration as a
predictor for haul-out probability.

We assessed whether the annual variation in maximum spring sea ice extent in the
Bering Sea influences the seasonal peak of seal haul-out behavior. In
particular, we used sea ice concentration data from the Nimbus-7 SMMR and DMSP
SSM/I-SSMIS Passive Microwave Dataset, Version 1 [@cavalieri1996] to calculate
maximum sea ice extent. All sea ice concentration grid cells (25 km^2^) in the
study area with greater than 15% concentration were counted daily to get the total sea
ice extent for each day between 15 February and 15 July across all years.
Maximum spring sea ice extent was simply the largest daily count of grid cells
with greater than 15 percent concentration for each year.

```{r tblCovariates, tab.id = "tblCovariates", include=TRUE, tab.cap = "Explanatory covariates used in analyses of binary haul-out records for bearded, spotted, and ribbon seals. Note that we also considered select interactions (see article text) between these primary covariates.  For instance, wind chill was represented by the interaction _temperature:wind_."}

covariate_tbl <- tibble::tribble(
  ~Covariate, ~Type, ~Source, ~Description,
  "Age-sex class",   "Categorical", "Field Assessment", 
  "young-of-the-year, subadult, adult male and adult female",
  "Hour", "Continuous; Fourier basis", "Bio-logger", 
  "local solar hour using 6 variables of a Fourier-series basis",
  "Day",   "Continuous", "Bio-logger", 
  "linear, quadratic, and cubic effects of day-of-year",
  "Precip", "Continuous", "NARR", 
  "convective precipitation (kg/m2)",
  "Pressure", "Continuous", "NARR", 
  "atmospheric pressure at sea level (kPa)",
  "Temp", "Continuous", "NARR",
  "air temperatures at 2m above the earth’s surface",
  "Wind", "Continuous", "NARR", 
  "northerly and easterly vector components for wind converted into a single wind speed via the Euclidean norm",
  "Northing", "Continuous", "Bio-logger","latitude divided by the mean latitude across all locations (for bearded seals only)"
)

covariate_tbl %>% 
  flextable() %>% 
  font(fontname = "Source Serif Pro Light", part = "all") %>% 
  fontsize(size = 10) %>% 
  set_table_properties(width = .95, layout = "autofit")
```

## Haul-out modeling

```{r modelfits}
tar_load(bearded_fit, store = here::here("_targets"))
tar_load(ribbon_fit, store = here::here("_targets"))
tar_load(spotted_fit, store = here::here("_targets"))
tar_load(spotted_year_fit, store = here::here("_targets"))
tar_load(ribbon_year_fit, store = here::here("_targets"))
```

Haul-out records for seals are often characterized by sequential hours spent
basking on ice alternating with long periods in the water (Figure
\@ref(fig:examplePlot)). Commonly used statistical models for binary data (e.g.
logistic regression) assume independence among responses, an assumption that is
clearly violated if hourly responses are modeled. Any analysis that ignores
temporal dependence in responses will thus have overstated precision
[@betts2006].

To properly account for temporal dependence within a computationally tractable
framework, we used generalized linear mixed pseudo-models (GLMPMs;
[@verhoef2010a]) to model variation in haul-out behavior as a function of (1)
covariate predictors, (2) temporally autocorrelated random effects, and (3)
individual random effects representing heterogeneity in individual behavior. We
used the {glmmLDTS} package [@verhoef2010a] to implement GLMPMs. We explored two
different model formulations for our data, and owing to the large number of
records, we fit separate models to bearded, ribbon, and spotted seal data sets.
In our first model formulation, for each species, we fitted a year-independent
model that predicted average haul-out behavior as a function of demographic,
environmental, seasonal, and diurnal effects. Second, for ribbon and spotted
seals (which had considerably more data than bearded seals), we fitted models
that included all the effects from the first model, but also permitted annual
variation in haul-out timing. This second class of models was used to examine
whether haul-out patterns varied by year and to determine the annual timing of
apparent peaks in haul-out behavior. For both models, the response variable was
the transmitted hourly percent-dry value.

We followed Ver Hoef et al. [-@verhoef2014a] in using linear, quadratic, and
cubic effects of day-of-year to represent temporal changes in behavior. However,
unlike previous models for harbor seals [@london2012] and ice-associated seals
[@verhoef2014a], which treated hour-of-day as a 24-level categorical variable to
capture diurnal cycles, we adopted a continuous formulation based on Fourier
series that provides a flexible model while preserving the inherent circularity
needed for time-of-day effects (i.e., hour 0 should be equal to hour 24). It
also represents hour-of-day with 6 parameters, which is a considerable reduction
when compared to a 24 parameter variable, especially when fitting models with
interactions between hour-of-day and other variables (e.g., age-sex class,
day-of-year). According to this approach, we used the following specification
for hour-of-day effects:

$$
H_t=\alpha_1 \cos(\frac{\pi t}{4}) + \alpha_2 \sin(\frac{\pi t}{4}) + \alpha_3 \cos(\frac{\pi t}{6}) + \alpha_4 \sin(\frac{\pi t}{6}) + \alpha_5 \cos(\frac{\pi t}{12}) + \alpha_6 \sin(\frac{\pi t}{12})
$$ 
where $H_t$ gives the effect for solar hour $t$ and $α_{i}$ are
estimated parameters (regression coefficients).

For the second set of models examining inter-annual variation in sea ice use, we
fitted models with year by day-of-year interactions. However, in this case we
only included *year:day* and *year:day^2^*, omitting the main effects of year as
well as *year:day^3^* interactions because models with the latter effects were
numerically unstable. However, the modeled interactions were sufficient to allow
shifts in haul-out distribution, as one can show mathematically that a simple
horizontal shift in timing of haul-out distributions does not affect the main
effects or cubic terms in a polynomial regression model. Bearded seals were
not included in this examination of inter-annual variation because of limited
data across many years in the study.

A typical model fitting exercise would also include a model selection process.
However, AIC (and similar criteria) is not suitable when using
pseudo-likelihoods, because pseudo-data generated in the model fitting process
[@verhoef2010a] differ between models [@teneyck2018]. After fitting GLMPM
models, we instead used "type III" $F$-tests to calculate $p$-values
[@verhoef2010a] to evaluate model performance and important terms. We also
produced predictions of haul-out behavior as a function of influential
predictors (e.g. solar hour, day-of-year, age-sex). Weather covariates for these
predictions were based on daily or hourly smoothed weather covariate values
across the study region. Such predictions were then used to develop haul-out
probability surfaces, explore conditional effects of weather covariates, and
determine annual peaks in haul-out activity. The timing of peak haul-out
behavior was further used to regress against the annual maximum sea ice extent
in the study region.

# Results

Models omitting year effects suggested that day-of-year, solar hour, age-sex
class, temperature, and wind substantially influenced haul-out behavior of all
three species, with $F$ tests producing $p$-values less than 0.05 for variables
embodying these effects and/or their interactions. Haul-out probabilities
typically increased throughout March and April, reaching a peak in May and early
June before declining again. Diurnal patterns were present, with maximum
haul-out behavior centered around local solar noon.

```{r type3}
ribbon_type3 <- tibble(
  ribbon_fit$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>% 
  column_to_rownames(var = "Effect") 
ribbon_type3 <- setNames(split(ribbon_type3, 
                               seq(nrow(ribbon_type3))), 
                         rownames(ribbon_type3))

spotted_type3 <- tibble(
  spotted_fit$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
spotted_type3 <- setNames(split(spotted_type3, 
                               seq(nrow(spotted_type3))), 
                         rownames(spotted_type3))

bearded_type3 <- tibble(
  bearded_fit$typeIII.hypoth
) %>% 
  mutate(Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
bearded_type3 <- setNames(split(bearded_type3, 
                               seq(nrow(bearded_type3))), 
                         rownames(bearded_type3))
```

## Bearded Seals

Age and sex class were not included in the model for bearded seals due to our
lower sample size for adult and young-of-year age classes. As such,
results are shown for all ages. Bearded seals had a bi-modal
distribution of haul-out probability across the day (Figure
\@ref(fig:beardedHOCal)). In addition to a peak around local solar noon, the
bearded seal model predicts additional haul-out activity around local midnight.
The haul-out behavior is also more protracted throughout the spring season
compared to ribbon and spotted seals.

```{r ebnewdata}
targets::tar_load(bearded_newdata, store = here::here("_targets"))
```

```{r beardedHOCal, fig.height = 3.5, include=TRUE, fig.id = "beardedHOCal", fig.cap="Predicted haul-out probability of bearded seals (all ages and sex classes) from 1 March to 15 July for all and sex classes."}

plot_df <- bearded_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df,aes(day,solar_hour,fill = ho_prob))+
  geom_tile(color= "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Bearded seals rest on sea ice around solar noon and solar midnight")
p
```

Bearded seal haul-out probability appears most influenced by wind
($F_{`r paste(bearded_type3$wind$Num.df, bearded_type3$wind$Den.df, sep=",")`}$
= `r bearded_type3$wind$F.value`; $p$ = `r bearded_type3$wind$Prob.F`)
and temperature
($F_{`r paste(bearded_type3$temp2$Num.df, bearded_type3$temp2$Den.df, sep=",")`}$
= `r bearded_type3$temp2$F.value`; $p$ =
`r bearded_type3$temp2$Prob.F`). Additionally, wind chill was also a
significant influence
($F_{`r paste(bearded_type3$temp2_wind$Num.df, bearded_type3$temp2_wind$Den.df, sep=",")`}$
= `r bearded_type3$temp2_wind$F.value`; $p$ =
`r bearded_type3$temp2_wind$Prob.F`). Any influence of barometric
pressure or precipitation is less apparent.

```{r beardedHOwx,fig.height = 5.5, include=FALSE, fig.id = "beardedHOwx", fig.cap="Variability in predicted haul-out probability of bearded seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df, aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df, aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df, aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df, aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

((p_temp + p_wind) / (p_pressure + p_precip)) +
  plot_annotation(
    title = 'Bearded seal haul-out behavior is less influeced by weather')
```

## Ribbon Seals

```{r hfnewdata}
targets::tar_load(ribbon_newdata, store = here::here("_targets"))
```

Ribbon seals show a pattern of gradually increasing haul-out probability
in April that peaks in late May for subadults and in early June for
adults (Figure \@ref(fig:ribbonHOCal)). The behavior is clearly centered
around local solar noon and expands to other hours later in the season
as seals enter their molting period. Subadults showed an earlier start
and more intense haul-out activity in April and May. The
young-of-the-year records begin after weaning and the model predictions
demonstrate the ontogeny of in-water activities (e.g. diving, foraging)
in May. Adult females have a more protracted haul-out season compared to
males, and more time spent resting in June and July. The model suggests
adult male ribbon seals complete their molt by the end of June. Lastly,
for adults of both sexes there is some indication for a shift to a
crepuscular haul-out pattern in late June and July.

```{r ribbonHOCal, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOCal", fig.cap="Predicted haul-out probability of ribbon seals from 1 March to 15 July for each age and sex class used in the model. Adult females are indicated by 'ADULT.F' and 'ADULT.M' indicates adult males. 'SUBADULT' and 'YOUNG OF YEAR' age classes include both sexes."}

plot_df <- ribbon_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label = TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df, aes(day,solar_hour,fill = ho_prob)) +
  geom_tile(color = "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour = "white")) +
  theme(axis.ticks = element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Ribbon seals rest on sea ice in late spring and around solar noon.")
p
```

The haul-out probability for ribbon seals was strongly influenced by
each of the weather covariates included within the model except
precipitation. The impacts of temperature
($F_{`r paste(ribbon_type3$temp2$Num.df, ribbon_type3$temp2$Den.df, sep=",")`}$
= `r ribbon_type3$temp2$F.value`; $p$ = `r ribbon_type3$temp2$Prob.F`),
wind
($F_{`r paste(ribbon_type3$wind$Num.df, ribbon_type3$wind$Den.df, sep=",")`}$
= `r ribbon_type3$wind$F.value`; $p$ = `r ribbon_type3$wind$Prob.F`),
and barometric pressure
($F_{`r paste(ribbon_type3$pressure$Num.df, ribbon_type3$pressure$Den.df, sep=",")`}$
= `r ribbon_type3$pressure$F.value`; $p$ =
`r ribbon_type3$pressure$Prob.F`) were especially noticeable as ribbon
seals were less likely to haul out at higher winds and lower pressure
values and more likely to haul out when temperatures were
relatively warm. Wind chill
($F_{`r paste(ribbon_type3$temp2_wind$Num.df, ribbon_type3$temp2_wind$Den.df, sep=",")`}$
= `r ribbon_type3$temp2_wind$F.value`; $p$ =
`r ribbon_type3$temp2_wind$Prob.F`) showed a moderately negative
influence on haul-out probability.

Visualizing the marginal effect of each weather covariate on haul-out
probability is difficult in this analysis because of the collinearity
between covariates as well as the temporal variation in covariate values
during the study period. The relationship of each weather covariate with
haul-out probability, averaged over the other weather conditions, is
more variable than simply showing model coefficients would imply. Figure
\@ref(fig:ribbonHOwx) presents the predicted haul-out probability of
ribbon seals across the range of weather conditions encountered in the
observed data. The collinearity and temporal variation are apparent in
the visual patterns that emerge.

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Variability in predicted haul-out probability of ribbon seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  geom_smooth(se = FALSE) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Ribbon seal haul-out behavior is influenced by weather')
```

## Spotted Seals

```{r plnewdata}
targets::tar_load(spotted_newdata, store = here::here("_targets"))
```

Compared to ribbon seals, spotted seals showed a longer spring haul-out season
that is less intensely centered on solar noon (Figure \@ref(fig:spottedHOCal)).
Adults of both sexes spend considerable time in April, May and June hauled out.
In contrast to ribibon seals, adult spotted seal males have a more protracted
haul-out season compared to females, and more time out of the water in June and
July (Figures \@ref(fig:ribbonHOCal) and \@ref(fig:spottedHOCal)). As with
ribbon seals, the young-of-the-year records begin after weaning and the model
predictions demonstrate the ontogeny of in-water activities (e.g. diving,
foraging) in May.

```{r spottedHOCal, fig.height = 5.5, include=TRUE, fig.id = "spottedHOCal", fig.cap="Predicted haul-out probability of spotted seals from 1 March to 15 July for each age and sex class used in the model. Adult females are indicated by 'ADULT.F' and 'ADULT.M' indicates adult males. 'SUBADULT' and 'YOUNG OF YEAR' age classes include both sexes."}

plot_df <- spotted_newdata %>%
  mutate(date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug")

p <- ggplot(plot_df,aes(day,solar_hour,fill=ho_prob))+
  geom_tile(color= "white",size = 0) +
  scale_fill_viridis_c(name = "haul-out probability", aesthetics = "fill",
                       guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                                              barheight = 0.5, title.hjust = 0.5)
  )
p <- p + facet_grid(age_sex~month)
p <- p + scale_x_continuous(breaks = c(1,10,20,30)) +
  scale_y_continuous(breaks = c(4,12,20))
p <- p + theme(legend.position = "bottom") +
  theme(strip.background = element_rect(colour="white")) +
  theme(axis.ticks=element_blank()) +
  xlab("day of month") + ylab("local solar hour") +
  ggtitle("Spotted seals rest on sea ice in late spring and around solar noon.")
p
```

Spotted seal haul-out behavior appears most strongly influenced by
temperature
($F_{`r paste(spotted_type3$temp2$Num.df, spotted_type3$temp2$Den.df, sep=",")`}$
= `r spotted_type3$temp2$F.value`; $p$ = `r spotted_type3$temp2$Prob.F`)
and wind
($F_{`r paste(spotted_type3$wind$Num.df, spotted_type3$wind$Den.df, sep=",")`}$
= `r spotted_type3$wind$F.value`; $p$ = `r spotted_type3$wind$Prob.F`)
with barometric pressure having a moderate influence
($F_{`r paste(spotted_type3$pressure$Num.df, spotted_type3$pressure$Den.df, sep=",")`}$
= `r spotted_type3$pressure$F.value`; $p$ =
`r spotted_type3$pressure$Prob.F`). Spotted seals were less likely to
haul out at higher winds and more likely to be on the ice when
temperatures were relatively warm. Wind chill (*temperature:wind*) and
precipitation were not as influential as the other covariates.
Differences in the magnitude of response between the age-sex classes are
present and consistent across each of the weather covariates (Figure
\@ref(fig:spottedHOwx)).

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Variability in the predicted haul-out probability of spotted seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability. Solid lines are smoothed fits to the predictions and provided to better indicate the marginal effect."}

library(ggsci)

p_temp <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(temp2*27, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("temperature (C)") + ylab("haul-out probability")

p_wind <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(wind, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) ),
                 name = "") +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top") +
  xlab("wind") + ylab("")

p_pressure <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(pressure, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("barometric pressure") + ylab("haul-out probability")

p_precip <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(precip, ho_prob, color = age_sex)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.01, stroke = 0) +
  scale_color_d3(guide = guide_legend(override.aes = list(alpha = 1) )) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "none") +
  xlab("precipitation") + ylab("")

library(patchwork)

p_temp + p_wind + p_pressure + p_precip +
  plot_annotation(
    title = 'Spotted seal haul-out behavior is influenced by weather')
```

## Annual variation in haul-out

```{r hfpvyear}
targets::tar_load(ribbon_newdata_year, store = here::here("_targets"))
targets::tar_load(spotted_newdata_year, store = here::here("_targets"))
```

```{r type3_year}
ribbon_year_type3 <- tibble(
  ribbon_year_fit$typeIII.hypoth
) %>% 
  mutate(# Prob.F = ifelse(Prob.F<0.001, "<0.01",Prob.F),
         Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>% 
  column_to_rownames(var = "Effect") 
ribbon_year_type3 <- setNames(split(ribbon_year_type3, 
                               seq(nrow(ribbon_year_type3))), 
                         rownames(ribbon_year_type3))

spotted_year_type3 <- tibble(
  spotted_year_fit$typeIII.hypoth
) %>% 
  mutate(# Prob.F = ifelse(Prob.F<0.001, "<0.01",Prob.F),
         Prob.F = scales::pvalue(Prob.F),
         Effect = janitor::make_clean_names(Effect)) %>%  
  column_to_rownames(var = "Effect")  
spotted_year_type3 <- setNames(split(spotted_year_type3, 
                               seq(nrow(spotted_year_type3))), 
                         rownames(spotted_year_type3))
```

The second set of models that included annual variation in haul-out
patterns uncovered significant contributions for linear and quadratic
interactions between day and year in both ribbon seals (day $\times$
year,
$F_{`r paste(ribbon_year_type3$day_year$Num.df, ribbon_year_type3$day_year$Den.df, sep=",")`}$
= `r ribbon_year_type3$day_year$F.value`; $p$ =
`r ribbon_year_type3$day_year$Prob.F`; day^2^:year,
$F_{`r paste(ribbon_year_type3$day2_year$Num.df, ribbon_year_type3$day2_year$Den.df, sep=",")`}$
= `r ribbon_year_type3$day2_year$F.value`; $p$ =
`r ribbon_year_type3$day2_year$Prob.F`) and spotted seals (day $\times$
year,
$F_{`r paste(spotted_year_type3$day_year$Num.df, spotted_year_type3$day_year$Den.df, sep=",")`}$
= `r spotted_year_type3$day_year$F.value`; $p$ =
`r spotted_year_type3$day_year$Prob.F`; day^2^:year,
$F_{`r paste(spotted_year_type3$day2_year$Num.df, spotted_year_type3$day2_year$Den.df, sep=",")`}$
= `r spotted_year_type3$day2_year$F.value`; $p$ =
`r spotted_year_type3$day2_year$Prob.F`). Predicted distributions of
haul-out activity were largely unimodal, but varied some among and
within years with respect to the timing and magnitude of haul-out peaks
(Figure \@ref(fig:annualHO)). It is important to note that predicted
variation in annual haul-out patterns likely reflects both process error
and sampling variability. While we did remove any years where only one
deployment in a species + age-sex group was present, there are still
some years where the pattern shown is informed by a small number of
individuals that may not represent population level processes.

```{r annualHO, fig.height = 5.5, include=TRUE, fig.id = "annualHO", fig.cap="Annual variability in the timing of peak haul-out probability for ribbon and spotted seals across 10 years. Predictions are shown for local solar noon and under smoothed weather conditions. Only those groups (age-sex + year) that included observations from more than one seal are shown. Additionally, any groups where the underlying data started after 1 June or ended before 1 May are not included."}

select_groups <- ribbon_year_fit$dataset %>%
  filter(age_sex != "YOUNG OF YEAR") %>%
  group_by(age_sex, year) %>%
  summarise(n = n_distinct(speno), group_id = paste0(age_sex,year)) %>%
  filter(n >1) %>% pull(group_id)

plot_df <- ribbon_newdata_year %>%
  mutate(species = "Ribbon seal",
         date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug", paste0(age_sex,year) %in% select_groups)

peak_ho_df <- plot_df %>% filter(age_sex != "YOUNG OF YEAR",
                                 yday > 100) %>%
  group_by(species, age_sex, year) %>% slice_max(ho_prob)

select_groups <- spotted_year_fit$dataset %>%
  filter(age_sex != "YOUNG OF YEAR") %>%
  group_by(age_sex, year) %>%
  summarise(n = n_distinct(speno), group_id = paste0(age_sex,year)) %>%
  filter(n >1) %>% pull(group_id)

plot_df2 <- spotted_newdata_year %>%
  mutate(species = "Spotted seal",
         date = lubridate::as_date(yday, origin = "2015-01-01"),
         month = lubridate::month(date,label=TRUE),
         day = lubridate::day(date)) %>%
  filter(month != "Aug", paste0(age_sex,year) %in% select_groups)

peak_ho_df2 <- plot_df2 %>% filter(age_sex != "YOUNG OF YEAR",
                                 yday > 100) %>%
  group_by(species, age_sex, year) %>% slice_max(ho_prob)

plot_df <- plot_df %>% 
  bind_rows(plot_df2) %>% 
  mutate(year = forcats::fct_relevel(.$year, sort))
peak_ho_df <- peak_ho_df %>% 
  bind_rows(peak_ho_df2) %>% ungroup() %>% 
  mutate(year = forcats::fct_relevel(.$year, sort))

pal <- c(RColorBrewer::brewer.pal(12, "Paired"), "black")

p <- ggplot(plot_df %>% filter(age_sex != "YOUNG OF YEAR"), aes(yday, ho_prob)) +
  geom_pointrange(aes(ymin = lower95, ymax = upper95), alpha = 0.25, 
                  stroke = 0, color = "grey80") +
  geom_pointrange(data = peak_ho_df,
                  mapping = aes(yday, ho_prob, ymin = 0, ymax = ho_prob, color = year),
                  shape = 15) +
  scale_color_discrete(type = pal) +
  scale_x_continuous(breaks = c(74,105,135,166),
                     labels = c("15-Mar", "15-Apr", "15-May", "15-Jun")) +
  facet_grid(age_sex~species) +
  theme(legend.position = "bottom") +
  xlab("") + ylab("haul-out probability") +
  ggtitle("Annual variability in the timing of peak haul-out probability")
 p
```

```{r}
load(here::here('data/seaice_extent.Rdata'))

extent_max <- extent_df %>% 
  group_by(year) %>% 
  slice_max(area_km2_epsg3571) %>% 
  ungroup() %>% 
  rename(max_extent_km2 = area_km2_epsg3571,
         max_extent_yday = yday) %>% 
  dplyr::select(-fdate)

extent_lookup <- extent_df %>% 
  dplyr::select(fdate, area_km2_epsg3571) %>% 
  rename(extent_km2 = area_km2_epsg3571)

peak_ho_df <- peak_ho_df %>% left_join(extent_max, 
                         by = "year") %>% 
  left_join(extent_lookup, by = c("date" = "fdate")) %>% 
  mutate(peak_diff_days = yday - max_extent_yday) %>% 
  distinct() %>% 
  mutate(max_extent_km2 = max_extent_km2/1000)

models <- peak_ho_df %>% 
  group_by(species, age_sex) %>%  
  nest() %>% 
  mutate(model = map(data, ~ lm(max_extent_km2 ~ yday, data = .x))) %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)

hf_adultF <- models %>% filter(species == "Ribbon seal" & age_sex == "ADULT.F") %>% 
  dplyr::select(r.squared, statistic, p.value)
hf_adultM <- models %>% filter(species == "Ribbon seal" & age_sex == "ADULT.M") %>% 
  dplyr::select(r.squared, statistic, p.value)
hf_subadult <- models %>% filter(species == "Ribbon seal" & age_sex == "SUBADULT") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_adultF <- models %>% filter(species == "Spotted seal" & age_sex == "ADULT.F") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_adultM <- models %>% filter(species == "Spotted seal" & age_sex == "ADULT.M") %>% 
  dplyr::select(r.squared, statistic, p.value)
pl_subadult <- models %>% filter(species == "Spotted seal" & age_sex == "SUBADULT") %>% 
  dplyr::select(r.squared, statistic, p.value)

```

The timing of peak haul-out probability for ribbon and spotted seals
appeared to have a limited to no relationship with the amount of
yearly maximum sea ice extent. Adult female and subadult spotted seals
show a negative trend line, but there is no indication that the observed
trend is meaningful (spotted seal adult female: $R^{2}$ =
`r pl_adultF$r.squared`, $p$ = `r pl_adultF$p.value`; spotted seal
subadults: $R^{2}$ = `r pl_subadult$r.squared`, $p$ =
`r pl_subadult$p.value`). For ribbon seals and adult male spotted seals,
$p$-values were substantially larger than 0.05 (ribbon seal adult
females; $R^{2}$ = `r hf_adultF$r.squared`, $p$ = `r hf_adultF$p.value`;
ribbon seal adult males: $R^{2}$ = `r hf_adultM$r.squared`, $p$ =
`r hf_adultM$p.value`; ribbon seal subadults: $R^{2}$ =
`r hf_subadult$r.squared`, $p$ = `r hf_subadult$p.value`; spotted seals
adult males: $R^{2}$ = `r pl_adultM$r.squared`, $p$ =
`r pl_adultM$p.value`).

```{r maxextentHO, fig.height = 5.5, include=FALSE, fig.id = "maxextent", fig.cap="Annual timing of peak haul-out behavior analyzed against the yearly maximum area of sea ice extent in the study area. While spotted seal adult females and subadults show a negative trend line, none of the relationships were determined to be significant."}

ggplot(data = peak_ho_df, aes(y = yday, x = max_extent_km2, color = species)) + 
  geom_smooth(method = "lm", alpha = 0.2, aes(fill = species)) +
  geom_point() +
  facet_grid(age_sex~.) +
  scale_y_continuous(breaks = c(125,135,145, 156),
                     labels = c("05-May", "15-May", "25-May", "05-Jun")) +
  scale_color_manual(values = c('#E66100','#5D3A9B'), name = "") +
  scale_fill_manual(values = c('#E66100','#5D3A9B'), name = "") +
  theme(legend.position = "top",
        legend.justification = "right") +
  labs(y = "date of peak haul-out", x = "maximum sea ice extent (km<sup>2</sup> / 1000)",
       title = "The connection between yearly maximum sea ice extent and the \ntiming of peak haul-out probability is limited") +
  theme(axis.title.x = element_markdown())

```

# Discussion

We modeled data from bio-loggers deployed on bearded, ribbon, and spotted seals
to examine factors affecting haul-out behavior on sea ice in the Bering and
Chukchi seas. Our analysis shows all three species of seal haul out
progressively more through the spring and peak near mid-May to early June before
declining again. This pattern aligns well with what has been previously
documented and confirms our haul-out data are likely capturing population-level
behavioral patterns. Seals preferentially haul out on ice shortly after solar
noon, which allows seals to maximize absorption of solar radiation thought to
facilitate the molting process [@feltz1966]. Interestingly, bearded seals appear
to have two peaks in haul-out activity within a day, one shortly after solar
noon, and one centered near solar midnight. A similar bi-modal pattern has been
seen in ringed seals [@vonduyke2020] and suggests that bearded and ringed seals
may be operating under different constraints than ribbon and spotted seals.
Bearded and ringed seals are distributed across higher latitudes and the
extended daylight hours may allow more flexibility in optimizing resting periods
with foraging. Other factors such as polar bear predation (which is rare for
ribbon and spotted seals in the Bering Sea) may also explain differing haul-out
patterns. Bearded seals also showed a more protracted season of haul-out
behavior compared to ribbon and spotted seals. This aligns with findings from
Thometz et al. [@thometz2021] who observed a mean molting period of 119$\pm$2
days and a relatively stable resting metabolic rate during that time. While
ribbon seals were not considered in the same study, spotted and ringed seals
underwent molt periods of 33$\pm$4 and 28$\pm$6 days and exhibited increased
resting metabolic rates.

Unlike previous analyses of seal haul-out behavior (e.g. [@verhoef2010a],
[@conn2014a]), we also investigated the influence of sex-age class on haul-out
probabilities for all species except bearded seals because of low sample size.
While both ribbon and spotted seals exhibited a unimodal diel haul-out pattern
generally centered around local solar noon, there were key differences across
species, age, and sex that match expectations given what we know of their
ecological behavior. Ribbon seals are not known to form triads and our model
predicts a progression of increased haul-out behavior with females starting
earlier in the season than males. Notably, female ribbon seals spend a large
portion of the day in the water during the pupping period, aligning with the
hypothesis that ribbon seal females continue foraging while nursing. Spotted
seals are known to form triads during the breeding season [@burns2009,
@mcclintock2015] where a female and dependent pup are accompanied on the ice by
a suitor male. The male waits for the female to wean the pup and enter estrus,
and fends off any other potential suitor males. Triad formation results in both
males and females spending a large portion of the day hauled out on ice and a
protracted spring haul-out season for both sexes. Because mating occurs in the
water, females may also be less inclined to interrupt their haul-out with
foraging trips while still nursing the pup. We see this reflected in the
predicted haul-out behavior, with both males and females exhibiting a broad
distribution of time out of the water throughout the solar day and the season.
In the case of both ribbon and spotted seals, subadults are the first to begin
consistent haul-out behavior and follow a typical phocid pattern where subadults
molt first as they do not have any reproductive constraints.

We also investigated the influence of environmental covariates on haul-out
probabilities, including wind speed, temperature, barometric pressure,
precipitation, and wind chill. These have been investigated for walruses (e.g.
Udevitz et al. [-@udevitz2009]) and a few select studies of ice-associated seals
[@hamilton2018, @perry2017]. Ribbon seals seemed to be the most influenced by
weather, with wind, temperature, barometric pressure, and precipitation all
being important components of the model. Spotted seals were most affected by
wind and barometric pressures. And, for bearded seals, the model indicated wind
and temperature had the greatest impact. In general, and as might be expected,
seals were more likely to haul-out when daily temperatures were warmer, winds
speeds were lower, barometric pressure was higher, and precipitation was lower.
These weather conditions are likely most beneficial for energetic considerations
and may also provide the best situation for predator detection. These results
highlight the importance for haul-out behavior models (and subsequent
calculations of availability during surveys) to incorporate weather covariates.

Our models detected small deviations in the timing and magnitude of annual peaks
in haul-out behavior for ribbon and spotted seals. The timing of peak haul-out
activity appears to fall within a relatively narrow time window of 3-4 weeks in
late May and early June. This consistency across 15 years is likely a reflection
of the relationship between a critical photoperiod and the timing of important
life history stages [@bronson2009; @temte1994]. However, along with a critical
photoperiod, ribbon and spotted seals are dependent upon the presence of sea ice
for pupping and molt. We didn't find any support in our models for a
relationship between the timing of peaks in haul-out behavior and the amount of
yearly maximum sea ice. This could indicate that, while the extent of spring sea
ice in the Bering sea varied widely during our study period, seals were still
able to locate sea ice and haul out. We should expect, however, that some
minimal threshold in the spatial extent or density of sea ice will have a
meaningful impact on the timing of peak haul-out behavior -- if there is no sea
ice, seals will not haul out. Additionally, while from an ecological perspective
the haul-out behavior appears consistent, the interannual differences in timing
and magnitude are large enough to have important ramifications on calculations
of abundance and trend. Those ramifications will only be exacerbated if climate
variability amplifies those interannual differences.

Previous attempts to estimate the abundance of phocid seals from aerial survey
data in the Bering and Chukchi seas (e.g. [@bengtson2005a], [@conn2014a],
[@verhoef2014a]) have used estimated haul-out probabilities to correct for the
proportion of animals that are in the water and thus unavailable to be counted.
Although several of these studies allowed haul-out probabilities to vary by
day-of-year and time-of-day, they have not accounted for variability among
years, in weather conditions, and in the age-sex class of the sample. In this
paper, we have shown that there can be considerable differences in the number of
seals hauled out on ice based on these factors. We recommend that future
abundance analyses employ availability models that account for these factors. For
instance, it is relatively straightforward to obtain weather reanalysis products
for times and locations that are surveyed and to construct a relevant correction
factor based on predictions of GLMPMs. The most challenging element in
developing availability correction factors is with annual variability. It can be
difficult to get a sufficient sample size to estimate year-specific correction
factors, particularly because research teams would likely need to tag seals and
conduct aerial surveys concurrently which requires considerably more personnel
and money. One possible suggestion is to estimate a "shift" parameter within
models for aerial survey counts that allow the peak of haul-out distributions to
be adjusted earlier or later in the year based on the frequency of counts
observed over time. Regardless, researchers should anticipate there being some
unmodeled heterogeneity in availability probability present in abundance
estimates obtained from aerial surveys. This may make trend detection difficult,
as one will not know if moderate differences in abundance estimates are
attributable to changes in abundance or changes in haul-out behavior.

Predictions of absolute haul-out probability in this paper were somewhat
different than those previously reported for these species, especially for
bearded seals. For instance, Ver Hoef et al. [-@verhoef2014a)] and Conn et al.
[-@conn2014a] used haul-out correction factors with maximums of 0.66 for bearded
seals, 0.62 for ribbon seals, and 0.54 for spotted seals, where maximums
corresponded to times near solar noon in mid-late May. Applying models that
ignore age, sex, and year effects, these probabilities were 0.38, 0.72, and
0.60, respectively, under the current analysis framework. Our current estimates
estimates of haul-out probability reflect increased sample sizes in terms of
number of animals, but also improvements to the way data were prepared prior to
analysis.

We focused this paper on haul-out behavior of bearded, ribbon, and spotted
seals. Ringed seals are also present in the Bering and Chukchi seas but exhibit
a unique complicating factor. Adult ringed seals build subnivean lairs in the
snow on top of the sea ice, where they haul out and pup until snow melt causes
their lairs to collapse [@kelly1990]. Thus, the wet-dry sensor on a bio-logger
could indicate that an animal is hauled out, but if it is within a lair, it is
not available to be detected during an aerial survey. We hope to address
availability of ringed seals using data from satellite tags, replicate survey
tracks, and auxiliary information about snow depth and timing of melt in a
future study.

# Author Contributions

**Josh M. London**: investigation, conceptualization, methodology,
formal analysis, validation, software, writing: original draft, writing:
review and editing, visualization, and data curation

**Paul B. Conn**: conceptualization, methodology, formal analysis,
software, validation, writing: original draft, writing: review and
editing

**Stacie K. Hardy**: investigation, data curation, methodology,
validation, writing: review and editing

**Erin L. Richmond**: data curation, investigation, methodology,
validation, writing: review and editing

**Jay M. Ver Hoef**: conceptualization, methodology, software, writing:
review and editing

**Irina S. Trukhanova**: conceptualization, writing: review and editing

**Michael F. Cameron**: investigation, project administration, writing:
review and editing

**Peter L. Boveng**: investigation, conceptualization, supervision,
project administration, writing: review and editing

# Data Availability

Data and code will be available as a complete research compendium on
GitHub and major versions archived at Zenodo. Original data sources for
telemetry are archived at the United States Animal Telemetry Network.

# Acknowledgements

We recognize that the species and ecosystems studied in this paper are within
the traditional range of the Inpuiaq and Yupik people. And, we would like to
acknowledge and thank them as the original stewards of these ecosystems. The
deployment of bio-logging devices used in this study were often done in
collaboration with Alaska Native seal hungers and their communities. We would
like to especially acknowledge the communities of Kotzebue, Koyuk, Utqiaġvik,
and Ulguniq (Wainwright) and the following individuals: James Adams, Jeff
Barger, David Barr, Wendell Booth, Cyrus Harris, Nereus 'Doc' Harris, Grover
Harris, Lee Harris, Tom Jones, Frank Garfield, Brenda Goodwin, Henry Goodwin,
John Goodwin, Pearl Goodwin, Willie Goodwin, Brett Kirk, Noah Naylor, Virgil
Naylor Jr., Virgil Naylor Sr., Dan Savetilik, Allen Stone, and Randy Toshavik
from Kotzebue, Alaska; Merlin Henry from Koyuk, Alaska; Billy Adams, Mary-Ellen
Ahmaogak, James Aiken, Tim Aiken, Sarah Coburn, Jason Herreman, Howard Kittick,
Gilbert Leavitt, Isaac Leavitt, J.R. Leavitt, Enoch Oktollik, Shawn Oktollik,
Stacey Osborn, Fred Rexford, Chuck Schaeffer, Ross Schaeffer, Bob Shears, and
Joe Skin from the North Slope Borough, Alaska;

We are grateful for the assistance in catching and sampling seals by
James Bailey, Michelle Barbieri, John Bengtson, Gavin Brady, Vladamir
Burkanov, Cynthia Christman, Shawn Dahle, Rob Delong, Stacy DiRocco, Deb
Fauquier, Shannon Fitzgerald, Kathy Frost, Scott Gende, Tracey
Goldstein, Jeff Harris, Markus Horning, John Jansen, Shawn Johnson,
Charles Littnan, Lloyd Lowry, Erin Moreland, Mark Nelson, Lorrie Rea, Gay Sheffield,
Brent Stewart, and Dave Withrow. We also appreciate the commitment to
science and safety by all officers and crew of the NOAA Ship Oscar
Dyson, the NOAA Ship MacArthur II, and the RV Thomas G. Thompson.

Telemetry data from the Alaska Department of Fish and Game (ADFG) and the North
Slope Borough Department of Wildlife Management (NSB) were important
contributions to the findings presented here. Both agencies collaborate
closely with Alaska Native communities and the contributions of local
hunters to the successful capture, deployment, and understanding of seal
ecology are invaluable. Deployments in the western Bering Sea and Sea of
Okhotsk were done in collaboration with Russian colleagues and North
Pacific Wildlife.

The findings and conclusions in the paper are those of the author(s) and do not
necessarily represent the views of the National Marine Fisheries Service, NOAA.
Any use of trade, product, or firm names does not imply an endorsement by the
U.S. Government. Funding for this study was provided by the U.S. National
Oceanic and Atmospheric Administration. The field work was conducted under the
authority of Marine Mammal Protection Act Research Permits Nos. 782-1676,
782-1765, 15126, and 19309 issued by the National Marine Fisheries Service, and
Letters of Assurance of Compliance with Animal Welfare Act regulations, Nos.
A/NW 2010-3 and A/NW 2016-1 from the Alaska Fisheries Science Center/Northwest
Fisheries Science Center Institutional Animal Care and Use Committee (IACUC). ).
Funding to ADFG for tagging seals was provided by the Bureau of Ocean Energy
Management (No. M13PC0015) and the Office of Naval Research (No.
N00014-16-1-3019). ADFG and NSB field work was covered by Research Permits Nos.
358-1585, 358-1787, 15324, and 20466 and by ADF&G IACUC permits Nos. 06-16,
09-21, 2014-03, 2015-25, 2016-23, 0027-2017-27, 0027-2018-29, 0027-2019-041.

# References

::: {#refs}
:::