### 2021-12-22 README.Rmd
UPDATE README
trying to include the targets dependency graph

---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4638221.svg)](https://doi.org/10.5281/zenodo.4638221)
<!-- badges: end -->

## Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas
Josh M. London^1^,✉, Paul B. Conn^1^, Stacie K. Hardy^1^, Erin L. Richmond^1^, Jay M. Ver Hoef^1^, Michael F. Cameron^1^, Justin Crawford^2^, Andrew L. Von Duyke^3^, Lori T. Quakenbush^2^, and Peter L. Boveng^1^

1. Marine Mammal Laboratory, Alaska Fisheries Science Center, NOAA Fisheries, Seattle, Washington, USA
2. Arctic Marine Mammals Program, Alaska Department of Fish and Game, Fairbanks, Alaska, USA
3. Department of Wildlife Management, North Slope Borough, Utqiaġvik, Alaska, USA

✉ Correspondence: Josh M. London <josh.london@noaa.gov>

This repository contains the research compendium in support of the above titled paper.
The manuscript is currently in preparation and the text, code, analysis, and results
are under active development.

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `London_HauloutBehavior.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: figures](/analysis/figures): figures generated
  - [:file\_folder: scripts](/analysis/scripts): R scripts in support of analysis.
  - [:file\_folder: templates](/analysis/templates): template files and scripts.

The **data** directory contains

  - [:file\_folder: data](/data): data used for model fits and final model output

### {targets} and {renv} packages

This repository relies heavily on the {targets} and {renv} packages for management
of analysis pipelines and reproducibility.

  - [:file\_folder: R](/R): pipeline functions developed in support of target pipeline
  - [:page\_facing\_up: _targets.R](/_targets.R): target pipeline
  - [:page\_facing\_up: renv.lock](/renv.lock): renv lockfile

The dependency graph for our pipeline is shown below

```{r, echo=FALSE}
library(targets)
targets::tar_glimpse()
```

### NOAA README

This repository is a scientific product and is not official communication of the National Oceanic and Atmospheric Administration, or the United States Department of Commerce. All NOAA GitHub project code is provided on an ‘as is’ basis and the user assumes responsibility for its use. Any claims against the Department of Commerce or Department of Commerce bureaus stemming from the use of this GitHub project will be governed by all applicable Federal law. Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by the Department of Commerce. The Department of Commerce seal and logo, or the seal and logo of a DOC bureau, shall not be used in any manner to imply endorsement of any commercial product or activity by DOC or the United States Government.