### 2022-03-29 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

<!-- badges: start -->
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4638221.svg)](https://doi.org/10.5281/zenodo.4638221)
<!-- badges: end -->

## Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas
Josh M. London^1^,✉, Paul B. Conn^1^, Stacie K. Hardy^1^, Erin L. Richmond^1^, Jay M. Ver Hoef^1^, Michael F. Cameron^1^, Justin Crawford^2^, Andrew L. Von Duyke^3^, Lori T. Quakenbush^2^, and Peter L. Boveng^1^

1. Marine Mammal Laboratory, Alaska Fisheries Science Center, National Marine Fisheries Service, NOAA, Seattle, Washington, USA
2. Arctic Marine Mammals Program, Alaska Department of Fish and Game, Fairbanks, Alaska, USA
3. Department of Wildlife Management, North Slope Borough, Utqiaġvik, Alaska, USA

✉ Correspondence: Josh M. London <josh.london@noaa.gov>

This repository serves as the research compendium in support of the above titled paper.
As the manuscript works its way through peer review and the publications process,
additional reproducibility features (e.g. Dockerfile, MyBinder.org) and
an R/Shiny application to explore results will be made developed and made available.

Major releases of the research compendium are published and archived with Zenodo [https://doi.org/10.5281/zenodo.4638221](https://doi.org/10.5281/zenodo.4638221)

## Draft Manuscript Under Active Development

Please note this manuscript is still in draft form and under active development.
Changes to results, code, and the manuscript are likely and this **should not
be cited without author permission**. We plan to provide a preprint to 
[bioRxiv](https://www.biorxiv.org/) once internal NOAA reviews and final 
revisions are completed.

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered PDF,
    [London_HauloutBehavior.pdf](https://github.com/jmlondon/berchukHaulout/blob/main/analysis/paper/London_HauloutBeahvior.pdf), suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: figures](/analysis/figures): figures generated
  - [:file\_folder: templates](/analysis/templates): template files and scripts.

The **data** directory contains

  - [:file\_folder: data](/data): data used for model fits and final model output
  - [:file\folder: data_raw](/data_raw): data provided by partners from ADFG and NSB

### {targets} and {renv} packages

This repository relies heavily on the {targets} and {renv} packages for management
of analysis pipelines and reproducibility.

  - [:file\_folder: R](/R): pipeline functions developed in support of target pipeline
  - [:page\_facing\_up: _targets.R](/_targets.R): target pipeline
  - [:page\_facing\_up: renv.lock](/renv.lock): renv lockfile

The dependency graph for our pipeline is shown below

![](./tar_glimpse.png)

## Acknowledgements

We recognize that the species and ecosystems we studied are within the ancestral
and present-day environs of the Inupiat and Yup'ik people who, through many
uncredited contributions of traditional knowledge, provided early western
naturalists and scientists with much of what gets described as the ‘basic
biology’ of Arctic seals. The deployment of bio-logging devices used in this
study were often done in collaboration with Alaska Native seal hunters near
their communities. We would like to especially acknowledge the communities of
Kotzebue, Koyuk, Nome, Nuiqsut, Scammon Bay, St. Michael, Utqiaġvik, and Ulguniq
(Wainwright) and the following individuals: James Adams, Jeff Barger, David
Barr, Wendell Booth, Cyrus Harris, Nereus 'Doc' Harris, Grover Harris, Lee
Harris, Tom Jones, Frank Garfield, Brenda Goodwin, Henry Goodwin, John Goodwin,
Pearl Goodwin, Willie Goodwin, Brett Kirk, Noah Naylor, Virgil Naylor Jr.,
Virgil Naylor Sr., Dan Savetilik, Chuck Schaeffer, Ross Schaeffer, Allen Stone,
and Randy Toshavik from Kotzebue, Alaska; Merlin Henry from Koyuk, Alaska; Tom
Gray from Nome; Vernon Long and Richard Tukle from Nuiqsuit, Alaska; Morgan
Simon, River Simon, and Al Smith from Scammon Bay, Alaska; Alex Niksik Jr. from
St. Michael, Alaska; Billy Adams, James Aiken, Tim Aiken, Howard Kittick,
Gilbert Leavitt, Isaac Leavitt, J.R. Leavitt, and Joe Skin from Utqiaġvik,
Alaska; Mary Ellen Ahmaogak, Enoch Oktollik, Shawn Oktollik, Stacey Osborn, and
Fred Rexford from Ulguniq, Alaska.

We are grateful for the assistance in catching and sampling seals by Ryan Adam,
James Bailey, Michelle Barbieri, John Bengtson, Gavin Brady, Vladamir Burkanov,
Cynthia Christman, Sarah Coburn, Shawn Dahle, Rob Delong, Stacy DiRocco, Deb
Fauquier, Shannon Fitzgerald, Kathy Frost, Scott Gende, Tracey Goldstein, Jeff
Harris, Jason Herreman, Markus Horning, John Jansen, Shawn Johnson, Charles
Littnan, Lloyd Lowry, Brett McClintock, Erin Moreland, Mark Nelson, Justin
Olnes, Lorrie Rea, Bob Shears, Gay Sheffield, Brent Stewart, Dave Withrow, and
Heather Ziel. We also appreciate the commitment to science and safety by all
officers and crew of the NOAA Ship Oscar Dyson, the NOAA Ship MacArthur II, and
the RV Thomas G. Thompson.

Telemetry data from the Alaska Department of Fish and Game (ADFG) and the North
Slope Borough Department of Wildlife Management (NSB) were important
contributions to the findings presented here. Both agencies collaborate closely
with Alaska Native communities and the contributions of local hunters to the
successful capture, deployment, and understanding of seal ecology are
invaluable. Deployments in the western Bering Sea were done in collaboration
with Russian colleagues and North Pacific Wildlife.

## Inspiration

I would like to provide special acknowledgement to Ben Marwick and Carl Boetinger.
This research compendium borrows heavily from Ben's [{rrtools}](https://github.com/benmarwick/rrtools)
package. Carl's example research compendiums ([https://github.com/cboettig/noise-phenomena](https://github.com/cboettig/noise-phenomena) and
[https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes)) were
also of great use when adapting the structure to meet my needs.