### 2023-02-02 analysis/paper/reviewer_response.md
initial reviewer responses

---
editor_options: 
  markdown: 
    wrap: 72
---

We would like to thank each of the reviewers for taking their time to
read and offer suggestions for improving the quality and readability of
the manuscript. Here, we respond directly to each of the suggested edits
and comments. These responses are organized by reviewer.

## Reviewer 1 (A. Fishbach)

> In their general comments, the reviewer mentioned that they would like
> to see a more direct framing of the problem and to see a specific
> example of how the modeled availability would have an impact on
> population abundance estimates.

We address this by ...

> **Title**\
> The reviewer suggests removing *and aerial survey detectability* from
> the title of the manuscript and suggests adding the seasonal time
> period our study focused on.

We have changed the title to: Spring haul-out behavior of seals in the
Bering and Chukchi seas

> **Abstract**\
> *Can you make a statement in your abstract that use of modeled haulout
> behavior using covariates available from regional reananalysis data
> improved the confidence intervals on availability estimates and that
> this would improve aerial survey abundance estimation efforts?*

We don't feel comfortable making a precision comparison, simply because
there are too many differences between the analyses, including (1) the
data sets are different (the present study has substantially more
records), and (2) the model structures are different (the present study
has more parameters). As such, it will be difficult to ascribe a
specific reason to precision increases or decreases.

> *Please use direct language. You are not making findings on how seals
> rely on sea ice, rather you are improving methods to estimate their
> availability for improved aerial survey population estimation efforts.
> Here you are focused on modeling ice sea availability to aerial survey
> so that aerial survey counts may be appropriately interpreted. Please
> be clear that you are considering ice seal behaviors that may be
> telemtered as "dry out of water behaviors" that may be interpreted as
> behaviors that place the seal on top of the ice and available for
> observation by an aerial survey.*

The opening sentence of the abstract is a statement regarding the
reliance of ice-associated seals on sea-ice for a range of important
life history behaviors. We think this statement is important
introduction for readers who are not familiar with ice-associated seals.
We do agree with the reviewer's suggestion for more direct language and
have made edits to the beginning of the abstract with this in mind.

> **Abstract**
>
> *Did you use previously published data, or did you collect it?  Be
> specific using an active voice so that readers may understand your
> effort.*
>
> *For better flow, please indicate the behavioral data logging time
> scale here when you describe the data collection, not later when you
> speak to the analytical construct.*

We have edited this to simply state *data from satellite-linked
bio-loggers* ... we have also altered this sentence to specifically
mention the hourly timescale for the bio-logger data

> *The term "basking" implies a subset of activities that ice seals may
> engage in when hauled out on sea ice and not an activity that can be
> directly telemetered using your data logger.  Yet, what you are
> describing is "haul-out behavior".  Please use this term here.*

edited to remove use of 'basking'

> *Can you help your naive reader understand why you consider these
> covariates to be of interest and why others were not considered.*

The main body of the paper includes this discussion. Given the limited
space and purpose of the abstract we feel it best to simply list the
covariates of interest

> **Introduction**
>
> *Please use specific terms and not metaphors so that second-language
> English readers may better understand your writing.*

We have edited 'ripple effects' to 'tangible effects'

> *What about human communities that don't live in the region, but that
> send large LNG tankers and other vessel through the region?  The loss
> of sea ice has effects on those communities and activities too.*

There is a seemingly endless list of impacts of loss of sea ice to human
communities (and ecosystems) beyond the Arctic region to the rest of the
world. However, for this sentence, we wanted to keep the reader focused
on the Arctic region.

> *Please use direct language you mean biologically important
> functions?*

The term 'life history function' is a commonly used, direct term that
refers to biologically important functions. To ensure the reader is
clearly aware of which behaviors/functions we are referring to we
provide specific examples within the sentence.

> *What does "readiness" mean in the context of a marine mammal?*

This is a good catch and we have edited the sentence to, now, read:

"Knowledge about evolutionary constraints on the timing of reproductive
and molting behavior is generally lacking, so it is difficult to predict
**how or if** ice-associated seal species **might adapt** to future
changes (e.g., by adjusting pupping or molting schedules to earlier
dates or different locales)"

> *Perhaps you mean to offer a topic sentence, stating "How
> ice-assoicated seals use sea ice as a haulout platform varies betweens
> species".*

We have adopted this suggestion

> There is so much to state about ice-associated seals in the Pacific
> Arctic that this reference to Antarctic work seems out of place here.

We have moved this reference to later in the paragraph so the first
references are about Arctic work. While the example is from the
Antarctic, it is an important, early example of analysis using
bio-logger data to estimate haul-out behavior.

> *Do you mean age-class? The term life-history implies knowledge of
> previous behavioral efforts or occupancy of regions of interest.  
> Here you are specifically interested in age-class.*

We have edited to use the term 'demographic'