### 2023-02-04 analysis/paper/reviewer_response.md
additional edits re: rev #1

---
editor_options: 
  markdown: 
    wrap: 72
---

We would like to thank each of the reviewers for taking their time to
read and offer suggestions for improving the quality and readability of
the manuscript. Here, we respond directly to each of the suggested edits
and comments. These responses are organized by reviewer.

## Reviewer 1 (A. Fishbach)

> In their general comments, the reviewer mentioned that they would like
> to see a more direct framing of the problem and to see a specific
> example of how the modeled availability would have an impact on
> population abundance estimates.

We address this by ...

> **Title**\
> The reviewer suggests removing *and aerial survey detectability* from
> the title of the manuscript and suggests adding the seasonal time
> period our study focused on.

We have changed the title to: Spring haul-out behavior of seals in the
Bering and Chukchi seas

> **Abstract**\
> *Can you make a statement in your abstract that use of modeled haulout
> behavior using covariates available from regional reananalysis data
> improved the confidence intervals on availability estimates and that
> this would improve aerial survey abundance estimation efforts?*

We don't feel comfortable making a precision comparison, simply because
there are too many differences between the analyses, including (1) the
data sets are different (the present study has substantially more
records), and (2) the model structures are different (the present study
has more parameters). As such, it will be difficult to ascribe a
specific reason to precision increases or decreases.

> *Please use direct language. You are not making findings on how seals
> rely on sea ice, rather you are improving methods to estimate their
> availability for improved aerial survey population estimation efforts.
> Here you are focused on modeling ice sea availability to aerial survey
> so that aerial survey counts may be appropriately interpreted. Please
> be clear that you are considering ice seal behaviors that may be
> telemtered as "dry out of water behaviors" that may be interpreted as
> behaviors that place the seal on top of the ice and available for
> observation by an aerial survey.*

The opening sentence of the abstract is a statement regarding the
reliance of ice-associated seals on sea-ice for a range of important
life history behaviors. We think this statement is important
introduction for readers who are not familiar with ice-associated seals.
We do agree with the reviewer's suggestion for more direct language and
have made edits to the beginning of the abstract with this in mind.

> **Abstract**
>
> *Did you use previously published data, or did you collect it?  Be
> specific using an active voice so that readers may understand your
> effort.*
>
> *For better flow, please indicate the behavioral data logging time
> scale here when you describe the data collection, not later when you
> speak to the analytical construct.*

We have edited this to simply state *data from satellite-linked
bio-loggers* ... we have also altered this sentence to specifically
mention the hourly timescale for the bio-logger data

> *The term "basking" implies a subset of activities that ice seals may
> engage in when hauled out on sea ice and not an activity that can be
> directly telemetered using your data logger.  Yet, what you are
> describing is "haul-out behavior".  Please use this term here.*

edited to remove use of 'basking'

> *Can you help your naive reader understand why you consider these
> covariates to be of interest and why others were not considered.*

The main body of the paper includes this discussion. Given the limited
space and purpose of the abstract we feel it best to simply list the
covariates of interest

> **Introduction**
>
> *Please use specific terms and not metaphors so that second-language
> English readers may better understand your writing.*

We have edited 'ripple effects' to 'tangible effects'

> *What about human communities that don't live in the region, but that
> send large LNG tankers and other vessel through the region?  The loss
> of sea ice has effects on those communities and activities too.*

There is a seemingly endless list of impacts of loss of sea ice to human
communities (and ecosystems) beyond the Arctic region to the rest of the
world. However, for this sentence, we wanted to keep the reader focused
on the Arctic region.

> *Please use direct language you mean biologically important
> functions?*

The term 'life history function' is a commonly used, direct term.

Here is a definition from Oxford:

"The life history of an organism is its pattern of survival and reproduction,
along with the traits that directly affect survival and the timing or amount of
reproduction. Rates of survival and reproduction can be estimated across age
classes, or across different stages in organisms with complex life cycles. Life
history traits include growth rate; age and size at sexual maturity; the
temporal pattern or schedule of reproduction; the number, size, and sex ratio of
offspring; the distribution of intrinsic or extrinsic mortality rates (e.g.,
patterns of senescence); and patterns of dormancy and dispersal."

To ensure the reader is clearly aware of which life-history behaviors/functions
we are referring to we provide specific examples within the sentence.

> *What does "readiness" mean in the context of a marine mammal?*

This is a good catch and we have edited the sentence to, now, read:

"Knowledge about evolutionary constraints on the timing of reproductive
and molting behavior is generally lacking, so it is difficult to predict
**how or if** ice-associated seal species **might adapt** to future
changes (e.g., by adjusting pupping or molting schedules to earlier
dates or different locales)"

> *Perhaps you mean to offer a topic sentence, stating "How
> ice-assoicated seals use sea ice as a haulout platform varies betweens
> species".*

We have adopted this suggestion

> There is so much to state about ice-associated seals in the Pacific
> Arctic that this reference to Antarctic work seems out of place here.

We have moved this reference to later in the paragraph so the first
references are about Arctic work. While the example is from the
Antarctic, it is an important, early example of analysis using
bio-logger data to estimate haul-out behavior.

> *Do you mean age-class? The term life-history implies knowledge of
> previous behavioral efforts or occupancy of regions of interest.  
> Here you are specifically interested in age-class.*

We have edited to use the term 'demographic'

> Please help your reader understand why are you doing your analysis and not 
> offer a wide and jargon rich inventory of concepts that may be considered.

> Please keep your introduction focused on establishing the knowledge gap that 
> you are seeking to fill.   Please do not lead your reader to the numerous other 
> knowledge gaps that are not being addressed by your current study.

We have removed the sentence in question

> Please use the term "hauled-out".  Neither the biologger nor the aerial 
> survey is able to distinguish among the various seal behaviors that result in 
> them being hauled out and available for view by an aerial survey.

noted, and the terms 'hauled out', 'to haul out', etc is used consistently
throughout the manuscript

> You have indicate that you would consider sea ice extent.  Here you introduce 
> the concept of icescape.  Please introduce this concept earlier and indicate 
> how you will incorporate this into your analysis.

changed `icescape` to `sea-ice extent`

> Please use an active voice for clarity.

This sentence has been changed to:

"For this study we used haul-out behavior data and location estimates from
bio-loggers deployed on bearded, ribbon, and spotted seals in the Bering,
Chukchi, and western Beaufort seas by multiple organizations as part of 
collaborative investigations from 2005 through 2020."

> You summarize the years during which land-based studies applied bio-loggers.  
> Please do the same for the offshore deployments.

This sentence now reads:

"Ship-based capture events occurred during spring near the southern ice edge in
the Bering Sea between 2005 and 2018."

> Please provide the geographic context of your tag deployments and data 
> collection using a plot similar to that used by Jay et al. 
> (2012:  10.3354/meps10057 figure 2).  Please consider using a tabulation that 
> enables readers to understand the geographic context of the deployments 
> similar to that used by Jay et al. 2012, table 1.

We feel the spatial distribution of the actual transmitted data -- compared to
the deployment location -- is most informative and relevant for the audience's
understanding of how our data are distributed spatially and how well data
represents the range of each species. We provide this visually in Figure 3 and
adding an additional layer to that figure would to show deployments would be
an additional complication. To ensure the audience as a clear understanding
of our study area, we have added an additional figure with a map of the Bering
and Chukchi sea region that indicates generally where deployments started. We
have also added a more detailed figure to the supplemental section with each
location specifically marked.

> Please use active voice for all actions that you took in your data collection, 
> preparation and analysis.

Noted and addressed

> Please move this to the discussion.  This is a summary of your data findings, 
> not a description of your methods.

We can see the reviewer's point. However, we do not feel this topic is a summary 
of our data findings in this paper. Instead, our view
is that this information provides the context and rationale for our
methodological choice regarding bio-logger type and attachment methods. It is
also important regarding the analytic methods -- preference for flipper tag
haul-out data over those from a tag adhered to the dorsal/head. If we were to 
move these important points to the discussion, we worry those points would be 
left disconnected.

> This is a discussion of your methods.  This is not your methods.  Please state 
> how the data was collected in your Methods section and discuss how data 
> collection have been uncertain or biased in your discussion section.

> Please move to your discussion.

We appreciate this suggestion from the reviewer and we have moved large
portions of this paragraph to the discussion. However, some information
regarding age/sex class determination is left here so the reader has context
for Table 1.

> Please indicate how this was determined during data collection.

We have edited this sentence some to clearly communicate that field 
determination of reproductive status (and, thus ideal delineation of
sub-adult and adult age class) was not possible.

> Please allow your readers the opportunity to understand the geographic 
> context of these deployments through use of symbols that maybe used to 
> cross-reference the deployment locations in a map that indicates deployment 
> locations along with filtered movement tracks.

As previously addressed, there are now two figures available that provide
geographic context of the deployments. Figure 3 shows the geographic
distribution of the data from the deployments. The paper is not presenting or
discussing movement ecology and, therefore, we do not present filtered
movement tracks. We keep the focus on haul-out behavior data.

> Please consider bar plot similar to that used by Jay et al. 
> (2012: 10.3354/meps10057 figure 3), replacing the region with the seal species.

This figure has been updated with this suggestion in mind

> Please state clearly how you determined the end of a deployment.  Here you 
> indicate various elements of the haulout chronology, movement and dive 
> chronology that may have been used, however not reader could replicate your 
> methods base don this text.

We've added the following text to the paragraph:

"For example, a data record that ends with several consecutive
days (~10+ days) of 100% dry observations and with locations indicating the tag
was on land would be truncated to not include the final stretch of 100% dry
observations. The vast majority of deployments end with the device detaching in the 
water and the deployment end date is obvious. There is no perfect algorithm for
identifying deployment end dates and each deployment in question must be considered 
separately. While not perfect, we are confident our reliance on expert opinion and
examination of multiple data streams provides the best option."

> Deployment activity is known to disrupt normal pinniped behavior.  Did you 
> censure some period of data collection during and after the deployment to 
> ensure that that disruptions from the deployment activity did not bias your 
> analysis?

For these phocid species, we have no empirical data (or even anecdotal 
observational data) on how captures may or may not disrupt 'normal' behavior.
This is likely to be influenced by the capture method, how long the animal
was restrained, whether the animal was with a dependent pup, whether the
animal was molting, environmental variables such as wind and temperature, etc
etc. Given all these unknowns, we have not applied a censure to the beginning of
our deployment data. We can't 100% ensure this choice doesn't result in some
level of bias. However, given the large sample size and duration of deployments
we expect any bias imposed is negligible.

> This is a result.
> Please indicate the method that you used to distinguish "plausible" from 
> "implausible" telemetered seal behavior.  In your Results section, you may 
> indicate how method was able to identify occasions of implausible seal behavior 
> and any patterns that arose relative to tag model or attachment.

This sentence has been moved to the results section. Addtionally, we've provided
a brief explanation of implausible. For these devices, the percent-dry timeline
reported 100% dry for several weeks followed by an additional period of time
that was more typical of expected bearded seal behavior. During the extended dry
time the locations indicated movement of the tag that would not be simply
drifting ice. We also observed increasing transmission rates. The working (i.e.
unconfirmed guess) is that the wet/dry sensor surface became obscured by some
corrosive element such that when in the water conductivity was prevented and,
thus, recorded as dry.

> Please move to the Results section.
> Please move the discussion of how molt timing affects termination of tag 
> deployments to the discussion section.

Figure 1 (original) has been moved to the results section along with additional
descriptive text. We have also added mention to the Discussion section.