### 2023-02-07 analysis/paper/reviewer_response.md
more edits

---
editor_options: 
  markdown: 
    wrap: 72
---

We would like to thank each of the reviewers for taking their time to
read and offer suggestions for improving the quality and readability of
the manuscript. Here, we respond directly to each of the suggested edits
and comments. These responses are organized by reviewer.

## Reviewer 1 (A. Fishbach)

> In their general comments, the reviewer mentioned that they would like
> to see a more direct framing of the problem and to see a specific
> example of how the modeled availability would have an impact on
> population abundance estimates.

We address this by ...

> **Title**\
> The reviewer suggests removing *and aerial survey detectability* from
> the title of the manuscript and suggests adding the seasonal time
> period our study focused on.

We have changed the title to: Spring haul-out behavior of seals in the
Bering and Chukchi seas

> **Abstract**\
> *Can you make a statement in your abstract that use of modeled haulout
> behavior using covariates available from regional reananalysis data
> improved the confidence intervals on availability estimates and that
> this would improve aerial survey abundance estimation efforts?*

We don't feel comfortable making a precision comparison, simply because
there are too many differences between the analyses, including (1) the
data sets are different (the present study has substantially more
records), and (2) the model structures are different (the present study
has more parameters). As such, it will be difficult to ascribe a
specific reason to precision increases or decreases.

> *Please use direct language. You are not making findings on how seals
> rely on sea ice, rather you are improving methods to estimate their
> availability for improved aerial survey population estimation efforts.
> Here you are focused on modeling ice sea availability to aerial survey
> so that aerial survey counts may be appropriately interpreted. Please
> be clear that you are considering ice seal behaviors that may be
> telemtered as "dry out of water behaviors" that may be interpreted as
> behaviors that place the seal on top of the ice and available for
> observation by an aerial survey.*

The opening sentence of the abstract is a statement regarding the
reliance of ice-associated seals on sea-ice for a range of important
life history behaviors. We think this statement is important
introduction for readers who are not familiar with ice-associated seals.
We do agree with the reviewer's suggestion for more direct language and
have made edits to the beginning of the abstract with this in mind.

> **Abstract**
>
> *Did you use previously published data, or did you collect it?  Be
> specific using an active voice so that readers may understand your
> effort.*
>
> *For better flow, please indicate the behavioral data logging time
> scale here when you describe the data collection, not later when you
> speak to the analytical construct.*

We have edited this to simply state *data from satellite-linked
bio-loggers* ... we have also altered this sentence to specifically
mention the hourly timescale for the bio-logger data

> *The term "basking" implies a subset of activities that ice seals may
> engage in when hauled out on sea ice and not an activity that can be
> directly telemetered using your data logger.  Yet, what you are
> describing is "haul-out behavior".  Please use this term here.*

edited to remove use of 'basking'

> *Can you help your naive reader understand why you consider these
> covariates to be of interest and why others were not considered.*

The main body of the paper includes this discussion. Given the limited
space and purpose of the abstract we feel it best to simply list the
covariates of interest

> **Introduction**
>
> *Please use specific terms and not metaphors so that second-language
> English readers may better understand your writing.*

We have edited 'ripple effects' to 'tangible effects'

> *What about human communities that don't live in the region, but that
> send large LNG tankers and other vessel through the region?  The loss
> of sea ice has effects on those communities and activities too.*

There is a seemingly endless list of impacts of loss of sea ice to human
communities (and ecosystems) beyond the Arctic region to the rest of the
world. However, for this sentence, we wanted to keep the reader focused
on the Arctic region.

> *Please use direct language you mean biologically important
> functions?*

The term 'life history function' is a commonly used, direct term.

Here is a definition from Oxford:

"The life history of an organism is its pattern of survival and
reproduction, along with the traits that directly affect survival and
the timing or amount of reproduction. Rates of survival and reproduction
can be estimated across age classes, or across different stages in
organisms with complex life cycles. Life history traits include growth
rate; age and size at sexual maturity; the temporal pattern or schedule
of reproduction; the number, size, and sex ratio of offspring; the
distribution of intrinsic or extrinsic mortality rates (e.g., patterns
of senescence); and patterns of dormancy and dispersal."

To ensure the reader is clearly aware of which life-history
behaviors/functions we are referring to we provide specific examples
within the sentence.

> *What does "readiness" mean in the context of a marine mammal?*

This is a good catch and we have edited the sentence to, now, read:

"Knowledge about evolutionary constraints on the timing of reproductive
and molting behavior is generally lacking, so it is difficult to predict
**how or if** ice-associated seal species **might adapt** to future
changes (e.g., by adjusting pupping or molting schedules to earlier
dates or different locales)"

> *Perhaps you mean to offer a topic sentence, stating "How
> ice-assoicated seals use sea ice as a haulout platform varies betweens
> species".*

We have adopted this suggestion

> There is so much to state about ice-associated seals in the Pacific
> Arctic that this reference to Antarctic work seems out of place here.

We have moved this reference to later in the paragraph so the first
references are about Arctic work. While the example is from the
Antarctic, it is an important, early example of analysis using
bio-logger data to estimate haul-out behavior.

> *Do you mean age-class? The term life-history implies knowledge of
> previous behavioral efforts or occupancy of regions of interest.  
> Here you are specifically interested in age-class.*

We have edited to use the term 'demographic'

> Please help your reader understand why are you doing your analysis and
> not offer a wide and jargon rich inventory of concepts that may be
> considered.

> Please keep your introduction focused on establishing the knowledge
> gap that you are seeking to fill. Please do not lead your reader to
> the numerous other knowledge gaps that are not being addressed by your
> current study.

We have removed the sentence in question

> Please use the term "hauled-out". Neither the biologger nor the aerial
> survey is able to distinguish among the various seal behaviors that
> result in them being hauled out and available for view by an aerial
> survey.

noted, and the terms 'hauled out', 'to haul out', etc is used
consistently throughout the manuscript

> You have indicate that you would consider sea ice extent. Here you
> introduce the concept of icescape. Please introduce this concept
> earlier and indicate how you will incorporate this into your analysis.

changed `icescape` to `sea-ice extent`

> Please use an active voice for clarity.

This sentence has been changed to:

"For this study we used haul-out behavior data and location estimates
from bio-loggers deployed on bearded, ribbon, and spotted seals in the
Bering, Chukchi, and western Beaufort seas by multiple organizations as
part of collaborative investigations from 2005 through 2020."

> You summarize the years during which land-based studies applied
> bio-loggers.\
> Please do the same for the offshore deployments.

This sentence now reads:

"Ship-based capture events occurred during spring near the southern ice
edge in the Bering Sea between 2005 and 2018."

> Please provide the geographic context of your tag deployments and data
> collection using a plot similar to that used by Jay et al. (2012:
> 10.3354/meps10057 figure 2). Please consider using a tabulation that
> enables readers to understand the geographic context of the
> deployments similar to that used by Jay et al. 2012, table 1.

We feel the spatial distribution of the actual transmitted data --
compared to the deployment location -- is most informative and relevant
for the audience's understanding of how our data are distributed
spatially and how well data represents the range of each species. We
provide this visually in Figure 3 and adding an additional layer to that
figure would to show deployments would be an additional complication. To
ensure the audience as a clear understanding of our study area, we have
added an additional figure with a map of the Bering and Chukchi sea
region that indicates generally where deployments started. We have also
added a more detailed figure to the supplemental section with each
location specifically marked.

> Please use active voice for all actions that you took in your data
> collection, preparation and analysis.

Noted and addressed

> Please move this to the discussion. This is a summary of your data
> findings, not a description of your methods.

We can see the reviewer's point. However, we do not feel this topic is a
summary of our data findings in this paper. Instead, our view is that
this information provides the context and rationale for our
methodological choice regarding bio-logger type and attachment methods.
It is also important regarding the analytic methods -- preference for
flipper tag haul-out data over those from a tag adhered to the
dorsal/head. If we were to move these important points to the
discussion, we worry those points would be left disconnected.

> This is a discussion of your methods. This is not your methods. Please
> state how the data was collected in your Methods section and discuss
> how data collection have been uncertain or biased in your discussion
> section.

> Please move to your discussion.

We appreciate this suggestion from the reviewer and we have moved large
portions of this paragraph to the discussion. However, some information
regarding age/sex class determination is left here so the reader has
context for Table 1.

> Please indicate how this was determined during data collection.

We have edited this sentence some to clearly communicate that field
determination of reproductive status (and, thus ideal delineation of
sub-adult and adult age class) was not possible.

> Please allow your readers the opportunity to understand the geographic
> context of these deployments through use of symbols that maybe used to
> cross-reference the deployment locations in a map that indicates
> deployment locations along with filtered movement tracks.

As previously addressed, there are now two figures available that
provide geographic context of the deployments. Figure 3 shows the
geographic distribution of the data from the deployments. The paper is
not presenting or discussing movement ecology and, therefore, we do not
present filtered movement tracks. We keep the focus on haul-out behavior
data.

> Please consider bar plot similar to that used by Jay et al. (2012:
> 10.3354/meps10057 figure 3), replacing the region with the seal
> species.

This figure has been updated with this suggestion in mind

> Please state clearly how you determined the end of a deployment. Here
> you indicate various elements of the haulout chronology, movement and
> dive chronology that may have been used, however not reader could
> replicate your methods base don this text.

We've added the following text to the paragraph:

"For example, a data record that ends with several consecutive days
(\~10+ days) of 100% dry observations and with locations indicating the
tag was on land would be truncated to not include the final stretch of
100% dry observations. The vast majority of deployments end with the
device detaching in the water and the deployment end date is obvious.
There is no perfect algorithm for identifying deployment end dates and
each deployment in question must be considered separately. While not
perfect, we are confident our reliance on expert opinion and examination
of multiple data streams provides the best option."

> Deployment activity is known to disrupt normal pinniped behavior. Did
> you censure some period of data collection during and after the
> deployment to ensure that that disruptions from the deployment
> activity did not bias your analysis?

For these phocid species, we have no empirical data (or even anecdotal
observational data) on how captures may or may not disrupt 'normal'
behavior. This is likely to be influenced by the capture method, how
long the animal was restrained, whether the animal was with a dependent
pup, whether the animal was molting, environmental variables such as
wind and temperature, etc etc. Given all these unknowns, we have not
applied a censure to the beginning of our deployment data. We can't 100%
ensure this choice doesn't result in some level of bias. However, given
the large sample size and duration of deployments we expect any bias
imposed is negligible.

> This is a result. Please indicate the method that you used to
> distinguish "plausible" from "implausible" telemetered seal behavior.
> In your Results section, you may indicate how method was able to
> identify occasions of implausible seal behavior and any patterns that
> arose relative to tag model or attachment.

This sentence has been moved to the results section. Additionally, we've
provided a brief explanation of implausible along with a reference for
more details. For these devices, the percent-dry timeline reported 100%
dry for several weeks followed by an additional period of time that was
more typical of expected bearded seal behavior. During the extended dry
time the locations indicated movement of the tag that would not be
simply drifting ice. We also observed increasing transmission rates. The
working (i.e. unconfirmed guess) is that the wet/dry sensor surface
became obscured by some corrosive element such that when in the water
conductivity was prevented and, thus, recorded as dry.

> Please move to the Results section. Please move the discussion of how
> molt timing affects termination of tag deployments to the discussion
> section.

This figure has been moved to the results section along with additional
descriptive text. We have left the discussion regarding molt timing and
termination of the data record here to better communicate the point to
the reader and prevent an isolated mention in the discussion that has no
anchor.

> Thus, the a successful transmission of data was required at no greater
> than 48 h intervals.

We're not sure how to interpret this comment. The main point of the
sentence is to explain the rationale/benefits of transmitting the
limited selection of percent dry values. We have expanded the text to
provide more details regarding the caching of messages and data
transmission.

> Do you mean "atmospheric" magnetic interference.\
> Presumably "extra-terrestrial atmospheric" events (such as events on
> Venus or Mars) played no role in the transmission of the data signal
> to the Argos receivers in orbit around Earth.

We were using a less strict definition of extra-terrestrial to refer to
interference that wasn't Earth/surface-based. We've edited to
'atmospheric interference'

> This note belongs in the Discussion, not the Results.

We feel this note is important to acknowledge in the methods because it
speaks to the rationale for our methodological choices (bio-logger type,
programming choices, analytical choices).

> Do you mean that each cell "encompasses" and area of 50 km\^2?

We have edited to the reviewer's suggested word 'encompasses'

> Please use the active voice.

Thank you for catching this. We have edited this sentence to active
voice

> For most tags, your daily locations are from the Argos system, not
> from your data loggers. However, you included some FastLoc tag
> deployments. Please re-write for clarity of what these location are.

The previous paragraph describes how we determined daily locations. We
see the reviewers point re: the data not technicaly coming from our data
loggers. We have edited the sentence for additional clarity.

> What is a "key" weather variable? Please write for clarity.

We've edited this sentence removed the word 'key'

> This is a procedural detail. Other readers may replicate your method
> without needing to do this. Please omit this detail.

> Readers can replicate your methods without the details on converting
> the NARR wind vectors into wind speed.

We're not sure why the reviewer would recommend omitting a 'procedural
detail' as part of our methods. We feel it is important to clearly
communicated our procedure for processing the large NARR data set and
that we used the native grid size of 32km. Regarding the calculation of
wind vectors, we include this detail because many readers, like us
initially, may not be well versed in how wind vectors components are
converted to wind speed.

> Please use an active voice: "For all seal species we considered the
> following variables when modeling the hourly haulout behavior..."

We have edited this sentence with the reviewer's suggestion

> Using the term "major focus" implies that there may be numerous minor
> foci. Can you not state that this is the goal of the study? Ancillary
> natural history insights can be discussed incidental to this, but are
> not driving the research. The statutory requirements of the MMPA and
> ESA are driving your study. These Federal laws require robust methods
> to monitor population status. Your study seeks to build this method.
> Discussion of natural history belong in your Discussion section and
> not in the framing of your study. Please help the reader focus on why
> the Federal government is doing this study.

> Understanding pinniped haulout behavior is essential to designing and
> interpreting population estimation aerial surveys. Understanding how
> they use sea ice during a time when sea ice availability is rapidly
> changing is also important, but is ancillary to the paper's focus.

We have removed the term 'major focus' and, as the reviewer suggests,
simply state that this is the goal of the study. We have also added
mention of the statutory requirements to paragraph 2 of the Introduction
section and expanded the text to acknowledge the importance of
understanding how seals use sea ice during this period of change but,
ultimately, it's ancillary to the focus of this paper.

As for providing the reader focus on why the Federal government is doing
this study: this is a collaborative study across multiple organizations
and the analysis/purpose is not limited to the Federal government. We've
incorporated the reviewers suggests to expand discussion of the
statutory requirements of the MMPA and ESA but don't feel we need to
single out the motivations for the Federal government.

> Seals can just as happily be on foraging trips within the pack and
> marginal sea ice as in the open water, depending on the species.
> Suggesting that foraging trips are associated with ice-free conditions
> is misleading.

We have removed 'e.g., foraging trips' from this sentence so as not to
mislead the reader.

> This is your primary reason for not including sea ice geography in
> your analysis: you simply lacked geolocation of much of your haulout
> behavior data.
>
> Please help the reader understand how coarse your tracking data was
> for the flipper tagged animals and what proportion of each species was
> tracked with these flipper tags. Please also help the reader
> understand how the tracking data relates geographically to regions
> that may be subject to aerial survey efforts (i.e. are your tracking
> data collected from across the regions that may be surveyed during
> range-wide spring aerial surveys?).

We have re-organized this paragraph to, hopefully, better communicate our
choice not to explore sea-ice related covariates. As the reviewer suggests,
we have limited geolocation data for at-sea periods of our haul-out behavior
data -- we know the seal was in the water but don't have any estimates of
location when only a flipper tag is attached. By leading with this statement,
we hope that communicates our primary rationale. But, the other rationales
are also important and we still include them.

We have also provided some statistics regarding the percentage of the
percent-dry timeline data that is not associated with a corresponding location
estimate because the seal was in the water and no estimates were available from
flipper mounted SPOT tags.

> Please help the reader understand why you have not included a basic
> metric of sea ice geography, such as distance from marginal ice edge
> (from the NIC's daily Marginal Ice Zone charts), which is a metric
> that would be available for all data collected during an aerial
> survey.

> It is a bit misleading to indicate that the NARR variables were
> sourced from NARR alone. Their values were interpolated onto a
> geogrpahic location that was widely variable in precision and that
> precision relates to the both the tag attachment (glued on or flipper
> attached) and the seal behavior (in water, hauled out).
>
> Indicating that your "Northing" variable was from the "Bio-logger" is
> also misleading because your locations was determined by either the
> Doppler inflection point time, satellite orbitography and a model of
> the Earth (the Argos relocation method) or by reprocessing of GNSS
> signals that were logged and then transmitted via the Argos system as
> a data message. Really, neither of these were determined by the
> "Bio-logger", rather they were determined by a very complex system of
> geolocation.

> Your readers may be better served by not including this "Source"
> field, given that you describe how you obtained or estimated these
> variable values in your methods text. Offering an over-simplified
> "Source" in the table obscures your actual methods.

We appreciate the reviewer's thoughts on this and agree with that removing
the 'Source' column better serves the readers.