### 2024-08-19 analysis/paper/Letter_Response_to_Reviewers_2.qmd
response letter, round 2

---
subject: "Additional Revisions to Maunscript 2022:04:72978"
sender: |
  Josh M. London, PhD \
  7600 Sand Point Way NE . Seattle, Washington, 98115 . josh.london@noaa.gov"
recipient: |
  To: Suzanne Prange \ 
  Academic Editor \ 
  PeerJ Life & Environment
sent: "19 August 2024"
format:
  letter-typst: default
---

Dr. Prange

I sincerely appreciate the additional feedback from the reviewers and and
pleased to hear you support publication of this article in *PeerJ*. I was,
unfortunately, on leave for a portion of the past few weeks and thus was unable
to make the initial 15 day deadline. I have revised the manuscript in response
to feedback from reviewer M. Carter and have provided additional details below.
I especially appreciated Matt's suggestion for a new title and some of this
thoughts regarding the ordering of topics.

I also want to take a moment and thank the *PeerJ* staff for their previous
efforts and dedication ensuring the formatting details, accessibility, and data
access were all finalized during the previous submission.

Sincerely

Josh M. London, PhD\
Wildlife Biologist\
Alaska Fisheries Science Center, NOAA Fisheries

--------------------------------------------------------------------------------

#### Response to Comments from M Carter

The reviewer provided a useful suggestion regarding the manuscript title and the
order in which particular topics are presented. We have re-titled the manuscript
per their suggestion to *Spring haul-out behavior of seals in the Bering and
Chukchi Seas: implications for abundance estimation*. Additionally we have made
minor edits to the Abstract and Introduction in response to the reviewer's
suggestions regarding topic order.

*L59: Typo? “Trends” rather than “trend”?*

Fixed

*L111: Maybe outline for non-specialist readers that adhering bio-loggers to
hair is the conventional approach for most devices on pinnipeds.*

The text now reads:

> These devices are designed to collect data through the molt period (when those
> adhered to the hair -- a more conventional method -- would fall off) and, in
> some situations, provide multiple years of data.

*L113: Given the final sentence of the previous paragraph, I think it’s
important to detail that you’re not using a continuous 16-year dataset from one
device here, rather you have compiled data collected from multiple devices
spanning a 16-year period.*

The text now reads:

> In this study, we used data collected from multiple bio-logging deployments
> spanning a 16-year period to investigate the haul-out behavior of bearded,
> ribbon, and spotted seals in the Bering and Chukchi seas.

*L138: What is meant by “primary literature” here? I can’t see any references
associated with the deployments.*

The supplemental table lists primary literature associated with each study where
additional details regarding the capture techniques, etc can be found. I've
update the text for additional clarification:

> We refer readers to the primary literature for detailed capture and bio-logger
> attachment methods (see publications listed in Supplemental Material S1)

*L156-157: Can you clarify this? Is it that you get erroneous haulout records
where the animal is actually at sea and thus there is a bias introduced into the
percentage time hauled out estimate that is used to scale the abundance
estimates? In cases where flipper-mounted tags were not deployed, did you use
any spatial cleaning method to identify and remove these (i.e. if they are
clearly offshore according to the associated location data)? Or is this
complicated by lack of knowledge about sea ice presence?*

The text currently states *bio-loggers attached to the head or dorsal region are
often dry while the seal is floating at the surface, inducing a slight positive
bias in the hourly percent-dry values reported by the bio-logger*. We're not
sure how we could efficiently improve the specifics of this statement. Because
seals must put their heads out of the water to breath, tags attached to the head
or dorsal region will naturally spend more time out of the water compared to the
flipper. This results in a slight bias.

*L365 and 383: Perhaps interesting to note from an ontogeny perspective that the
diurnal pattern of haulout seems to develop in YOY from a very young age for
both ribbon and spotted seals.*

An additional sentence was added to highlight this

*L430: Repetition of “higher latitudes”.*

Fixed

*L463: Can you comment on the ontogeny of haulout behaviour seen in pups for
these species? It seems interesting that they adopt the diel pattern in haulout
behaviour so soon after weaning as presumably they are not moulting? Can you
unpack this in the context of existing knowledge, if any exists? If none exists,
perhaps draw attention to its novelty?*

Additional text was added to note this ontogeny.

*L474: “Both of which provide” (not “provides”).*

Fixed

*L523-524: Can you give an idea of the magnitude of this bias?*

The Conn & Trukhanova 2023 paper describes the bias in detail related to the
inclusion of age and stage so the text was edited to more directly encourage the
readers to see that paper for more details. We don't currently have the
framework to measure the magnitude of the bias but it might be possible after
we've completed updated abundance estimates from recent surveys.

*L567-569: I think this statement would be much more impactful if we had some
idea of the magnitude of bias in abundance estimates from not accounting for
these factors.*

Agreed. Unfortunately, including this and conducting the additional analysis
would exceed the expectations of 'minor revisions' and delay publication. We
will keep this in mind for future analyses and manuscripts.